### 2014-10-08 article.tex
Initial commit

% !TeX spellcheck = ru_ru_en_us
\documentclass[]{TAACpaper}

\begin{document}

%This is the comment. Further under the text features of use of style of theses are described

%We are specifying language of the paper
\selectlanguage{russian}
\def\dd#1#2{\frac{\partial#1}{\partial#2}}
\section{
%%%%%%%%%%%%%%% You have to specify name of your paper here
``О связи задачи маршрутизации с жесткими временными ограничениями с задачей о назначениях'' 
}

\authors{
Р.~А.~Шафеев
}

\abstract{
%%%%%%%%%%%%%%% You have to insert abstract of your paper here
В статье приводится метод определения решения задачи маршрутизации с учетом временных ограничений, которое можно использовать в вероятностном методе поиска с запретами как нижнюю границу для формирования окрестности потенциальных решений, что может значительно уменьшить время поиска оптимума. В случае если целевая функция не зависит от простоя автомобиля перед посещением следующего пункта назначения, то такое решение совпадает с оптимумом исходной задачи.
}

%%%%%%%%%%%%%%% To define new section use a command \subsection
% We are defining "Introduction" here
\subsection{Введение}
%%%%% Paragraph text. To separate paragraphs use empty lines.
Задача маршрутизации с жесткими временными ограничениями принадлежит к классу NP трудных\cite{VRPTW}, для решения которых широко применяются метаэвристические алгоритмы, в частности различные модификации алгоритма вероятностного поиска с запретами \cite{Tabu_search}. Данные алгоритмы позволяют находить качественные решения, однако при решении динамических задач маршрутизации, для которых время нахождения оптимума является критическим, малопригодны. Если  провести анализ исходных данных задачи, то часто  возникает ситуация, что граф имеет статическую структуру в не зависимости от временных ограничений и задачу можно решить за полиномиальное время \cite{Assignment_Problems}. Такое решение можно использовать, как нижнюю границу при поиске решения исходной задачи. А в случае, когда простой автомобиля перед посещением следующего клиента не влияет на целевую функцию, то такое решение совпадает с оптимумом.

\subsection{Постановка задачи и дискретная модель}
Пусть $C,dim(C)=n$ – множество точек, соответствующих текущему положению автотранспортных средств(АТС), $Q,dim(Q)=m$ – множество пунктов назначения. Прибытие в пункт назначения должно быть выполнено в промежутке времени $[t_j,t_j + \Delta t_j]$. Временные затраты, которые могут быть связаны с разгрузкой товаров, равны $w_j, \forall j \in Q $.

Необходимо составить оптимальные маршруты передвижения АТС для посещения всех пунктов назначения.

Зададим последовательность неизвестных матриц $\{X^k\}^n_{k=1}$, элементы  которых принимают следующие значения:
\begin{equation}
  x^{(k)}_{i,j} = 
    \begin{cases}
	  1,&\text{авт. $k$ движется из $i$-го узла в $j$-й, 
	           $i \in Q \cup \{k\}, j \in Q$ }\\
	  0,&\text{в противном случае.}
    \end{cases}
\end{equation}
Целевая функция примет следующий вид:
\begin{equation} \label{main_objective}
  F(X) = 
    \sum_{k \in C}[
     \sum_{i \in Q \cup  \{k\},j \in Q} 
     (\Omega_{i,j}+\varphi(t_j - t_i - \omega_i - t_{i,j})\cdot x^{(k)}_{i,j})
     ]
     \to min
\end{equation}
При ограничениях:
\begin{align} 
& \sum_{k \in C}\sum_{j \in Q}x^{(k)}_{i,j} = 1, 
  \forall i \in C \cup Q, \forall k \in C \label{main_cond_1}\\ 
& \sum_{j \in Q} ( 
       x^{(k)}_{k,j} -x^{(k)}_{i,j} ) \geq 0, 
       \forall i \in Q ,  \forall k \in C \label{main_cond_2}\\
& \sum_{i \in Q \cup  \{k\}} x^{(k)}_{i,\omega} - 
  \sum_{j \in Q} x^{(k)}_{\omega,j} = 0, 
  \forall \omega \in Q ,  \forall k \in C \label{main_cond_3}\\
&  \sum_{i \in S}\sum_{j \notin S}x^{(k)}_{i,j} > 0, 
  S=\{s \in Q: \sum_{j \in Q \cup C}x^{(k)}_{j,s}>0 \}  ,\forall k \in C \label{main_cond_4}\\
& t_j \leq \tilde{t_j} \leq t_j + \Delta t_j, \forall j \in Q, 
\tilde{t_j} \text {- время прибытия в j-й пункт назначения} \label{main_cond_5}
\end{align}

\subsection{Анализ исходных данных}
В случае, когда простой автомобиля перед посещением следующего клиента не влияет на целевую функцию, а именно $\varphi=const=0$, тогда в некоторых случаях временные ограничения могут не влиять на структуру графа и задача сводится к задаче о назначениях. В силу того, что $NP \neq P$, однозначный переход невозможен, но из-за наличия временных ограничений граф задачи может не иметь локальных циклов и в этом случае  ограничения (\ref{main_cond_4}) и (\ref{main_cond_5}) возможно опустить после удаления дуг $(i,j)$ по которым невозможно достижение вершины $j$ в не зависимости от времени прибытия в вершину $i$.

Для того, чтобы проверить возможность такого перехода, введем дополнительные величины. 
\begin{enumerate}
\item Учитывая наложенные на целевую функцию ограничения, время прибытия в $j$–й пункт назначения равно:
\begin{equation} \label{t_j_input}
   \exists h_i \in [0,1] : \tilde{t_i} = t_i + h_i \cdot \Delta t_i
\end{equation}

\item Введем матрицу $U$, элементы которой показывают в количественной форме нарушение временного окна:
\begin{equation} \label{matrix_U}
 U(\vec{h}) : u_{i,j} = 
    \begin{cases}
	  0, i=j, \\
      t_i + h_i \cdot \Delta t_i + \omega_i + t_{i,j} - t_j - \Delta t_j, i \neq j
    \end{cases}   
\end{equation}
\end{enumerate}

Можно сказать, что если элемент матрицы $U(\vec{h}=\vec{1})$ отрицательный, то передвижение между пунктами возможно выполнить в не зависимости от временных ограничений. Поэтому построим граф $G$ для задачи маршрутизации, в котором уже будут учтены ограничения (\ref{main_cond_4}) и (\ref{main_cond_5}), задав следующую матрицу инцидентности:
\begin{equation} \label{matrix_I}
 I(\vec{h}) : I_{i,j} = 
    \begin{cases}
	  0, &\text{если $u_{i,j} \geq 0$,} \\
      1, &\text{если $u_{i,j}<0$}
    \end{cases}, i \in C \cup Q, j \in Q
\end{equation}
\newtheorem{Def}{Определение} 
\begin{Def}\label{def_main}
Если:
  \begin{enumerate}
  \item $\varphi=const=0$;
  \item матрица $I(\vec{0})$ равна матрице $I(\vec{1})$;
  \item граф $G$, построенный по матрице инцидентности $I(\vec{0})$ не является мультиграфом,
  \end{enumerate}
 тогда задача маршрутизации с жесткими ограничениями по времени сводится к задаче о назначениях.
\end{Def}
Это утверждение вытекает из следующих лемм.

\newtheorem{Lem}{Лемма}
\begin{Lem}
Если матрица $I(\vec{0}) = I(\vec{1})$, тогда $I(\vec{0})=I(\vec{h}), \forall \vec{h}: h_i \in [0,1]$
\end{Lem}
Доказательство. 
Элементы матрицы $I_{i,j}$ не зависят от вектора $\vec{h}$ при $i \in C, j \in Q$, так как вершины из множества C соответствуют местоположениям автомобилей в начальный момент времени и им не заданы временные окна ($\Delta t_i = 0, \forall i \in C$). Остальные элементы матрицы можно представить в виде формулы:
\begin{equation} 
I_{i,j}(h_i) = \frac{1 - sign(u_{i,j}(h_i))}{2}
\end{equation}
По условию $I(\vec{0}) = I(\vec{1})$, тогда:

\begin{flalign*}
&\frac{1 - sign(u_{i,j}(0))}{2} =  \frac{1 - sign(u_{i,j}(1))}{2} \Rightarrow
sign(u_{i,j}(0)) = sign(u_{i,j}(1))&
\end{flalign*}
Или:
\begin{flalign*}
&\left[\begin{array}{l}
     \begin{cases}
	  u_{i,j}(0)>0  \\
      u_{i,j}(1)>0
    \end{cases}\\
    \begin{cases}
	  u_{i,j}(0)<0  \\
      u_{i,j}(1)<0
    \end{cases}
\end{array}\right.&
\end{flalign*}

Функция $u_{i,j}(h_i)$ на промежутке $h_i \in [0,1]$ является монотонно возрастающей, потому что производная $\dd{u_{i,j}(h_i)}{h_i} = \Delta t_i >0$ положительна на всем промежутке. Поэтому:

\begin{flalign*}
&\left[\begin{array}{l}
     \begin{cases}
	  u_{i,j}(0)>0  \\
      u_{i,j}(1)>0
    \end{cases} \Rightarrow u_{i,j}(h_i)>0 \\
    \begin{cases}
	  u_{i,j}(0)<0  \\
      u_{i,j}(1)<0
    \end{cases} \Rightarrow u_{i,j}(h_i)<0
\end{array}\right.
\Rightarrow sign(u_{i,j}(0)) = sign(u_{i,j}(h_i)), \forall h_i \in [0,1] &
\end{flalign*}
Тогда $sign(u_{i,j}(0)) = sign(u_{i,j}(h_i)) \Rightarrow  I(\vec{h})=I(\vec{0})$. Доказано.

\newtheorem{Lem2}{Лемма}
\begin{Lem}
Если граф  $G$, построенный по матрице инцидентности $I(\vec{h}=const)$ не является мульграфом, то такой граф не имеет циклов.
\end{Lem}
Доказательство(от противного). Пусть дан граф $G(V={C \cup Q}, E)$, построенный по матрице инцидентности  $I(\vec{h}=const)$ и в нем имеется цикл: \begin{align*}
&E_{cicle}=\{(v_{s_1},v_{s_2}),(v_{s_2},v_{s_3}),...,(v_{s_{k-1}},v_{s_k}),(v_{s_k},v_{s_1})\}
\end{align*}
Предположим, что при наличии цикла $E_{cicle}$ в графе $G$ нет дуги $(v_{s_1},v_{s_k})$, т.е. АТС не успеет добраться из вершины $v_{s_1}$ в вершину $v_{s_k}$ не нарушив временные ограничения. Поэтому:
\begin{equation} \label{error_inequality}
 t_{s_k} + \Delta t_{s_k} <  \tilde{t}_{s_1} + \omega_{s_1} + t_{s_1,s_k}
\end{equation}
Согласно предположению о существовании цикла $E_{cicle}$ верно следующее неравенство ( $E_{route}=E_{cicle} / \{(v_{s_k},v_{s_1}) \}$):
\begin{equation} \label{time_s_k_inequality}
 t_{s_k} + \Delta t_{s_k} > \tilde{t}_{s_{k-1}} + \omega_{s_{k-1}} + t_{s_{k-1},s_k} = \tilde{t}_{s_1} +  \sum_{(i,j) \in E_{route}} (t_{i,j} +  \omega_{s_i}) = A
\end{equation}
Пусть передвижение от вершины $v_{s_1}$ в вершину $v_{s_k}$ будет выполняться по дугам $E_{route}$. Тогда:
\begin{equation} \label{time_route}
 \tilde{t}_{s_1} + \omega_{s_1} + t_{s_1,s_k} \leq \tilde{t}_{s_1} + \omega_{s_1} + \sum_{(i,j) \in E_{route}} t_{i,j} \leq A
\end{equation}
Подставим (\ref{time_route}) в (\ref{time_s_k_inequality}):
\begin{equation} \label{true_inequality}
 t_{s_k} + \Delta t_{s_k} > \tilde{t}_{s_1} + \omega_{s_1} + t_{s_1,s_k}
\end{equation}
Пришли к противоречию, следовательно, неравенство (\ref{error_inequality}) неверно. Доказано.

\subsection{Переход к задаче о назначениях}
Если выполняются правила из определения \ref{def_main}, тогда можно свести исходную задачу к задаче о назначениях.

Введем замену переменных X :
\begin{equation} \label{y_replace}
\forall k \in C: ~Y^k = \{ I_{i,j}(0) \cdot x^{(k)}_{i,j} , i \in Q \cup  \{k\} , j \in Q  \}
\end{equation}

Теперь задачу можно сформулировать таким образом, что ее можно разрешить за полиномиальное время:
\begin{align} 
  &F(Y) = 
    \sum_{k \in C}
     \sum_{i \in Q \cup  \{k\},j \in Q} 
     \Omega_{i,j}\cdot y^{(k)}_{i,j})
     \to min\\
& \sum_{k \in C}\sum_{j \in Q}y^{(k)}_{i,j} = 1, 
  \forall i \in C \cup Q, \forall k \in C \label{mod_cond_1}\\ 
& \sum_{j \in Q} ( 
       y^{(k)}_{k,j} -y^{(k)}_{i,j} ) \geq 0, 
       \forall i \in Q ,  \forall k \in C \label{mod_cond_2}\\
& \sum_{j \in Q \cup  \{k\}} y^{(k)}_{i,\omega} - 
  \sum_{i \in Q} y^{(k)}_{\omega,j} = 0, 
  \forall \omega \in Q ,  \forall k \in C \label{mod_cond_3}
\end{align}

Чтобы получить задачу о назначениях в классическом ее виде, построим на основе графа $G$ двудольный граф $G_b(V_{begin},V_{end},E_b)$ по следующему принципу:
\begin{enumerate}
\item $V_{begin} = \{v \in C \cup Q\ : \exists (v,\omega) \in E, \omega \in Q \}$ - состоит из множества вершин, для которых найдется хотя бы одна выходящая дуга из множества E;
\item $V_{end} = \{v  \in Q\ : \exists (\omega,v) \in E, \omega \in C \cup Q \}$ - состоит из множества вершин, для которых найдется хотя бы одна входящая дуга из множества E;
\item $E_b = \{(v,\omega) \in E: v \in V_{begin}, \omega \in V_{end} \}$.
\end{enumerate}

\begin{figure}[h]
\hfil\includegraphics[height=2.0in]{sign_graph}\hfil
\caption
{
  Пример преобразования графа $G$ в двудольный граф $G_b$.
}
\label{aba:fig1}
\end{figure}

\subsection{Результаты}
Преобразованную задачу(рис. \ref{aba:fig1}) рекомендуется решать с помощью метода Голдберга и Кеннеди,  основанного на технике масштабирования при переходе к задаче поиска потока минимальной стоимости \cite{Goldberg_Kennedy}. Сложность алгоритма равна $O(\sqrt{n}mlog(nC))$.

В случае, если функция $\varphi(t)$ присутствует в целевой функции, то решение преобразованной задачи может послужить нижней границей, которую можно использовать для отсеивания заведомо худших решений при формировании окрестности вокруг текущего решения в работе вероятностного алгоритма поиска с запретами.

Проверка метода была выполнена на тестовых задачах Кристофидеса, Голдберга  и Тейларда с временными ограничениями, удовлетворяющими условиям, которые были изложены в определении \ref{def_main} \cite{problems_Christofides,problems_Golden,problems_Taillard}.

\subsection{Выводы}
Предложены правила перехода от задачи маршрутизации с учетом временных ограничений к задаче о назначениях. Данную технику целесообразно использовать для задач маршрутизации, в которых не учитывается простой автомобиля. Также  метод можно использовать для инициализации метаэвристического алгоритма или в виде более качественной нижней границы для определения оптимума.

\begin{thebibliography}{99}
%In brackets we writes the name which is used for referencing.

\bibitem{VRPTW} T.~Babb, Pickup and Delivery Problem with Time Windows // Coordinated Transportation Systems: The State of the Art. Department of Computer Science University of Central Florida Orlando, Florida, 2005, 38 p.

\bibitem{Tabu_search} O.~Braysy, M.~Gendreau, Vehicle Routing Problem with Time Windows, Part I: Route Constuction and local algorithms // Transportation science Vol.39 No. 1, 2005, p. 104-118.

\bibitem{Assignment_Problems} E.~Rainer, Assignment problems, 2009, 402 p.

\bibitem{Goldberg_Kennedy} V.~Goldberg, R.~Kennedy, An Efficient cost scaling algotirhm for the assignment problem, Math. Program., 1995, p. 153--177.

\bibitem{problems_Christofides}  N. Christofides, S. Eilon, An algorithm for the vehicle dispatching problem //Operational Research Quarterly, 20, 1969, p. 309–318.
\bibitem{problems_Golden}  B. Golden, E. Wasil, J. Kelly, I-M. Chao. The impact of metaheuristics on solving the vehicle routing problem: Algorithms, problem sets, and computational results. In T. Crainic and G. Laporte, editors // Fleet Management and Logistics, Kluwer, Boston, 1998 p. 33–56.
\bibitem{problems_Taillard}  E. Taillard. VRP benchmarks.\\ http://mistic.heig-vd.ch/taillard/problemes.dir/vrp.dir/vrp.html, 1993.
\end{thebibliography}

\subsection{Автор}
\author{Роман Александрович Шафеев}{магистр 2-го года обучения, факультет информатики и управления, Национальный технический университет <<Харьковский политехнический институт>>, Харьков, Украина}
{rs@premiumgis.com}

\end{document}

### 2014-10-08 article.tex
txs auto checkin

@@
 \subsection{Постановка задачи и дискретная модель}
-Пусть $C,dim(C)=n$ – множество точек, соответствующих текущему положению автотранспортных средств(АТС), $Q,dim(Q)=m$ – множество пунктов назначения.
-Прибытие в пункт назначения должно быть выполнено в промежутке времени $[t_j,t_j + \Delta t_j]$.
-Временные затраты, которые могут быть связаны с разгрузкой товаров, равны $w_j, \forall j \in Q $.
+Пусть $C,dim(C)=n$ – множество точек, соответствующих текущему положению автотранспортных средств(АТС), $Q,dim(Q)=m$ – множество пунктов назначения. $D$ $dim(D)=k$ - множество складов.\\
+Пусть $V=C\cup{Q}\cup{D}$.
+$P_i(t)={lat,lon}$, где $i\inV$ - местоположением $i$-го узла в момент времени $t$.\\
+$S_k,k\inC$ - грузоподъемность АТС;
+$t_{A}^{j}$ - время поступления $j$-ой заявки.\\
+$\bigtriangleup\tilde{t}_{\omega}^{j}$-время обработки $j$-ой заявки на складе (прогнозируемое время обработки)\\
+$\tilde{t}^{j}$-время доставки груза $j$-му клиенту (прогнозируемое время, настроечное)\\
+$depot_j$-склад, на котором находится груз $j$-го клиента\\
+$\nu_j$-груз, который необходимо доставить $j$-му клиенту.
 

### 2014-10-08 article.tex
txs auto checkin

--- 2014-10-08 article.tex
@@
 \subsection{Постановка задачи и дискретная модель}
-Пусть $C,dim(C)=n$ – множество точек, соответствующих текущему положению автотранспортных средств(АТС), $Q,dim(Q)=m$ – множество пунктов назначения.
-Прибытие в пункт назначения должно быть выполнено в промежутке времени $[t_j,t_j + \Delta t_j]$.
-Временные затраты, которые могут быть связаны с разгрузкой товаров, равны $w_j, \forall j \in Q $.
+Пусть $C,dim(C)=n$ – множество точек, соответствующих текущему положению автотранспортных средств(АТС), $Q,dim(Q)=m$ – множество пунктов назначения. $D$ $dim(D)=k$ - множество складов.\\
+Пусть $V=C\cup{Q}\cup{D}$.
+$P_i(t)={lat,lon}$, где $i\inV$ - местоположением $i$-го узла в момент времени $t$.\\
+$S_k k\inC$ - грузоподъемность АТС;
+$t_{A}^{j}$ - время поступления $j$-ой заявки.\\
+$\bigtriangleup\tilde{t}_{\omega}^{j}$-время обработки $j$-ой заявки на складе (прогнозируемое время обработки)\\
+$\tilde{t}^{j}$-время доставки груза $j$-му клиенту (прогнозируемое время, настроечное)\\
+$depot_j$-склад, на котором находится груз $j$-го клиента\\
+$\nu_j$-груз, который необходимо доставить $j$-му клиенту.
 

### 2014-10-08 article.tex
txs auto checkin

--- 2014-10-08 article.tex
@@
 \subsection{Постановка задачи и дискретная модель}
-Пусть $C,dim(C)=n$ – множество точек, соответствующих текущему положению автотранспортных средств(АТС), $Q,dim(Q)=m$ – множество пунктов назначения.
-Прибытие в пункт назначения должно быть выполнено в промежутке времени $[t_j,t_j + \Delta t_j]$.
-Временные затраты, которые могут быть связаны с разгрузкой товаров, равны $w_j, \forall j \in Q $.
+Пусть $C,dim(C)=n$ – множество точек, соответствующих текущему положению автотранспортных средств(АТС), $Q,dim(Q)=m$ – множество пунктов назначения. $D$ $dim(D)=k$ - множество складов.\\
+Пусть $V=C\cup{Q}\cup{D}$.
+$P_i(t)={lat,lon}$, где $i\inV$ - местоположением $i$-го узла в момент времени $t$.\\
+$S_k$ $k\inC$ - грузоподъемность АТС;
+$t_{A}^{j}$ - время поступления $j$-ой заявки.\\
+$\bigtriangleup\tilde{t}_{\omega}^{j}$-время обработки $j$-ой заявки на складе (прогнозируемое время обработки)\\
+$\tilde{t}^{j}$-время доставки груза $j$-му клиенту (прогнозируемое время, настроечное)\\
+$depot_j$-склад, на котором находится груз $j$-го клиента\\
+$\nu_j$-груз, который необходимо доставить $j$-му клиенту.
 

### 2014-10-08 article.tex
txs auto checkin

--- 2014-10-08 article.tex
@@
 $P_i(t)={lat,lon}$, где $i\inV$ - местоположением $i$-го узла в момент времени $t$.\\
-$S_k k\inC$ - грузоподъемность АТС;
+$S_k$  - грузоподъемность АТС;
 $t_{A}^{j}$ - время поступления $j$-ой заявки.\\

### 2014-10-08 article.tex
txs auto checkin

@@
 Пусть $V=C\cup{Q}\cup{D}$.
-$P_i(t)={lat,lon}$, где $i\inV$ - местоположением $i$-го узла в момент времени $t$.\\
+$P_i(t)={lat,lon}$, где - местоположением $i$-го узла в момент времени $t$.\\
 $S_k$  - грузоподъемность АТС;

### 2014-10-08 article.tex
txs auto checkin

@@
 Пусть $C,dim(C)=n$ – множество точек, соответствующих текущему положению автотранспортных средств(АТС), $Q,dim(Q)=m$ – множество пунктов назначения. $D$ $dim(D)=k$ - множество складов.\\
-Пусть $V=C\cup{Q}\cup{D}$.
+Пусть $V=C\cup{Q}\cup{D}$:\\
 $P_i(t)={lat,lon}$, где - местоположением $i$-го узла в момент времени $t$.\\

### 2014-10-08 article.tex
txs auto checkin

--- 2014-10-08 article.tex
@@
 Пусть $C,dim(C)=n$ – множество точек, соответствующих текущему положению автотранспортных средств(АТС), $Q,dim(Q)=m$ – множество пунктов назначения. $D$ $dim(D)=k$ - множество складов.\\
-Пусть $V=C\cup{Q}\cup{D}$.
+Пусть $V=C\cup{Q}\cup{D}$:\\
 $P_i(t)={lat,lon}$, где - местоположением $i$-го узла в момент времени $t$.\\
-$S_k$  - грузоподъемность АТС;
+$S_k$, где $k\in{C}$ - грузоподъемность АТС;\\
 $t_{A}^{j}$ - время поступления $j$-ой заявки.\\

### 2014-10-08 article.tex
txs auto checkin

--- 2014-10-08 article.tex
@@
 Пусть $C,dim(C)=n$ – множество точек, соответствующих текущему положению автотранспортных средств(АТС), $Q,dim(Q)=m$ – множество пунктов назначения. $D$ $dim(D)=k$ - множество складов.\\
-Пусть $V=C\cup{Q}\cup{D}$.
-$P_i(t)={lat,lon}$, где - местоположением $i$-го узла в момент времени $t$.\\
-$S_k$  - грузоподъемность АТС;
+Пусть $V=C\cup{Q}\cup{D}$:\\
+$P_i(t)={lat,lon}$, где $i\in{V}$- местоположением $i$-го узла в момент времени $t$.\\
+$S_k$, где $k\in{C}$ - грузоподъемность АТС;\\
 $t_{A}^{j}$ - время поступления $j$-ой заявки.\\

### 2014-10-08 article.tex
txs auto checkin

--- 2014-10-08 article.tex
@@
 Пусть $V=C\cup{Q}\cup{D}$:\\
-$P_i(t)={lat,lon}$, где - местоположением $i$-го узла в момент времени $t$.\\
-$S_k$  - грузоподъемность АТС;
+$P_i(t)={lat,lon}$, где $i\in{V}$- местоположением $i$-го узла в момент времени $t$.\\
+$S_k$, где $k\in{C}$ - грузоподъемность АТС;\\
 $t_{A}^{j}$ - время поступления $j$-ой заявки.\\
-$\bigtriangleup\tilde{t}_{\omega}^{j}$-время обработки $j$-ой заявки на складе (прогнозируемое время обработки)\\
-$\tilde{t}^{j}$-время доставки груза $j$-му клиенту (прогнозируемое время, настроечное)\\
-$depot_j$-склад, на котором находится груз $j$-го клиента\\
+$\bigtriangleup\tilde{t}_{\omega}^{j}$-время обработки $j$-ой заявки на складе (прогнозируемое время обработки);\\
+$\tilde{t}^{j}$-время доставки груза $j$-му клиенту (прогнозируемое время, настроечное);\\
+$depot_j$-склад, на котором находится груз $j$-го клиента;\\
 $\nu_j$-груз, который необходимо доставить $j$-му клиенту.

### 2014-10-13 article.tex
txs auto checkin

@@
 	  1,&\text{авт. $k$ движется из $i$-го узла в $j$-й, 
-	           $i \in Q \cup \{k\}, j \in Q$ }\\
+	           $i\in{C}\cup{Q}\cup{D}, j \in Q\cup{D}$ }\\
+	  0,&\text{в противном случае.}
+    \end{cases}
+\end{equation}
+\begin{equation}
+	x^{(k)}_{j} = 
+	\begin{cases}
+		1,&\text{авт. $k$-ая машина взяла груз для $j$-го клиента, $i\in{C}\cup{Q}\cup{D}, j \in Q\cup{D}$ }\\
 		0,&\text{в противном случае.}
@@
   F(X) = 
-    \sum_{k \in C}[
-     \sum_{i \in Q \cup  \{k\},j \in Q} 
-     (\Omega_{i,j}+\varphi(t_j - t_i - \omega_i - t_{i,j})\cdot x^{(k)}_{i,j})
-     ]
+    \sum_{k \in C}
+     \sum_{i,j\cup{V}} 
+     x_{ij}^{k}\cdot\Omega_{ij}+\sum_{j\in{Q}}\tau(j)
      \to min
@@
 \begin{align} 
+& \sum_{i\in{V}}x_{ij}^{k}=y_{j,d}^{k}, \forall{k}\in{C}, \forall{j}\in{Q},\forall{d}\in{D}\\
+& sign[\sum_{j\in{Q}}y_{j,d}^{k}]=\sum_{i\in{V}}x_{i,d}^{k}, \forall{k}\in{C}, \forall{d}\in{D}\\
+& y_{j,d}^{k}(\tilde{t}_{j}^{k}-\tilde{t}_{d}^{k})\ge{0}\\
 & \sum_{k \in C}\sum_{j \in Q}x^{(k)}_{i,j} = 1, 

### 2014-11-24 article.tex
write problem definition

@@
 
-\subsection{Постановка задачи и дискретная модель}
-Пусть $C,dim(C)=n$ – множество точек, соответствующих текущему положению автотранспортных средств(АТС), $Q,dim(Q)=m$ – множество пунктов назначения. $D$ $dim(D)=k$ - множество складов.\\
-Пусть $V=C\cup{Q}\cup{D}$:\\
-$P_i(t)={lat,lon}$, где $i\in{V}$- местоположением $i$-го узла в момент времени $t$.\\
-$S_k$, где $k\in{C}$ - грузоподъемность АТС;\\
-$t_{A}^{j}$ - время поступления $j$-ой заявки.\\
-$\bigtriangleup\tilde{t}_{\omega}^{j}$-время обработки $j$-ой заявки на складе (прогнозируемое время обработки);\\
-$\tilde{t}^{j}$-время доставки груза $j$-му клиенту (прогнозируемое время, настроечное);\\
-$depot_j$-склад, на котором находится груз $j$-го клиента;\\
-$\nu_j$-груз, который необходимо доставить $j$-му клиенту.
+\subsection{Постановка задачи}
+Пусть $C,dim(C)=n$ – множество транспортных средств, $Q,dim(Q)=m$ – множество заявок, полученных от клиентов на текущий момент времени.
 
-Необходимо составить оптимальные маршруты передвижения АТС для посещения всех пунктов назначения.
+Предположим, что известна следующая информация о транспортных средствах из множества $C$: \\
+$P_c$ -- местоположение транспортного средства $c \in C$;\\
+$L_c$ -- грузоподъемность транспортного средства $c \in C$.
+
+Пусть $S$ -- множество отправителей, $dim(S) = dim(Q) = m$), $R$ -- множество получателей, $dim(R) = dim(Q) = m$. Тогда каждая клиентская заявка $q \in Q$ включает в себя следующую информацию:\\
+$s_q$ -- отправитель груза, $s \in S$; \\
+$r_q$ -- получатель груза, $r \in R$; \\
+$P_{s_q}$ -- местоположение отправителя груза;\\
+$P_{r_q}$ -- местоположение получателя груза; \\
+$w_q$ -- груз, который требуется доставить от отправителя к получателю; \\
+$[t_{s}^{q}, t_{s}^{q}+ \Delta{t_{s}^{q}}]$ -- временное окно, в течении которого требуется забрать груз у отправителя;\\ 
+$[t_{r}^{q}, t_{r}^{q}+ \Delta{t_{r}^{q}}]$ -- временное окно, в течении которого требуется доставить груз получателю.
+
+Таким образом, каждую заявку можно представить в виде кортежа:
+\begin{equation}
+\forall q \in Q: \exists q = (s_q,r_q, P_{s_q}, P_{r_q}, w_q, t_{s}^{q}, \Delta{t_{s}^{q}}, t_{r}^{q}, \Delta{t_{r}^{q}})
+\end{equation}
+Для оценки стоимости передвижения между пунктами назначения определена функция стоимости $\Omega$: 
+\begin{equation}
+\forall i,j \in S \cup R: \exists \Omega_{i,j} = \Omega(P_i,P_j)
+\end{equation}
+
+Необходимо составить оптимальные маршруты передвижения транспортных средств для перевоза грузов от отправителя к получателю для всех клиентских заявок.
 
@@
 \end{equation}
-\begin{equation}
-	x^{(k)}_{j} = 
-	\begin{cases}
-		1,&\text{авт. $k$-ая машина взяла груз для $j$-го клиента, $i\in{C}\cup{Q}\cup{D}, j \in Q\cup{D}$ }\\
-		0,&\text{в противном случае.}
-	\end{cases}
-\end{equation}
+
 Целевая функция примет следующий вид:

### 2014-11-30 article.tex
delete pdf file
write problem statement

--- 2014-10-08 article.tex
@@
-% !TeX spellcheck = ru_ru_en_us
+% !TeX spellcheck = en_US
 \documentclass[]{TAACpaper}
-
+\usepackage[linesnumbered,boxed]{algorithm2e}
 \begin{document}
+each code-block optionally)
 
@@
 %We are specifying language of the paper
-\selectlanguage{russian}
+\selectlanguage{english}
 \def\dd#1#2{\frac{\partial#1}{\partial#2}}
@@
 %%%%%%%%%%%%%%% You have to specify name of your paper here
-``О связи задачи маршрутизации с жесткими временными ограничениями с задачей о назначениях'' 
+``Investigation of tuning parameters and modification of tabu search algorithm for solving the Static Routing Courier Delivery Problem.'' 
 }
@@
 \authors{
-Р.~А.~Шафеев
+R.~Shafeyev , L.~Lyubchik 
 }
@@
 %%%%%%%%%%%%%%% You have to insert abstract of your paper here
-В статье приводится метод определения решения задачи маршрутизации с учетом временных ограничений, которое можно использовать в вероятностном методе поиска с запретами как нижнюю границу для формирования окрестности потенциальных решений, что может значительно уменьшить время поиска оптимума.
-В случае если целевая функция не зависит от простоя автомобиля перед посещением следующего пункта назначения, то такое решение совпадает с оптимумом исходной задачи.
+
+In the paper ....
 }
@@
 % We are defining "Introduction" here
-\subsection{Введение}
+\subsection{Introduction}
 %%%%% Paragraph text. To separate paragraphs use empty lines.
-Задача маршрутизации с жесткими временными ограничениями принадлежит к классу NP трудных\cite{VRPTW}, для решения которых широко применяются метаэвристические алгоритмы, в частности различные модификации алгоритма вероятностного поиска с запретами \cite{Tabu_search}.
-Данные алгоритмы позволяют находить качественные решения, однако при решении динамических задач маршрутизации, для которых время нахождения оптимума является критическим, малопригодны.
-Если  провести анализ исходных данных задачи, то часто  возникает ситуация, что граф имеет статическую структуру в не зависимости от временных ограничений и задачу можно решить за полиномиальное время \cite{Assignment_Problems}.
-Такое решение можно использовать, как нижнюю границу при поиске решения исходной задачи.
-А в случае, когда простой автомобиля перед посещением следующего клиента не влияет на целевую функцию, то такое решение совпадает с оптимумом.
+ The VRP \cite{VRPTW} ...
 
-\subsection{Постановка задачи и дискретная модель}
-Пусть $C,dim(C)=n$ – множество точек, соответствующих текущему положению автотранспортных средств(АТС), $Q,dim(Q)=m$ – множество пунктов назначения. $D$ $dim(D)=k$ - множество складов.\\
-Пусть $V=C\cup{Q}\cup{D}$:\\
-$P_i(t)={lat,lon}$, где $i\in{V}$- местоположением $i$-го узла в момент времени $t$.\\
-$S_k$, где $k\in{C}$ - грузоподъемность АТС;\\
-$t_{A}^{j}$ - время поступления $j$-ой заявки.\\
-$\bigtriangleup\tilde{t}_{\omega}^{j}$-время обработки $j$-ой заявки на складе (прогнозируемое время обработки);\\
-$\tilde{t}^{j}$-время доставки груза $j$-му клиенту (прогнозируемое время, настроечное);\\
-$depot_j$-склад, на котором находится груз $j$-го клиента;\\
-$\nu_j$-груз, который необходимо доставить $j$-му клиенту.
+\subsection{Problem Statement}
+Let $C,dim(C)=n$ is a set of vehicles and $Q,dim(Q)=m$ is a set of requests received from customers at the current time.
 
-Необходимо составить оптимальные маршруты передвижения АТС для посещения всех пунктов назначения.
+Suppose that the following information is known about vehicles from set $C$: \\
+$\vec{P}_c$ -- vehicle position, $c \in C$;\\
+$L_c$ -- vehicle capacity, $c \in C$.
 
-Зададим последовательность неизвестных матриц $\{X^k\}^n_{k=1}$, элементы  которых принимают следующие значения:
+Let $S$ is a set of senders, ($dim(S) = dim(Q) = m$), $R$ is a set of receivers, $dim(R) = dim(Q) = m$. Then each client request $q \in Q$ includes the following information:\\
+$s_q$ -- a sender of client shipment, $s \in S$; \\
+$r_q$ -- a receiver of client shipment, $r \in R$; \\
+$\vec{P}_{s_q}$ -- sender position;\\
+$\vec{P}_{r_q}$ -- receiver position; \\
+$w_q$ -- client shipment that is required to deliver from the sender to the recipient; \\
+$[t_{s}^{q}, t_{s}^{q}+ \Delta{t_{s}^{q}}]$ -- the time window within which worker must to pick up the goods from the sender;\\ 
+$[t_{r}^{q}, t_{r}^{q}+ \Delta{t_{r}^{q}}]$ -- the time window within which worker must deliver the goods to the receiver.
+
+Thus, each application can be represented as a tuple:
+\begin{equation}
+\forall q \in Q: \exists q = (s_q,r_q, \vec{P}_{s_q}, \vec{P}_{r_q}, w_q, t_{s}^{q}, \Delta{t_{s}^{q}}, t_{r}^{q}, \Delta{t_{r}^{q}})
+\end{equation}
+To estimate the cost of transportation between destinations defined cost function $\Omega$:
+\begin{equation}
+\forall i,j \in S \cup R: \exists \Omega_{i,j} = \Omega(\vec{P}_i,\vec{P}_j)
+\end{equation}
+
+Necessary to construct the optimal routes of vehicles movement for the transportation of goods from the sender to the receiver for all client requests.
+
+\subsection{Discrete model}
+Routing problem can be represented as a directed graph $G=G(V,E)$.
+The set $V=C\cup{S}\cup{R}$ -- nodes of the graph $G$, consisting of elements of vehicles, senders and receivers.
+$E$ -- dynamic set of arcs of the graph $G$, such that:
+\begin{equation}
+\forall e(X) \in E: e(X) = (v_i,v_j), \exists v_i \in V, v_j \in V/C
+\end{equation}
+
+Let $\{X^k\}^n_{k=1}$ is a sequence of variable matrices for each vehicle $k \in C$. Elements of the matrices take the following values:
 \begin{equation}
@@
     \begin{cases}
-	  1,&\text{авт. $k$ движется из $i$-го узла в $j$-й, 
-	           $i \in Q \cup \{k\}, j \in Q$ }\\
-	  0,&\text{в противном случае.}
+	  1,&\text{the vehicle $k \in C$ moves from $i$ node to the $j$}\\
+	  0,&\text{otherwise.}
     \end{cases}
 \end{equation}
-Целевая функция примет следующий вид:
+where: $i\in{V/C \cup {k}}, j \in V/C$.
+
+Let us introduce the vector  of variables $ \vec{Y}  ^ k(X) $ for each vehicle $ k \ in C $. Vector elements have the following values:
+\begin{equation}
+\vec{y}^{(k)}_{j}(X) = 
+\begin{cases}
+1,&\text{the request $j \in Q$ is processed by vehicle $k \in C$}\\
+0,&\text{otherwise.}
+\end{cases}
+\end{equation}
+where: $i\in{V/C \cup {k}}, j \in V/C$.
+
+Let $t^k_j(X^{(k)})$ -- arrival time of the vehicle $k \in C$ at the destination $j \in S \cup R$.
+
+The objective function takes the following form:
 \begin{equation} \label{main_objective}
   F(X) = 
-    \sum_{k \in C}[
-     \sum_{i \in Q \cup  \{k\},j \in Q} 
-     (\Omega_{i,j}+\varphi(t_j - t_i - \omega_i - t_{i,j})\cdot x^{(k)}_{i,j})
-     ]
+    \sum_{k \in C}
+     \sum_{i,j\cup{V}} 
+     \Omega_{ij} \cdot x_{ij}^{(k)} 
      \to min
 \end{equation}
-При ограничениях:
+
+We define constraints on the objective function (\ref{main_objective}), which provided the continuity of routes:
 \begin{align} 
-& \sum_{k \in C}\sum_{j \in Q}x^{(k)}_{i,j} = 1, 
-  \forall i \in C \cup Q, \forall k \in C \label{main_cond_1}\\ 
-& \sum_{j \in Q} ( 
-       x^{(k)}_{k,j} -x^{(k)}_{i,j} ) \geq 0, 
-       \forall i \in Q ,  \forall k \in C \label{main_cond_2}\\
-& \sum_{i \in Q \cup  \{k\}} x^{(k)}_{i,\omega} - 
-  \sum_{j \in Q} x^{(k)}_{\omega,j} = 0, 
-  \forall \omega \in Q ,  \forall k \in C \label{main_cond_3}\\
-&  \sum_{i \in S}\sum_{j \notin S}x^{(k)}_{i,j} > 0, 
-  S=\{s \in Q: \sum_{j \in Q \cup C}x^{(k)}_{j,s}>0 \}  ,\forall k \in C \label{main_cond_4}\\
-& t_j \leq \tilde{t_j} \leq t_j + \Delta t_j, \forall j \in Q, 
-\tilde{t_j} \text {- время прибытия в j-й пункт назначения} \label{main_cond_5}
+& \sum_{k \in C}\sum_{j \in S \cup R}x^{(k)}_{i,j} \leq 1, 
+\forall i \in V \label{main_cond_1}\\
+& \sum_{k \in C}\sum_{i \in S \cup R \cup \{k\} } x^{(k)}_{i,j} = 1, 
+\forall j \in S \cup R \label{main_cond_2}\\
+& \sum_{i \in S \cup R \cup \{k\} } x^{(k)}_{i,\omega} - 
+\sum_{j \in S \cup R} x^{(k)}_{\omega,j} \leq 1, 
+\forall \omega \in S \cup R,  \forall k \in C \label{main_cond_3}\\
+&  \sum_{i \in S \cup R / Z}\sum_{j \in Z } x^{(k)}_{i,j} > 0, 
+Z=\{z \in Z: \sum_{j \in S \cup R}x^{(k)}_{j,z}>0 \}  ,\forall k \in C \label{main_cond_4}
 \end{align}
 
-\subsection{Анализ исходных данных}
-В случае, когда простой автомобиля перед посещением следующего клиента не влияет на целевую функцию, а именно $\varphi=const=0$, тогда в некоторых случаях временные ограничения могут не влиять на структуру графа и задача сводится к задаче о назначениях.
-В силу того, что $NP \neq P$, однозначный переход невозможен, но из-за наличия временных ограничений граф задачи может не иметь локальных циклов и в этом случае  ограничения (\ref{main_cond_4}) и (\ref{main_cond_5}) возможно опустить после удаления дуг $(i,j)$ по которым невозможно достижение вершины $j$ в не зависимости от времени прибытия в вершину $i$.
+The restriction (\ref{main_cond_1}) prohibits to a node in the graph $G$ has more than one output arc.
+The restriction (\ref{main_cond_2}) prohibits to a node has more than one input arc.
+The constraint (\ref{main_cond_3}) indicates that the number of input arcs to the node can not be less than the output arcs (this constraint considers the fact that the vehicle can leave the destination only if it has visited this node).
+The restriction (\ref{main_cond_4}) excludes local loops.
 
-Для того, чтобы проверить возможность такого перехода, введем дополнительные величины. 
-\begin{enumerate}
-\item Учитывая наложенные на целевую функцию ограничения, время прибытия в $j$–й пункт назначения равно:
-\begin{equation} \label{t_j_input}
-   \exists h_i \in [0,1] : \tilde{t_i} = t_i + h_i \cdot \Delta t_i
-\end{equation}
-
-\item Введем матрицу $U$, элементы которой показывают в количественной форме нарушение временного окна:
-\begin{equation} \label{matrix_U}
- U(\vec{h}) : u_{i,j} = 
-    \begin{cases}
-	  0, i=j, \\
-      t_i + h_i \cdot \Delta t_i + \omega_i + t_{i,j} - t_j - \Delta t_j, i \neq j
-    \end{cases}   
-\end{equation}
-\end{enumerate}
-
-Можно сказать, что если элемент матрицы $U(\vec{h}=\vec{1})$ отрицательный, то передвижение между пунктами возможно выполнить в не зависимости от временных ограничений.
-Поэтому построим граф $G$ для задачи маршрутизации, в котором уже будут учтены ограничения (\ref{main_cond_4}) и (\ref{main_cond_5}), задав следующую матрицу инцидентности:
-\begin{equation} \label{matrix_I}
- I(\vec{h}) : I_{i,j} = 
-    \begin{cases}
-	  0, &\text{если $u_{i,j} \geq 0$,} \\
-      1, &\text{если $u_{i,j}<0$}
-    \end{cases}, i \in C \cup Q, j \in Q
-\end{equation}
-\newtheorem{Def}{Определение} 
-\begin{Def}\label{def_main}
-Если:
-  \begin{enumerate}
-  \item $\varphi=const=0$;
-  \item матрица $I(\vec{0})$ равна матрице $I(\vec{1})$;
-  \item граф $G$, построенный по матрице инцидентности $I(\vec{0})$ не является мультиграфом,
-  \end{enumerate}
- тогда задача маршрутизации с жесткими ограничениями по времени сводится к задаче о назначениях.
-\end{Def}
-Это утверждение вытекает из следующих лемм.
-
-\newtheorem{Lem}{Лемма}
-\begin{Lem}
-Если матрица $I(\vec{0}) = I(\vec{1})$, тогда $I(\vec{0})=I(\vec{h}), \forall \vec{h}: h_i \in [0,1]$
-\end{Lem}
-Доказательство. 
-Элементы матрицы $I_{i,j}$ не зависят от вектора $\vec{h}$ при $i \in C, j \in Q$, так как вершины из множества C соответствуют местоположениям автомобилей в начальный момент времени и им не заданы временные окна ($\Delta t_i = 0, \forall i \in C$).
-Остальные элементы матрицы можно представить в виде формулы:
-\begin{equation} 
-I_{i,j}(h_i) = \frac{1 - sign(u_{i,j}(h_i))}{2}
-\end{equation}
-По условию $I(\vec{0}) = I(\vec{1})$, тогда:
-
-\begin{flalign*}
-&\frac{1 - sign(u_{i,j}(0))}{2} =  \frac{1 - sign(u_{i,j}(1))}{2} \Rightarrow
-sign(u_{i,j}(0)) = sign(u_{i,j}(1))&
-\end{flalign*}
-Или:
-\begin{flalign*}
-&\left[\begin{array}{l}
-     \begin{cases}
-	  u_{i,j}(0)>0  \\
-      u_{i,j}(1)>0
-    \end{cases}\\
-    \begin{cases}
-	  u_{i,j}(0)<0  \\
-      u_{i,j}(1)<0
-    \end{cases}
-\end{array}\right.&
-\end{flalign*}
-
-Функция $u_{i,j}(h_i)$ на промежутке $h_i \in [0,1]$ является монотонно возрастающей, потому что производная $\dd{u_{i,j}(h_i)}{h_i} = \Delta t_i >0$ положительна на всем промежутке. Поэтому:
-
-\begin{flalign*}
-&\left[\begin{array}{l}
-     \begin{cases}
-	  u_{i,j}(0)>0  \\
-      u_{i,j}(1)>0
-    \end{cases} \Rightarrow u_{i,j}(h_i)>0 \\
-    \begin{cases}
-	  u_{i,j}(0)<0  \\
-      u_{i,j}(1)<0
-    \end{cases} \Rightarrow u_{i,j}(h_i)<0
-\end{array}\right.
-\Rightarrow sign(u_{i,j}(0)) = sign(u_{i,j}(h_i)), \forall h_i \in [0,1] &
-\end{flalign*}
-Тогда $sign(u_{i,j}(0)) = sign(u_{i,j}(h_i)) \Rightarrow  I(\vec{h})=I(\vec{0})$. Доказано.
-
-\newtheorem{Lem2}{Лемма}
-\begin{Lem}
-Если граф  $G$, построенный по матрице инцидентности $I(\vec{h}=const)$ не является мульграфом, то такой граф не имеет циклов.
-\end{Lem}
-Доказательство(от противного). Пусть дан граф $G(V={C \cup Q}, E)$, построенный по матрице инцидентности  $I(\vec{h}=const)$ и в нем имеется цикл: \begin{align*}
-&E_{cicle}=\{(v_{s_1},v_{s_2}),(v_{s_2},v_{s_3}),...,(v_{s_{k-1}},v_{s_k}),(v_{s_k},v_{s_1})\}
-\end{align*}
-Предположим, что при наличии цикла $E_{cicle}$ в графе $G$ нет дуги $(v_{s_1},v_{s_k})$, т.е.
-АТС не успеет добраться из вершины $v_{s_1}$ в вершину $v_{s_k}$ не нарушив временные ограничения.
-Поэтому:
-\begin{equation} \label{error_inequality}
- t_{s_k} + \Delta t_{s_k} <  \tilde{t}_{s_1} + \omega_{s_1} + t_{s_1,s_k}
-\end{equation}
-Согласно предположению о существовании цикла $E_{cicle}$ верно следующее неравенство ( $E_{route}=E_{cicle} / \{(v_{s_k},v_{s_1}) \}$):
-\begin{equation} \label{time_s_k_inequality}
- t_{s_k} + \Delta t_{s_k} > \tilde{t}_{s_{k-1}} + \omega_{s_{k-1}} + t_{s_{k-1},s_k} = \tilde{t}_{s_1} +  \sum_{(i,j) \in E_{route}} (t_{i,j} +  \omega_{s_i}) = A
-\end{equation}
-Пусть передвижение от вершины $v_{s_1}$ в вершину $v_{s_k}$ будет выполняться по дугам $E_{route}$. Тогда:
-\begin{equation} \label{time_route}
- \tilde{t}_{s_1} + \omega_{s_1} + t_{s_1,s_k} \leq \tilde{t}_{s_1} + \omega_{s_1} + \sum_{(i,j) \in E_{route}} t_{i,j} \leq A
-\end{equation}
-Подставим (\ref{time_route}) в (\ref{time_s_k_inequality}):
-\begin{equation} \label{true_inequality}
- t_{s_k} + \Delta t_{s_k} > \tilde{t}_{s_1} + \omega_{s_1} + t_{s_1,s_k}
-\end{equation}
-Пришли к противоречию, следовательно, неравенство (\ref{error_inequality}) неверно. Доказано.
-
-\subsection{Переход к задаче о назначениях}
-Если выполняются правила из определения \ref{def_main}, тогда можно свести исходную задачу к задаче о назначениях.
-
-Введем замену переменных X :
-\begin{equation} \label{y_replace}
-\forall k \in C: ~Y^k = \{ I_{i,j}(0) \cdot x^{(k)}_{i,j} , i \in Q \cup  \{k\} , j \in Q  \}
-\end{equation}
-
-Теперь задачу можно сформулировать таким образом, что ее можно разрешить за полиномиальное время:
+Then next constraints synchronize values of variables $X$ and $\vec{y}$ for each request $q \in Q$ and prohibits the service of receiver before a visit to the sender:
 \begin{align} 
-  &F(Y) = 
-    \sum_{k \in C}
-     \sum_{i \in Q \cup  \{k\},j \in Q} 
-     \Omega_{i,j}\cdot y^{(k)}_{i,j})
-     \to min\\
-& \sum_{k \in C}\sum_{j \in Q}y^{(k)}_{i,j} = 1, 
-  \forall i \in C \cup Q, \forall k \in C \label{mod_cond_1}\\ 
-& \sum_{j \in Q} ( 
-       y^{(k)}_{k,j} -y^{(k)}_{i,j} ) \geq 0, 
-       \forall i \in Q ,  \forall k \in C \label{mod_cond_2}\\
-& \sum_{j \in Q \cup  \{k\}} y^{(k)}_{i,\omega} - 
-  \sum_{i \in Q} y^{(k)}_{\omega,j} = 0, 
-  \forall \omega \in Q ,  \forall k \in C \label{mod_cond_3}
+& x^{(k)}_{s_q} + x^{(k)}_{r_q} = 2 \cdot y^{(k)}_{q}(X), \forall k \in C, q \in Q  \\
+& y^{(k)}_{q}(X) \cdot (\tilde{t}^k_{r_q}(X^{(k)})-\tilde{t}^k_{s_q}(X^{(k)})\ge{0}, \forall k \in C, q \in Q
 \end{align}
 
-Чтобы получить задачу о назначениях в классическом ее виде, построим на основе графа $G$ двудольный граф $G_b(V_{begin},V_{end},E_b)$ по следующему принципу:
-\begin{enumerate}
-\item $V_{begin} = \{v \in C \cup Q\ : \exists (v,\omega) \in E, \omega \in Q \}$ - состоит из множества вершин, для которых найдется хотя бы одна выходящая дуга из множества E;
-\item $V_{end} = \{v  \in Q\ : \exists (\omega,v) \in E, \omega \in C \cup Q \}$ - состоит из множества вершин, для которых найдется хотя бы одна входящая дуга из множества E;
-\item $E_b = \{(v,\omega) \in E: v \in V_{begin}, \omega \in V_{end} \}$.
-\end{enumerate}
+Define restrictions for accounting of vehicle capacity and time windows:
+\begin{align} 
+& \sum_{j\in{Q}} \omega_j \cdot y_{j}^{k} \leq L_k, \forall{k}\in{C}\\
+& t_{s}^{q} \leq \tilde{t}^k_{s_q}(X^{(k)} \leq t_{s}^{q}+ \Delta{t_{s}^{q}}, \forall q \in Q, \label{tws_cond} \\
+& t_{r}^{q} \leq \tilde{t}^k_{r_q}(X^{(k)} \leq t_{r}^{q}+ \Delta{t_{r}^{q}}, \forall q \in Q, \label{twr_cond}
+\end{align}
 
-\begin{figure}[h]
-\hfil\includegraphics[height=2.0in]{sign_graph}\hfil
-\caption
+\subsection{Description of the Algorithm}
+
+\begin{algorithm}[H]
+	\textbf{function TabuSearch}($x^0,l,p,N_{stop}$) \\
+	// Initialize variables:	\\
+	$x^{opt} = x^0$;
+	$F^{opt} = F^0$;
+	$\phi^{0} =  \emptyset$;
+	$i=0;s=0$; \\
+	\While{the breakpoint is not triggered }{
+	   $N_l = N_l(x^i,\phi^i,p) $; \\
+		\eIf{$N_l \ne   \emptyset $}{
+			$ x^{i+1} = x^i; i = i + 1;$ \\
+			\textbf{continue;}
+		}
 		{
-  Пример преобразования графа $G$ в двудольный граф $G_b$.
+			// find optimum into the neighbord $N_l$: \\
+			$ x^{i+1} : F(x^{i+1}) = min\{ F(y): y \in N_l \}$;
 		}
-\label{aba:fig1}
-\end{figure}
 
-\subsection{Результаты}
-Преобразованную задачу(рис.
-\ref{aba:fig1}) рекомендуется решать с помощью метода Голдберга и Кеннеди,  основанного на технике масштабирования при переходе к задаче поиска потока минимальной стоимости \cite{Goldberg_Kennedy}.
-Сложность алгоритма равна $O(\sqrt{n}mlog(nC))$.
+		\If{$F(x^{i+1}) < F_{opt})$}{
+			$ F_{opt} = F(x^{i+1})$;
+			$ x^{opt} = x^{i+1}$;
+		}
 
-В случае, если функция $\varphi(t)$ присутствует в целевой функции, то решение преобразованной задачи может послужить нижней границей, которую можно использовать для отсеивания заведомо худших решений при формировании окрестности вокруг текущего решения в работе вероятностного алгоритма поиска с запретами.
+		$\phi^{i+1} = update(\phi^i)$;\\
+		i = i + 1;\\
 
-Проверка метода была выполнена на тестовых задачах Кристофидеса, Голдберга  и Тейларда с временными ограничениями, удовлетворяющими условиям, которые были изложены в определении \ref{def_main} \cite{problems_Christofides,problems_Golden,problems_Taillard}.
+	}
+	\textbf{return} $x^{opt}$;
 
-\subsection{Выводы}
-Предложены правила перехода от задачи маршрутизации с учетом временных ограничений к задаче о назначениях.
-Данную технику целесообразно использовать для задач маршрутизации, в которых не учитывается простой автомобиля.
-Также  метод можно использовать для инициализации метаэвристического алгоритма или в виде более качественной нижней границы для определения оптимума.
+\caption{Pseudo-code for probabilistic tabu-search algorithm.}
+\label{alg:generalGP}
+\end{algorithm}
+
+\subsection{Initialization}
+
+\subsection{A Method of Forming Solutions Neighborhood}
+
+\subsection{Modification}
+The first generalized diagram is as follows.
+First, the set of solutions is constructed by Solomon.
+Then every solution is improved by tabu search and choose the best solution according to the objective function.
+Modification of the scheme is based on the hypothesis of "On a large valley." According to this hypothesis, the average local optima are located much closer to the global than a randomly chosen point.
+There is a certain concentration of local optima in a small part of the feasible region, which is figuratively called a large valley.
+If this assumption is true, then it is advisable to remember the best solutions and based on them design new original decision.
+We use this idea to solve the  Routing Courier Delivery Problem.
+
+From these solutions generated population.
+To do this, first sort by the value of the objective function in ascending order.
+Each solution gets into the population with a given probability, the probability of selection decreases with increasing atomic number.
+It then crosses the decisions by the following rule: the first solution is randomly selected route, then the other solutions chosen a route that does not include the customer first route, etc.
+Other clients added by Solomon in the selected route.
+Then the process is repeated.
+
+\subsection{Conclusion}
+Text...
 
@@
 
-\subsection{Автор}
-\author{Роман Александрович Шафеев}{магистр 2-го года обучения, факультет информатики и управления, Национальный технический университет <<Харьковский политехнический институт>>, Харьков, Украина}
-{rs@premiumgis.com}
-
 \end{document}

### 2015-02-03 article.tex
add modify section

@@
 \begin{document}
-each code-block optionally)
 
@@
 %%%%%%%%%%%%%%% You have to specify name of your paper here
-``Investigation of tuning parameters and modification of tabu search algorithm for solving the Static Routing Courier Delivery Problem.'' 
+``Investigation of tuning parameters of tabu search algorithm and its modification  for solving the static Routing Courier Delivery Problem.'' 
 }
@@
 
-In the paper ....
+The paper considers the routing courier delivery problem with the service time for which the discrete model was constructed and the calculation scheme based on metaheuristic algorithm.
+The efficiency  of the proposed algorithm‘s scheme has been tested on a large-scale problems, generated on the basis of different classes known model routing problems.
+
 }
@@
 %%%%% Paragraph text. To separate paragraphs use empty lines.
- The VRP \cite{VRPTW} ...
+In this study, we consider a static vehicle routing problem with time windows. This variant of the routing problem is a common problem in the courier delivery industry.
+
+The resulting solution is a master plan that solves the problem, and takes into account the daily events that take place according to plan.
+We evaluated the model and strategy decisions through simulation of randomly generated test data and known problems.
 
@@
 
+\begin{figure}[h]
+	\hfil\includegraphics[height=4.0in]{images/scheme}\hfil
+	\caption
+	{
+		Example of Routing Delivery Problem.
+	}
+	\label{aba:fig1}
+\end{figure}
+
 \subsection{Discrete model}
@@
 
-Let $\{X^k\}^n_{k=1}$ is a sequence of variable matrices for each vehicle $k \in C$. Elements of the matrices take the following values:
+Let $X = \{X^k\}^n_{k=1}$ is a sequence of variable matrices for each vehicle $k \in C$. Elements of the matrices take the following values:
 \begin{equation}
@@
 \subsection{Description of the Algorithm}
+Present the set of matrices $\{X^k\}$ in the form of a vector defined on the hypercube $E^\eta=\{0.1\}^\eta$, where $\eta=n\cdot(2m+1)\cdot 2m$.
+The aim of the task is to minimize the objective function:
+\begin{align} 
+& F(\vec{u})\to min,\vec{u}\in E^{\eta}
+\end{align}
+
+Denote by $\delta(\vec{u},\vec{v})$ the Hamming distance between $\vec{u}$ and $\vec{v}$. Through $N_l(\vec{u})$ we denote the neighborhood of a point $\vec{u}$ radius $l$, ie:
+\begin{align} 
+& N_l(\vec{u})=\{\vec{v} \in E^{\eta}:\delta(\vec{u},\vec{v})\le l \}, l=\bar{1,\eta}
+\end{align}
+
+When $l=\eta$ a set $ N_l(\vec{u})$ for any vector $\vec{u}$ coincides with the set $E^{\eta}$ and being in this neighborhood vector with a minimum value of the objective function is equivalent to solving the original problem.
+The standard algorithm local search algorithm starts with a randomly selected vector $\vec{u^0}$.
+On the $i$ step of the algorithm move from the current vector in the neighbouring that the minimum value of the objective function in the neighborhood of the given vector:
+\begin{align} 
+& F(\vec{u}^{i+1})=min\{F(\vec{v}):\vec{v} \in N_l(\vec{u}^i)\}
+\end{align}
+
+The algorithm terminates at a local optimum, when $F(\vec{x^{i+1}})=F(\vec{x^i})$.
+When solving the transportation problem is a typical situation when many local optimums and only one of them is global:
+\begin{align} 
+& F_{opt}=min\{F(\vec{v}):v \in E^{\eta}\}
+\end{align}
+
+To ensure that the algorithm did not stop in a local minimum, and passed from one local minimum to another, with edges removed the Central point and when searching for the minimum of the following rule applies.
+Let $l=2$ and in the transition from $\vec{u^{i+1}}$ to $\vec{u^i}$ change the values in the coordinate $(u_\lambda^{i}, u_\omega^{i})$.
+The algorithm stores such pair for the last $h$ couple of steps and in the next step prohibits the movement in these directions.
+An ordered list of such pairs
+\begin{align} 
+& \phi^i=\{(u_\lambda^{i},u_\omega^{i}),(u_\lambda^{i-1},u_\omega^{i-1}),\cdots,(u_\lambda^{i-h+1},u_\omega^{i-h+1}) \} 
+\end{align}	
+called the list of prohibitions.
+When building the list of all pairs of different and a pair of $(u_\lambda,u_\omega),\lambda \ne \omega$ does not prohibit the movement of pairs $(u_\lambda,u_\lambda)$ and $(u_\omega,u_\omega)$.
+When $l>2$ similarly constructed three coordinates, fours, etc.
+A set of non-restricted vectors denote by $N_l(\vec{u^i},\vec{\phi^i})$.
+In order that the search was effectively, it is advisable to use small values $h$ and to control this parameter in the course of the algorithm.
+
+Denote by $N_l(\vec{u^i},\vec{\phi^i},p)$ probabilistic edge that goes from a deterministic $N_l(\vec{u^i},\vec{\phi^i})$ as follows.
+Each vector $\vec{v} \in N_l(\vec{u^i},\vec{\phi^i})$ with probability $p$ included in the neighborhood $N_l(\vec{u^i},\vec{\phi^i},p)$ regardless of other points.
+Note that this set may be empty or contain only one point.
+The General scheme of the probabilistic search algorithm with a list of prohibitions referred to as pseudo-code.
+
+As the stopping criterion is based on the total number of steps $N_stop$, in the course of which does not change the value $F_{opt}$.
+Values $l,p,h$ are the control parameters of the algorithm.
+Their choice depends on the problem dimension.
 
 \begin{algorithm}[H]
-	\textbf{function TabuSearch}($x^0,l,p,N_{stop}$) \\
+	\textbf{function TabuSearch}($u^0,l,p,h$) \\
 	// Initialize variables:	\\
-	$x^{opt} = x^0$;
+	$u^{opt} = u^0$;
 	$F^{opt} = F^0$;
 	$\phi^{0} =  \emptyset$;
-	$i=0;s=0$; \\
+	$i=0$; \\
 	\While{the breakpoint is not triggered }{
-	   $N_l = N_l(x^i,\phi^i,p) $; \\
+	   $N_l = N_l(u^i,\phi^i,p,h) $; \\
 		\eIf{$N_l \ne   \emptyset $}{
-			$ x^{i+1} = x^i; i = i + 1;$ \\
+			$ u^{i+1} = u^i; i = i + 1;$ \\
 			\textbf{continue;}
@@
 			// find optimum into the neighbord $N_l$: \\
-			$ x^{i+1} : F(x^{i+1}) = min\{ F(y): y \in N_l \}$;
+			$ u^{i+1} : F(u^{i+1}) = min\{ F(y): y \in N_l \}$;
 		}
 
-		\If{$F(x^{i+1}) < F_{opt})$}{
-			$ F_{opt} = F(x^{i+1})$;
-			$ x^{opt} = x^{i+1}$;
+		\If{$F(u^{i+1}) < F_{opt})$}{
+			$ F_{opt} = F(u^{i+1})$;
+			$ u^{opt} = u^{i+1}$;
 		}
@@
 	}
-	\textbf{return} $x^{opt}$;
+	\textbf{return} $u^{opt}$;
 
 \caption{Pseudo-code for probabilistic tabu-search algorithm.}
-\label{alg:generalGP}
+\label{alg:TabuSearch}
 \end{algorithm}
 
+In the presented scheme, it was assumed that the value of $h$ (the dimension of the deny list) does not change in the course of the algorithm.
+This creates certain difficulties in the implementation of the scheme, as it is unknown how long is to take a list of prohibitions.
+At small $h$ the algorithm can start walking cycle.
+At large $h$ the search becomes inefficient.
+One of the simple rules regulating the length of the list of prohibitions is the following.
+
+If at the next step of the algorithm, the edge  $N_l(\vec{u^i},\vec{\phi^i},p)$ turned out to be empty, it is added to the list of fictitious vector $\vec{0}$.
+This vector nothing forbids, but reduces the number of bans per unit.
+
 \subsection{Initialization}
+To use a search algorithm is restricted by the need to build the initial solution.
+To obtain the initial solution $\vec{u^0}$ you can use a heuristic method to construct the route.
+The essence of these methods is the following.
+Every car has consistently built a route by adding another is not considered clients according to specified rules.
+
+The most known method of construction of the route proposed by Solomon in 1987 year and found in the literature as a method of Solomon.
+This algorithm is fast enough (computational complexity $O(n^3)$, therefore, it is useful to initialize a vector $\vec{u^0}$.
+
+At the beginning of the algorithm, you add one to a customer on the route according to time limits.
+The choice of the first client is randomly or chosen one, you want to serve first.
+Then, each valid client $u$ from the set of outstanding clients $Q^{'}$ (the set $Q^{'}\subset{Q}$) evaluated insertion between two adjacent nodes route.
+For selection of inserts used criterion:
+\begin{align} 
+& c(v_i,u,v_i+1)=[\min_{p=2,\dots,n^k}(\Omega_{p-1,u}^k+\Omega_{u,p}^k-\mu \Omega_{p-1,p}^k) ]
+\end{align}	
+where:
+
+$k$ - vehicle route $k\in C$;
+$n^k$ - the number of nodes in the route;
+$\mu$ - arbitrary parameter, $\mu \ge 0 $.
+
+Added the client, which is $c(v_i,u,v_i+1)$ least:
+\begin{align} 
+& u^*:c(v_i,u^*,v_i+1)=min[c(v_i,u,V_i+1)]
+\end{align}
+
+As a result the client $u^*$ is added to the current route $k$ the position between nodes $v_i$ and $v_i+1$.
 
 \subsection{A Method of Forming Solutions Neighborhood}
+The most difficult part of the local search algorithm with the prohibitions are viewing the surroundings $N_l(\vec{u^i},\vec{\phi^i})$ of the current point $\vec{us^i}$.
+For even larger values of $l$ and the parameter $p=1$ in the outskirts plenty of points.
+But due to the nature of the model's constraints many of them are not the solution.
+Therefore, due to the nature of the considered discrete tasks, we used the following method of determining the area.
+On the $i$ step of the algorithm compiled from routes pair: $\{R_1,R_2\},\{R_1,R_3\},\dots,\{R_1,R_n\},\{R_2,R_1\}, \{R_2,R_3\},\dots,\{R_2,R_n\},\dots,\{R_{n-1},R_{n}\}$.
+Then over each pair of routes is performed in moving requests.
+The resulting displacement vectors $\vec{u}$ are neighborhood $N_l(\vec{u^i},\vec{\phi^i})$.
+For moving  we used change-move  and swap-move operations.
 
-\subsection{Modification}
-The first generalized diagram is as follows.
-First, the set of solutions is constructed by Solomon.
-Then every solution is improved by tabu search and choose the best solution according to the objective function.
-Modification of the scheme is based on the hypothesis of "On a large valley." According to this hypothesis, the average local optima are located much closer to the global than a randomly chosen point.
-There is a certain concentration of local optima in a small part of the feasible region, which is figuratively called a large valley.
-If this assumption is true, then it is advisable to remember the best solutions and based on them design new original decision.
-We use this idea to solve the  Routing Courier Delivery Problem.
+\subsubsection{The optimization of the algorithm parameters}
 
@@
 
+\subsection{Analysis of the algorithm parameters}
+Modified tabu search algorithm has 3 main parameters: $p$ -- the probability of hitting solutions in a neighborhood, $h$ -- size of the list of prohibitions and $N_{neighbors}$ -- accepted count limit of size of neighbordhood $N_l(\vec{u})$.
+For the experiments, we constructed the test tasks on the basis of the capacity routing problems developed by Breedam, Fisher, Christofides and Eilon.
+To determine trends in time and cost of the decision to use exponential smoothing by a factor of $0.1$.
+
+In the first computational experiment we made finding the best solutions for different values of the tabu size parameter $h \in [0, 40]$ (pict.
+1 an pict.
+2).
+In this experiment we use two different breakpoint strategies:
+\begin{enumerate}
+	\item $StepCount$ termination strategy --  terminates when an amount of steps has been reached;
+	\item $TimeMillisSpent$ termination strategy -- terminates when an amount of time has been used.
+\end{enumerate}
+
+\begin{figure}[h]
+	\hfil\includegraphics[height=3.0in]{images/tabuSize_stepCount}\hfil
+	\caption
+	{
+	  The dependence of the quality of solutions on the size of the tabu list when using $ StepCount $ termination strategy
+	}
+	\label{aba:fig1}
+\end{figure}
+
+As should have been expected, for large values of the parameter a jam occurs in local optimum due to the large number of restrictions of movement in space.
+Otherwise if you select a too small tabu size, algorithm can still get stuck in a local optimum.
+\begin{figure}[h]
+	\hfil\includegraphics[height=2.0in]{images/tabuSize_time}\hfil
+	\caption
+	{
+	  The dependence of the quality of solutions on the size of the tabu list when using $ TimeMillisSpent $ termination strategy
+	}
+	\label{aba:fig1}
+\end{figure}
+
+In the second experiment we built a relationship between the number $N_{neighbors}$ of viewed solutions neighborhood $N_l(\vec{u})$ using $TimeMillisSpent$ termination strategy. 
+\begin{figure}[h]
+	\hfil\includegraphics[height=2.0in]{images/acceptedCountLimit}\hfil
+	\caption
+	{
+	  The dependence of the quality of solutions on the neighborhood size when using $ TimeMillisSpent $ termination strategy
+	}
+	\label{aba:fig1}
+\end{figure}
+
+\subsection{Modification}
+
+The first generalized diagram is as follows.
+First, the set of solutions is constructed by Solomon.
+Then every solution is improved by tabu search and choose the best solution according to the objective function.
+
+Modification of the scheme is based on the hypothesis of "On a large valley." According to this hypothesis, the average local optima are located much closer to the global than a randomly chosen point.
+There is a certain concentration of local optimum in a small part of the feasible region, which is figuratively called a large valley.
+If this assumption is true, then it is advisable to remember the best solutions and based on them design new original decision.
+We use this idea to solve the  Routing Courier Delivery Problem.
+
+\begin{algorithm}[H]
+
+	\textbf{function constructSolution}( $U^{opt}, p_u, n^{min}_s$) \\
+	// $U^{opt}$ -- an ordered set of optimal solutions \\
+
+	\If{$ sizeof(U^{opt}) <  n^{min}_s$}{
+		// construct solution using heuristics(for example, Solomon alg.)\\
+		$u^0 = heuristicsConscruction();$
+		return $u^0$;
+	}
+
+	\While{$i < sizeof(U^{opt})$}{
+		\If{$random(0.0,1.0) > p^i_u$}{
+		   i = i + 1; \\
+		   goto 7;
+		}
+		$R = $randomly select a route from $u^{opt}_i$ solution;
+		$u^0 = u^0 \cup R$;
+		i = i + 1;\\
+
+	}
+	\textbf{return} $u^0$;
+
+	\caption{Pseudo-code for heuristics construction algorithm.}
+	\label{alg:constructSolution}
+\end{algorithm}
+
+Let $\vec{\psi_i} = (\psi^1_i,\psi^2_i,\psi^3_i) = (\tilde{l_i},\tilde{p_i},\tilde{l_h})$ -- a set of values of TabuSearch configuration parameters which was using for solving the problem in the step $i$.
+
+Let $[\psi^j_{begin}, \psi^j_{end}]$ -- an optimal interval of the TabuSearch $j$ -parameter.
+(For example: if $j=3$, then: $[\psi^31_{begin}, \psi^3_{end}]= [h_{begin}, h_{end}] = [10, 40] $).
+These intervals are initial data and they are set in the initialization block.
+
+\begin{algorithm}[H]
+	\textbf{function SolveCDP}( $n^{min}_s$) \\
+	// Initialize variables:	\\
+
+	$\psi = \emptyset$;
+	$p_u = \emptyset$;
+	$U^{opt} = \emptyset$;
+	$i=0$; \\
+
+	\While{the breakpoint is not triggered }{
+		// set TabuSearch parameters \\
+		\For{$j=1..3$}{
+	  	// $f_j$ -- the density distribution of the random \\
+	  	// variable $\psi^j$ (TabuSearch parameter $l,p $ or $ h$) \\
+	  	$\psi^j_i = random(\psi^j_{begin}, \psi^j_{end}, f_j); $
+	  	}
+
+		// Make solution using euristics construction algorithms \\
+		$u^0 = constructSolution(U^{opt} ,n^{min}_s)$
+
+		// Improve solution $u^0$ using TabuSearch algorithm 
+		$u^{opt}_i = TabuSearch(u^0, \vec{\psi_i});$ \\
+
+		$U^{opt}  = U^{opt} \cup \{u^{opt}_i\};$ \\
+
+		// Sort ascending optimums $U$ of the objective function \\
+		$U^{opt} = sort(U^{opt});$\\
+		$p^{i+1}_u = update(p^i_u)$;\\
+
+		i = i + 1;\\
+
+	}
+	\textbf{return} $u^{opt}_1$;
+
+	\caption{Pseudo-code for modified tabu-search algorithm.}
+	\label{alg:modTabuSearch}
+\end{algorithm}
+
 \subsection{Conclusion}
-Text...
+In the process, has been studied and implemented tabu search algorithm to solve the static problem with time constraints.
+Modification of the algorithm to optimize the computational scheme.
+Modified algorithm gave the best solutions in terms of the balance between the number of vehicles and the distance traveled.
+such a feature of this algorithm can be used for decision support in intelligent systems.
+In practice, the more advantageous to use more vehicles if this will reduce the time to use them, this will reduce fuel costs and depreciation of transport;
+
+The analysis of the parameters of implemented algorithms allowed us to determine their optimal values for this class of routing problems.
+With the modified algorithm was found solutions of model problems, which in most cases have an acceptable deviation from the global optimum;
 

### 2015-02-04 article.tex
write alpha

@@
 
-The paper considers the routing courier delivery problem with the service time for which the discrete model was constructed and the calculation scheme based on metaheuristic algorithm.
-The efficiency  of the proposed algorithm‘s scheme has been tested on a large-scale problems, generated on the basis of different classes known model routing problems.
+The paper considers the routing courier delivery problem with the service time for which the discrete model was constructed and the calculation scheme based on tabu search algorithm.
+The efficiency  of the proposed algorithm‘s scheme has been tested on a large-scale problems, which were generated on the basis of different classes known model routing problems.
 
@@
 %%%%% Paragraph text. To separate paragraphs use empty lines.
-In this study, we consider a static vehicle routing problem with time windows. This variant of the routing problem is a common problem in the courier delivery industry.
+In the research we consider a static vehicle routing problem with time windows.
+This variant of the routing problem is a common problem in the courier delivery industry.
+Effective solution to this problem affects the quality of services provided by the company, as well as the magnitude of potential losses caused by the excess of the agreed terms of delivery.
+The proposed solution to the problem is based on the automation of the process of planning the route and involves the use of distributed information systems for the possible implementation of these approaches in practice.
+Another important criterion for good transport plan are low transport costs, which affect the cost of services.
 
-The resulting solution is a master plan that solves the problem, and takes into account the daily events that take place according to plan.
-We evaluated the model and strategy decisions through simulation of randomly generated test data and known problems.
+The Routing Courier Delivery Problem with time windows requires that a set of vehicles satisfy a collection of customer requests.
+Each client request requires the use of a single vehicle to receive and load a specified amount of goods at one source location(sender) and to deliver them to destination location(receiver).
+All requests must be performed without violating either the vehicle capacity or the customer time window.
+The solution is a master plan that solves the problem, and takes into account the daily events that take place according to plan.
+In this paper, we present a discrete model and a modified tabu search algorithm  to solve the  Courier Delivery Problem with time windows.
+
+In order to validate our strategies decisions technique effectiveness, we have constructed a new set of benchmark problems for the pickup and delivery problem with time windows based on which were generated on the basis of different classes known model routing problems.
+Computational results substantiate the solution quality and efficiency of our modified tabu search algorithm.
 
@@
 	{
-		Example of Routing Delivery Problem.
+		Example of the Routing Delivery Problem.
 	}
@@
 
+\begin{algorithm}[H]
+	\textbf{function TabuSearch}($u^0,l,p,h$) \\
+	// Initialize variables:	\\
+	$u^{opt} = u^0$
+	$F^{opt} = F^0$
+	$\phi^{0} =  \emptyset$
+	$i=0$ \\
+	\While{the breakpoint is not triggered }{
+	   $N_l = N_l(u^i,\phi^i,p,h) $ \\
+		\eIf{$N_l \ne   \emptyset $}{
+			$ u^{i+1} = u^i$ \\
+			$i = i + 1$ \\
+			\textbf{goto} 5
+		}
+		{
+			// find optimum into the neighbord $N_l$: \\
+			$ u^{i+1} : F(u^{i+1}) = min\{ F(y): y \in N_l \}$
+		}
+
+		\If{$F(u^{i+1}) < F_{opt})$}{
+			$ F_{opt} = F(u^{i+1})$ \\
+			$ u^{opt} = u^{i+1}$
+		}
+
+		$\phi^{i+1} = update(\phi^i)$\\
+		i = i + 1\\
+
+	}
+	\textbf{return} $u^{opt}$
+
+\caption{Pseudo-code for probabilistic tabu-search algorithm.}
+\label{alg:TabuSearch}
+\end{algorithm}
+
 Denote by $N_l(\vec{u^i},\vec{\phi^i},p)$ probabilistic edge that goes from a deterministic $N_l(\vec{u^i},\vec{\phi^i})$ as follows.
@@
 Note that this set may be empty or contain only one point.
-The General scheme of the probabilistic search algorithm with a list of prohibitions referred to as pseudo-code.
+The General scheme of the probabilistic search algorithm with a list of prohibitions referred to as pseudo-code in algorithm \\
+scheme \ref{alg:TabuSearch}.
 
@@
 Their choice depends on the problem dimension.
-
-\begin{algorithm}[H]
-	\textbf{function TabuSearch}($u^0,l,p,h$) \\
-	// Initialize variables:	\\
-	$u^{opt} = u^0$;
-	$F^{opt} = F^0$;
-	$\phi^{0} =  \emptyset$;
-	$i=0$; \\
-	\While{the breakpoint is not triggered }{
-	   $N_l = N_l(u^i,\phi^i,p,h) $; \\
-		\eIf{$N_l \ne   \emptyset $}{
-			$ u^{i+1} = u^i; i = i + 1;$ \\
-			\textbf{continue;}
-		}
-		{
-			// find optimum into the neighbord $N_l$: \\
-			$ u^{i+1} : F(u^{i+1}) = min\{ F(y): y \in N_l \}$;
-		}
-
-		\If{$F(u^{i+1}) < F_{opt})$}{
-			$ F_{opt} = F(u^{i+1})$;
-			$ u^{opt} = u^{i+1}$;
-		}
-
-		$\phi^{i+1} = update(\phi^i)$;\\
-		i = i + 1;\\
-
-	}
-	\textbf{return} $u^{opt}$;
-
-\caption{Pseudo-code for probabilistic tabu-search algorithm.}
-\label{alg:TabuSearch}
-\end{algorithm}
 
@@
 
-\subsubsection{The optimization of the algorithm parameters}
-
-From these solutions generated population.
-To do this, first sort by the value of the objective function in ascending order.
-Each solution gets into the population with a given probability, the probability of selection decreases with increasing atomic number.
-It then crosses the decisions by the following rule: the first solution is randomly selected route, then the other solutions chosen a route that does not include the customer first route, etc.
-Other clients added by Solomon in the selected route.
-Then the process is repeated.
-
 \subsection{Analysis of the algorithm parameters}
@@
 To determine trends in time and cost of the decision to use exponential smoothing by a factor of $0.1$.
+
+\begin{figure}[h]
+	\hfil\includegraphics[height=3.0in]{images/tabuSize_stepCount}\hfil
+	\caption
+	{
+	  The dependence of the quality of solutions on the size of the tabu list when using $ StepCount $ termination strategy
+	}
+	\label{aba:fig2}
+\end{figure}
 
@@
 
-\begin{figure}[h]
-	\hfil\includegraphics[height=3.0in]{images/tabuSize_stepCount}\hfil
-	\caption
-	{
-	  The dependence of the quality of solutions on the size of the tabu list when using $ StepCount $ termination strategy
-	}
-	\label{aba:fig1}
-\end{figure}
-
 As should have been expected, for large values of the parameter a jam occurs in local optimum due to the large number of restrictions of movement in space.
@@
 	}
-	\label{aba:fig1}
+	\label{aba:fig3}
 \end{figure}
@@
 	}
-	\label{aba:fig1}
+	\label{aba:fig4}
 \end{figure}
@@
 
+We proceed to the description of the algorithm in scheme \ref{alg:constructSolution}. Let $U^{opt}$ is a sorted array of optimal solutions by value of the objective function ascending,  ie:
+\begin{equation} \label{u_sorted}
+F(u^{opt}_1) \leq F(u^{opt}_2) \leq \ldots < F(u^{opt}_i) \leq \ldots \leq F(u^{opt}_{sizeof(U^{opt})})
+\end{equation}
+Each solution $u^{opt}_i$ gets into the population with a given probability $p^i_u$, the probability of selection decreases with increasing sequence number:
+\begin{equation} \label{p_sorted}
+p^1_u = 1 > p^2_u > \ldots > p^i_u> \ldots   > p^{sizeof(U^{opt})}_u,  p^{sizeof(U^{opt})}_u > 0
+\end{equation}
+
 \begin{algorithm}[H]
@@
 	\textbf{function constructSolution}( $U^{opt}, p_u, n^{min}_s$) \\
-	// $U^{opt}$ -- an ordered set of optimal solutions \\
+	// $U^{opt}$ -- a sorted set array optimal solutions \\
 
@@
 		// construct solution using heuristics(for example, Solomon alg.)\\
-		$u^0 = heuristicsConscruction();$
-		return $u^0$;
+		$u^0 = heuristicsConscruction()$ \\
+		\textbf{return} $u^0$
 	}
@@
 		\If{$random(0.0,1.0) > p^i_u$}{
-		   i = i + 1; \\
-		   goto 7;
+		   i = i + 1 \\
+		   \textbf{goto} 8
 		}
-		$R = $randomly select a route from $u^{opt}_i$ solution;
-		$u^0 = u^0 \cup R$;
-		i = i + 1;\\
+		$R_i = $randomly select a route from $u^{opt}_i$ solution;
+		$u^0 = u^0 \cup R_i$
+		i = i + 1\\
 
 	}
-	\textbf{return} $u^0$;
+
+	\If{$u^0 is not complete$}{
+	   	$u^0 = heuristicsConscruction(u^0)$
+	}
+
+	\textbf{return} $u^0$
 
@@
 
-Let $\vec{\psi_i} = (\psi^1_i,\psi^2_i,\psi^3_i) = (\tilde{l_i},\tilde{p_i},\tilde{l_h})$ -- a set of values of TabuSearch configuration parameters which was using for solving the problem in the step $i$.
+It then crosses solutions by the following rule: with the first solution is randomly selected route $R_1$.
+Then another solution is selected from a route $R_2$ that does not include client requests from the first  route etc.
+If the client requests were not included in the solution $u^0$, then these requests are added by the construction heuristics algorithm (for example, Solomon algorithm).
+
+Now we describe the primary function $SolveCDP()$ of the modified algorithm presented in pseudocode form in Scheme \ref{alg:modTabuSearch}.
+In this function using a loop is a sequential formation of an array of the best solutions $U^{opt}$ by means TabuSearch algorithm.
+
+Based on the results of the experiments described in the preceding section, we decided to set the parameters of the algorithm are not fixed values, but as a random variable of predetermined period.
+For example, the size of the tabu list $h$ is defined as the interval $[10,35]$(see pict.
+\ref{aba:fig2} and  pict.
+\ref{aba:fig3}) and the neighbordhood size $l=N_{neighbordhood}$ is defined as the interval $[600\cdot n,850\cdot n]$, where $n$ -- problem size (see pict.
+\ref{aba:fig4} ).
+
+Let $\vec{\psi_i} = (\psi^1_i,\psi^2_i,\psi^3_i) = (\tilde{l_i},\tilde{p_i},\tilde{l_h})$ -- a set of values of the TabuSearch configuration parameters which was using for solving the problem in the step $i$.
 
 Let $[\psi^j_{begin}, \psi^j_{end}]$ -- an optimal interval of the TabuSearch $j$ -parameter.
-(For example: if $j=3$, then: $[\psi^31_{begin}, \psi^3_{end}]= [h_{begin}, h_{end}] = [10, 40] $).
+(For example: if $j=3$, then: $[\psi^3_{begin}, \psi^3_{end}]= [h_{begin}, h_{end}] = [10, 40] $).
 These intervals are initial data and they are set in the initialization block.
+
+The configuration parameters  $\psi^j_i $ are randomly selected from the interval $[\psi^j_{begin}, \psi^j_{end}]$ according to a distribution law.
+We proposed to build the distribution density $f_j(\psi^j_i)$ of the random variable $\psi^j_i$ as follows.
+At first, we define anchor points:
+
+ \begin{equation} \label{anchor_points}
+     \rho^j_i =  \dfrac{F(u^{opt}_1)}{avg[F(u^{opt}_k)]}, \forall k: \psi^j_k = \psi^j_i
+ \end{equation}
+ The number of anchor points $n_{\rho}$  is smaller than the set of optimal solutions, because a set of reference anchor points are removed the same point $(n_{\rho} \leq  n_u)$.
+
+The main idea is that the distribution density should be larger at those points (parameters $\psi^j_i$) in which high quality solutions than the points at which the low quality of solutions.
+Therefore, we presented the distribution function $f_j(z)$ as a sum of kernel functions:
+ \begin{equation} \label{dist_density}
+ f_j(z) = \alpha_j * \sum\limits_{i=1}^{n_{\rho}}  \rho^j_i * K(z -  \psi^j_i)
+ \end{equation}
+ where: $K(x)$ -- kernel function, $ \alpha$ -- a normalizing parameter
+
+ We have chosen as the kernel the function of parabolic type(known as an Epanechnikov function),  because during the experiments the best results were obtained using this function:
+ \begin{equation} \label{dist_density}
+  K(z) = 3/4 \cdot (1-z^2)
+ \end{equation}
 
@@
 
-	$\psi = \emptyset$;
-	$p_u = \emptyset$;
-	$U^{opt} = \emptyset$;
-	$i=0$; \\
+	$\psi = \emptyset$,
+	$p_u = \emptyset$,
+	$U^{opt} = \emptyset$,
+	$i=0$ \\
 
@@
 		// set TabuSearch parameters \\
-		\For{$j=1..3$}{
+		\For{$j=1 \ldots 3$}{
 	  	// $f_j$ -- the density distribution of the random \\
 	  	// variable $\psi^j$ (TabuSearch parameter $l,p $ or $ h$) \\
-	  	$\psi^j_i = random(\psi^j_{begin}, \psi^j_{end}, f_j); $
+	  	$\psi^j_i = random(\psi^j_{begin}, \psi^j_{end}, f_j) $
 	  	}
@@
 		// Improve solution $u^0$ using TabuSearch algorithm 
-		$u^{opt}_i = TabuSearch(u^0, \vec{\psi_i});$ \\
+		$u^{opt}_i = TabuSearch(u^0, \vec{\psi_i})$ \\
 
-		$U^{opt}  = U^{opt} \cup \{u^{opt}_i\};$ \\
+		$U^{opt}  = U^{opt} \cup \{u^{opt}_i\}$ \\
 
 		// Sort ascending optimums $U$ of the objective function \\
-		$U^{opt} = sort(U^{opt});$\\
-		$p^{i+1}_u = update(p^i_u)$;\\
+		$U^{opt} = sort(U^{opt})$\\
+		$p^{i+1}_u = update(p^i_u)$\\
 
-		i = i + 1;\\
+		i = i + 1\\
 
 	}
-	\textbf{return} $u^{opt}_1$;
+	\textbf{return} $u^{opt}_1$
 
@@
 
+ The parameter $\alpha_j$ was introduced in the density function to perform the normalization condition:
+ \begin{equation} \label{norm_cond}
+ \int\limits_{\psi^j_{begin}}^{\psi^j_{end}} f_j(z)dz = 1  =>  (\alpha_j)^{-1} = \sum\limits_{i=1}^{n_{\rho}} \rho^j_i \cdot  \int\limits_{\psi^j_{begin}}^{\psi^j_{end}} K(z -  \psi^j_i) dz 
+ \end{equation}
+
+ The parameter $\alpha_j$ can be written in explicit form:
+  \begin{equation} \label{norm_cond}
+ \begin{split}
+ (\alpha_j)^{-1} & = \dfrac{3}{4} \cdot \sum\limits_{i=1}^{n_{\rho}} \rho^j_i \cdot [(1 - (\psi^j_i)^2) \cdot (\psi^j_{end} - \psi^j_{begin}) + \\
+ & + \psi^j_i \cdot ((\psi^j_{end})^2 - (\psi^j_{begin})^2) - 1/3 \cdot (\psi^j_{end})^3 - (\psi^j_{begin})^3)]
+ \end{split}
+ \end{equation}
+
+ In figure \ref{aba:fig5} the results of an experiment in which  were presented a dependence of the quality of the solutions obtained by the classical and modified algorithms and  the breaking point time.
+As seen from the graph, with an increase in operating time of the computational scheme, the quality of solutions obtained using the modified algorithm is growing faster than using classical Tabu Search algorithm.
+
+ \begin{figure}[h]
+ 	\hfil\includegraphics[height=2.0in]{images/acceptedCountLimit}\hfil
+ 	\caption
+ 	{
+ 		The dependence of the quality of solutions and  the breaking point time using $ TimeMillisSpent $ termination strategy
+ 	}
+ 	\label{aba:fig5}
+ \end{figure}
+
 \subsection{Conclusion}
-In the process, has been studied and implemented tabu search algorithm to solve the static problem with time constraints.
-Modification of the algorithm to optimize the computational scheme.
-Modified algorithm gave the best solutions in terms of the balance between the number of vehicles and the distance traveled.
-such a feature of this algorithm can be used for decision support in intelligent systems.
-In practice, the more advantageous to use more vehicles if this will reduce the time to use them, this will reduce fuel costs and depreciation of transport;
+During the research has been investigated and implemented tabu search algorithm to solve the Static Routing Courier Delivery problem with time constraints.
+Also a modified algorithm was developed  based on the Tabu Search algorithm to improve the quality of the solutions.
+Modified algorithm gave the best solutions in terms of the balance between the number of vehicles and the cost traveled.
+In practice this algorithm can be used for decision support in intelligent systems for improving the quality of customer service and for reducing waiting the time to use vehicle, this will reduce fuel costs and depreciation of transport.
 
 The analysis of the parameters of implemented algorithms allowed us to determine their optimal values for this class of routing problems.
-With the modified algorithm was found solutions of model problems, which in most cases have an acceptable deviation from the global optimum;
+With the modified algorithm was found solutions of model problems, which in most cases have an acceptable deviation from the global optimum.
 
@@
 
+\bibitem{RCDS} F.~Ordonez, Chen Wang, A New Approach for Routing Courier Delivery Services //METRANS Transportation Center:University of Southern California
+Los Angeles,  2012, 81--115.
+
 \bibitem{VRPTW} T.~Babb, Pickup and Delivery Problem with Time Windows // Coordinated Transportation Systems: The State of the Art.
@@
 
+\bibitem{DELIVERY_2000} W~Barnes, Solving the pickup and delivery problem with time windows using reactive tabu search // Transportation Research Part B: Methodological, Vol.
+34 Issue 2, 2000, p.
+107--121.
+% http://www.sciencedirect.com/science/article/pii/S0191261599000168
+
 \bibitem{Tabu_search} O.~Braysy, M.~Gendreau, Vehicle Routing Problem with Time Windows, Part I: Route Constuction and local algorithms // Transportation science Vol.39 No. 1, 2005, p. 104-118.
-
-\bibitem{Assignment_Problems} E.~Rainer, Assignment problems, 2009, 402 p.
 
@@
 \bibitem{problems_Christofides}  N. Christofides, S. Eilon, An algorithm for the vehicle dispatching problem //Operational Research Quarterly, 20, 1969, p. 309–318.
+
 \bibitem{problems_Golden}  B.
@@
 33–56.
+
 \bibitem{problems_Taillard}  E. Taillard. VRP benchmarks.\\ http://mistic.heig-vd.ch/taillard/problemes.dir/vrp.dir/vrp.html, 1993.

### 2015-02-05 article.tex
edit literature list

@@
 
-In the first computational experiment we made finding the best solutions for different values of the tabu size parameter $h \in [0, 40]$ (pict.
-1 an pict.
-2).
 In this experiment we use two different breakpoint strategies: 
@@
 	\item $StepCount$ termination strategy --  terminates when an amount of steps has been reached;
-	\item $TimeMillisSpent$ termination strategy -- terminates when an amount of time has been used.
+	\item $TimeSpent$ termination strategy -- terminates when an amount of time has been used.
 \end{enumerate}
@@
 Otherwise if you select a too small tabu size, algorithm can still get stuck in a local optimum.
+In the first computational experiment we made finding the best solutions for different values of the tabu size parameter from interval $h \in [0, 40]$ (pict.
+1 an pict.
+2).
 \begin{figure}[h]
-	\hfil\includegraphics[height=2.0in]{images/tabuSize_time}\hfil
+	\hfil\includegraphics[height=2.2in]{images/tabuSize_time}\hfil
 	\caption
 	{
-	  The dependence of the quality of solutions on the size of the tabu list when using $ TimeMillisSpent $ termination strategy
+	  The dependence of the quality of solutions on the size of the tabu list when using $ TimeSpent $ termination strategy
 	}
@@
 
-In the second experiment we built a relationship between the number $N_{neighbors}$ of viewed solutions neighborhood $N_l(\vec{u})$ using $TimeMillisSpent$ termination strategy. 
 \begin{figure}[h]
-	\hfil\includegraphics[height=2.0in]{images/acceptedCountLimit}\hfil
+	\hfil\includegraphics[height=2.2in]{images/acceptedCountLimit}\hfil
 	\caption
 	{
-	  The dependence of the quality of solutions on the neighborhood size when using $ TimeMillisSpent $ termination strategy
+	  The dependence of the quality of solutions on the neighborhood size when using $ TimeSpent $ termination strategy
 	}
@@
 \end{figure}
+
+In the second experiment (see figure \ref{aba:fig4}) we built a relationship between the number $N_{neighborhood}$ of viewed solutions neighborhood size $N_l(\vec{u})$ using $TimeSpent$ termination strategy.
+In this experiment we made finding the best solutions for different values of the neighborhood size parameter from interval $N_{neighborhood} \in [750, 850]$ (pict.
+1 an pict.
+2).
 
@@
  \begin{figure}[h]
- 	\hfil\includegraphics[height=2.0in]{images/acceptedCountLimit}\hfil
+ 	\hfil\includegraphics[height=3.0in]{images/tabuSize_time.png}\hfil
  	\caption
  	{
- 		The dependence of the quality of solutions and  the breaking point time using $ TimeMillisSpent $ termination strategy
+ 		The dependence of the quality of solutions and  the breaking point time by classic and modified algorithms using $ TimeSpent $ termination strategy
  	}
@@
 
-\bibitem{VRPTW} T.~Babb, Pickup and Delivery Problem with Time Windows // Coordinated Transportation Systems: The State of the Art.
+\bibitem{Delivery_Babb_2005} T.~Babb, Pickup and Delivery Problem with Time Windows // Coordinated Transportation Systems: The State of the Art.
 Department of Computer Science University of Central Florida Orlando, Florida, 2005, 38 p.
 
-\bibitem{DELIVERY_2000} W~Barnes, Solving the pickup and delivery problem with time windows using reactive tabu search // Transportation Research Part B: Methodological, Vol.
+\bibitem{Delivery_Barnes_2000} W~Barnes, Solving the pickup and delivery problem with time windows using reactive tabu search // Transportation Research Part B: Methodological, Vol.
 34 Issue 2, 2000, p.
 107--121.
+
+\bibitem{Delivery_Masson_2014} R.~Masson, F.~Lehued, O.~ Peton,The dial-a-ride problem with transfers. // Computers and Operations Research, Vol. 41, 2014, p. 12--23.
 % http://www.sciencedirect.com/science/article/pii/S0191261599000168
+
+\bibitem{jlogistics}  R. Shafeyev. Java-based framework for sloving routing problems.\\ url: http://jlogistics.net, 2015.
+
+\bibitem{Delivery_Coltin_2014} B.~Coltin, M.~Veloso, Scheduling for Transfers in Pickup and Delivery Problems with Very Large Neighborhood Search // The Twenty-Eighth Conference on Artificial Intelligence, Quebec City, Canada, 2014, 7 p.
+% http://www.cs.cmu.edu/~mmv/papers/14aaai-ColtinVeloso.pdf
 
@@
 
-\bibitem{problems_Taillard}  E. Taillard. VRP benchmarks.\\ http://mistic.heig-vd.ch/taillard/problemes.dir/vrp.dir/vrp.html, 1993.
+\bibitem{problems_Taillard}  E. Taillard. VRP benchmarks.\\ url: http://mistic.heig-vd.ch/taillard/problemes.dir/vrp.dir/vrp.html, 1993.
+
+\bibitem{opt_Goncharov} E.~Goncharov, Y~Kochetov, Probabilistic search with exclusions for discrete unconstrained optimization  //Discrete Analysis and Operations Research, Moscow, Serial 2.
+Vol.
+9, 2002, p.
+13-30.
+%Гончаров Е.Н., Кочетов Ю.А. Вероятностный поиск с запретами для дискретных задач безусловной оптимизации // Дискрет. анализ и исслед. операций. Сер. 2. Т.9, № 2. 2002, с. 13-30.
+
+\bibitem{opt_Kochetov}   Y~Kochetov, Probabilistic methods of local search for discrete optimization problems //Discrete Mathematics and Its Applications: Proceedings of youth lectures and scientific schools in discrete mathematics and its applications, Publishing House of the Center for Applied Research at the Mechanics and Mathematics.
+Faculty of Moscow State University, 2001, p.
+84-117.
+%Кочетов Ю.А.
+Вероятностные методы локального поиска для задач дискретной оптимизации // Дискретная математика и ее приложения: Сб.
+лекций молодежных и научных школ по дискретной математике и ее приложениям.
+М.: Изд-во центра прикл.
+исслед.
+при мех-мат.
+фак.
+МГУ, 2001, с.
+84-117;
+
 \end{thebibliography}

### 2015-02-10 article.tex
add assign problem section

@@
 \usepackage[linesnumbered,boxed]{algorithm2e}
+\pagestyle{plain}
 \begin{document}
@@
 
-Denote by $\delta(\vec{u},\vec{v})$ the Hamming distance between $\vec{u}$ and $\vec{v}$. Through $N_l(\vec{u})$ we denote the neighborhood of a point $\vec{u}$ radius $l$, ie:
+Denote by $\delta(\vec{u},\vec{v})$ the Hamming distance between $\vec{u}$ and $\vec{v}$. Through $N_l(\vec{u})$ we denote the neighborhood of a point $\vec{u}$ radius $l$, ie \cite{vrp_shafeyev}:
 \begin{align} 
@@
 The standard algorithm local search algorithm starts with a randomly selected vector $\vec{u^0}$.
-On the $i$ step of the algorithm move from the current vector in the neighbouring that the minimum value of the objective function in the neighborhood of the given vector:
+On the $i$ step of the algorithm move from the current vector in the neighboring that the minimum value of the objective function in the neighborhood of the given vector:
 \begin{align} 
@@
 		{
-			// find optimum into the neighbord $N_l$: \\
+			// find optimum into the neighborhood $N_l$: \\
 			$ u^{i+1} : F(u^{i+1}) = min\{ F(y): y \in N_l \}$
@@
 
-\caption{Pseudo-code for probabilistic tabu-search algorithm.}
+\caption{Pseudo-code for probabilistic Tabu Search algorithm.}
 \label{alg:TabuSearch}
@@
 \subsection{Analysis of the algorithm parameters}
-Modified tabu search algorithm has 3 main parameters: $p$ -- the probability of hitting solutions in a neighborhood, $h$ -- size of the list of prohibitions and $N_{neighbors}$ -- accepted count limit of size of neighbordhood $N_l(\vec{u})$.
+The Tabu Search algorithm has two main parameters: $h$ -- size of the list of prohibitions and $N_{neighbors}$ -- accepted count limit of size of neighbordhood $N_l(\vec{u})$.
 For the experiments, we constructed the test tasks on the basis of the capacity routing problems developed by Breedam, Fisher, Christofides and Eilon.
-To determine trends in time and cost of the decision to use exponential smoothing by a factor of $0.1$.
+To determine trends in time and cost of the decision we used exponential smoothing by a factor of $0.1$.
 
@@
 	\item $StepCount$ termination strategy --  terminates when an amount of steps has been reached;
-	\item $TimeSpent$ termination strategy -- terminates when an amount of time has been used.
+	\item $TimeSpent$ termination strategy -- terminates when an amount of time has been reached.
 \end{enumerate}
@@
 In the first computational experiment we made finding the best solutions for different values of the tabu size parameter from interval $h \in [0, 40]$ (pict.
-1 an pict.
-2).
+\ref{aba:fig2} an pict.
+\ref{aba:fig3}).
 \begin{figure}[h]
@@
 
-In the second experiment (see figure \ref{aba:fig4}) we built a relationship between the number $N_{neighborhood}$ of viewed solutions neighborhood size $N_l(\vec{u})$ using $TimeSpent$ termination strategy.
-In this experiment we made finding the best solutions for different values of the neighborhood size parameter from interval $N_{neighborhood} \in [750, 850]$ (pict.
-1 an pict.
-2).
+In the second experiment (see figure \ref{aba:fig4}) we built a relationship between the number $N_{neigh.}$ of viewed solutions neighborhood size $N_l(\vec{u})$ using $TimeSpent$ termination strategy.
+In this experiment we made finding the best solutions for different values of the neighborhood size parameter from interval $N_{neigh.}\in[750, 850]$.
 
@@
 
-Modification of the scheme is based on the hypothesis of "On a large valley." According to this hypothesis, the average local optima are located much closer to the global than a randomly chosen point.
+Modification of the scheme is based on the hypothesis of "On a large valley"\cite{opt_Kochetov}.
+According to this hypothesis, the average local optima are located much closer to the global than a randomly chosen point.
 There is a certain concentration of local optimum in a small part of the feasible region, which is figuratively called a large valley.
@@
 	\textbf{function constructSolution}( $U^{opt}, p_u, n^{min}_s$) \\
-	// $U^{opt}$ -- a sorted set array optimal solutions \\
+	// $U^{opt}$ -- a sorted set array of optimal solutions \\
 
@@
  \end{equation}
- The number of anchor points $n_{\rho}$  is smaller than the set of optimal solutions, because a set of reference anchor points are removed the same point $(n_{\rho} \leq  n_u)$.
+ The number of anchor points $n_{\rho}$  is smaller than the set of optimal solutions, because a set of  anchor points are removed the same points $(n_{\rho} \leq  n_u)$.
 
@@
  \begin{figure}[h]
- 	\hfil\includegraphics[height=3.0in]{images/tabuSize_time.png}\hfil
+ 	\hfil\includegraphics[height=3.0in]{images/tabu_mod.png}\hfil
  	\caption
@@
 
+\bibitem{Delivery_Masson_2014} R.~Masson, F.~Lehued, O.~ Peton,The dial-a-ride problem with transfers. // Computers and Operations Research, Vol. 41, 2014, p. 12--23.
+% http://www.sciencedirect.com/science/article/pii/S0191261599000168
+
 \bibitem{Delivery_Babb_2005} T.~Babb, Pickup and Delivery Problem with Time Windows // Coordinated Transportation Systems: The State of the Art.
 Department of Computer Science University of Central Florida Orlando, Florida, 2005, 38 p.
+
+\bibitem{vrp_shafeyev} R.~Shafeyev ,L.~Lyubchik, A some realization of Tabu Search algorithm for Solving the Transportation Problem with Time Constraints // Vestnik NTU "KhPI". – Kharkov: NTU
+"KhPI", 2013. – No3 (977). – p. 35–39.
+
+\bibitem{jlogistics}  R. Shafeyev. Java-based optimixation framework for solving routing problems.\\ url: http://jlogistics.net, 2015.
 
@@
 
-\bibitem{Delivery_Masson_2014} R.~Masson, F.~Lehued, O.~ Peton,The dial-a-ride problem with transfers. // Computers and Operations Research, Vol. 41, 2014, p. 12--23.
-% http://www.sciencedirect.com/science/article/pii/S0191261599000168
-
-\bibitem{jlogistics}  R. Shafeyev. Java-based framework for sloving routing problems.\\ url: http://jlogistics.net, 2015.
-
 \bibitem{Delivery_Coltin_2014} B.~Coltin, M.~Veloso, Scheduling for Transfers in Pickup and Delivery Problems with Very Large Neighborhood Search // The Twenty-Eighth Conference on Artificial Intelligence, Quebec City, Canada, 2014, 7 p.
 % http://www.cs.cmu.edu/~mmv/papers/14aaai-ColtinVeloso.pdf
+
+\bibitem{opt_Goncharov} E.~Goncharov, Y~Kochetov, Probabilistic search with exclusions for discrete unconstrained optimization  //Discrete Analysis and Operations Research, Moscow, Serial 2.
+Vol.
+9, 2002, p.
+13-30.
+%Гончаров Е.Н., Кочетов Ю.А. Вероятностный поиск с запретами для дискретных задач безусловной оптимизации // Дискрет. анализ и исслед. операций. Сер. 2. Т.9, № 2. 2002, с. 13-30.
+
+\bibitem{opt_Kochetov}   Y~Kochetov, Probabilistic methods of local search for discrete optimization problems //Discrete Mathematics and Its Applications: Proceedings of youth lectures and scientific schools in discrete mathematics and its applications, Publishing House of the Center for Applied Research at the Mechanics and Mathematics.
+Faculty of Moscow State University, 2001, p.
+84-117.
+%Кочетов Ю.А.
+Вероятностные методы локального поиска для задач дискретной оптимизации // Дискретная математика и ее приложения: Сб.
+лекций молодежных и научных школ по дискретной математике и ее приложениям.
+М.: Изд-во центра прикл.
+исслед.
+при мех-мат.
+фак.
+МГУ, 2001, с.
+84-117;
 
@@
 
-\bibitem{opt_Goncharov} E.~Goncharov, Y~Kochetov, Probabilistic search with exclusions for discrete unconstrained optimization  //Discrete Analysis and Operations Research, Moscow, Serial 2.
-Vol.
-9, 2002, p.
-13-30.
-%Гончаров Е.Н., Кочетов Ю.А. Вероятностный поиск с запретами для дискретных задач безусловной оптимизации // Дискрет. анализ и исслед. операций. Сер. 2. Т.9, № 2. 2002, с. 13-30.
-
-\bibitem{opt_Kochetov}   Y~Kochetov, Probabilistic methods of local search for discrete optimization problems //Discrete Mathematics and Its Applications: Proceedings of youth lectures and scientific schools in discrete mathematics and its applications, Publishing House of the Center for Applied Research at the Mechanics and Mathematics.
-Faculty of Moscow State University, 2001, p.
-84-117.
-%Кочетов Ю.А.
-Вероятностные методы локального поиска для задач дискретной оптимизации // Дискретная математика и ее приложения: Сб.
-лекций молодежных и научных школ по дискретной математике и ее приложениям.
-М.: Изд-во центра прикл.
-исслед.
-при мех-мат.
-фак.
-МГУ, 2001, с.
-84-117;
-
 \end{thebibliography}

### 2016-05-03 article.tex
begin to translate article to russian
merge

@@
 
-Thus, each application can be represented as a tuple:
+Thus, each request can be represented as a tuple:
 \begin{equation}
-\forall q \in Q: \exists q = (s_q,r_q, \vec{P}_{s_q}, \vec{P}_{r_q}, w_q, t_{s}^{q}, \Delta{t_{s}^{q}}, t_{r}^{q}, \Delta{t_{r}^{q}})
+\forall q \in Q: q = (s_q,r_q, \vec{P}_{s_q}, \vec{P}_{r_q}, w_q, t_{s}^{q}, \Delta{t_{s}^{q}}, t_{r}^{q}, \Delta{t_{r}^{q}})
 \end{equation}
@@
 
-The algorithm terminates at a local optimum, when $F(\vec{x^{i+1}})=F(\vec{x^i})$.
+The algorithm terminates at a local optimum, when $F(\vec{u^{i+1}})=F(\vec{u^i})$.
 When solving the transportation problem is a typical situation when many local optimums and only one of them is global:
@@
 
-As a result the client $u^*$ is added to the current route $k$ the position between nodes $v_i$ and $v_i+1$.
+As a result the client $u^*$ is added to the current route $k$ the position between nodes $v_i$ and $v_(i+1)$.
 
@@
 	\hfil\includegraphics[height=2.2in]{images/tabuSize_time}\hfil
-	\caption
-	{
+	\caption{
 	  The dependence of the quality of solutions on the size of the tabu list when using $ TimeSpent $ termination strategy
@@
 	\hfil\includegraphics[height=2.2in]{images/acceptedCountLimit}\hfil
-	\caption
-	{
+	\caption{
 	  The dependence of the quality of solutions on the neighborhood size when using $ TimeSpent $ termination strategy
 	}
+
 	\label{aba:fig4}
@@
 
-\bibitem{Goldberg_Kennedy} V.~Goldberg, R.~Kennedy, An Efficient cost scaling algotirhm for the assignment problem, Math. Program., 1995, p. 153--177.
+\bibitem{Goldberg_Kennedy} V.~Goldberg, R.~Kennedy, An Efficient cost scaling algorithm for the assignment problem, Math. Program., 1995, p. 153--177.
 

### 2016-05-03 article_ru.tex

--- 2015-02-10 article.tex
@@
 %We are specifying language of the paper
-\selectlanguage{english}
+\selectlanguage{russian}
 \def\dd#1#2{\frac{\partial#1}{\partial#2}}
 \section{
-%%%%%%%%%%%%%%% You have to specify name of your paper here
-``Investigation of tuning parameters of tabu search algorithm and its modification  for solving the static Routing Courier Delivery Problem.'' 
+``Исследование  настроечных параметров алгоритма поиска с запретами и его модификация для решения курьерской задачи маршрутизации.'' 
 }
@@
 \authors{
-R.~Shafeyev , L.~Lyubchik 
+L.~Lyubchik, R.~Shafeyev 
 }
@@
 
-The paper considers the routing courier delivery problem with the service time for which the discrete model was constructed and the calculation scheme based on tabu search algorithm.
-The efficiency  of the proposed algorithm‘s scheme has been tested on a large-scale problems, which were generated on the basis of different classes known model routing problems.
-
+В статье рассматривается задача маршрутизации курьерской доставки  с долгим сроком службы, для которой была построена дискретная модель и вычислительная схема, в основу которой был положен алгоритм поиска с запретами.
+Эффективность схемы предложенного алгоритма была проверена на задачах большой размерности, которые были составлены на базе известных модельных задачах маршрутизации на транспорте различных классов.
 }
@@
 %%%%% Paragraph text. To separate paragraphs use empty lines.
-In the research we consider a static vehicle routing problem with time windows.
-This variant of the routing problem is a common problem in the courier delivery industry.
-Effective solution to this problem affects the quality of services provided by the company, as well as the magnitude of potential losses caused by the excess of the agreed terms of delivery.
-The proposed solution to the problem is based on the automation of the process of planning the route and involves the use of distributed information systems for the possible implementation of these approaches in practice.
-Another important criterion for good transport plan are low transport costs, which affect the cost of services.
+Одним из средств экономии ресурсов при транспортировке грузов является применение систем поддержки принятия решений в сфере транспортной логистики.
+Одной из ключевых функций систем поддержки принятия решений в сфере транспортной логистики является возможность расчета и построения эффективных с точки зрения стоимости объезда маршрутов различного назначения на транспортной сети.
+Математическая формулировка этой задачи широко известно как задача маршрутизации транспорта.
+Существует ряд разновидностей ЗМТ с различными дополнительными условиями, которые позволяют учитывать грузоподъемность транспортных средств и другие ограничения для более полного представления деталей реальной действительности.
 
-The Routing Courier Delivery Problem with time windows requires that a set of vehicles satisfy a collection of customer requests.
-Each client request requires the use of a single vehicle to receive and load a specified amount of goods at one source location(sender) and to deliver them to destination location(receiver).
-All requests must be performed without violating either the vehicle capacity or the customer time window.
-The solution is a master plan that solves the problem, and takes into account the daily events that take place according to plan.
-In this paper, we present a discrete model and a modified tabu search algorithm  to solve the  Courier Delivery Problem with time windows.
+Данная статья посвящена рассмотрению задачи курьерской маршрутизации, которая является подзадачей ЗМТ.
+Данная задача состоит в нахождении маршрутов для посещения заданного множества адресов некоторым количеством единиц транспортных средств.
+Такая проблема наиболее часто встречается среди компаний, выполняющих развозки товара от отправителя к получателю.
 
-In order to validate our strategies decisions technique effectiveness, we have constructed a new set of benchmark problems for the pickup and delivery problem with time windows based on which were generated on the basis of different classes known model routing problems.
-Computational results substantiate the solution quality and efficiency of our modified tabu search algorithm.
+Автоматизация процесса планирования маршрутов будет актуальной для:  интернет-магазинов, крупных оптовых компаний (например, для развозки товаров, которые быстро портятся или оптимизации схемы доставки товаров в пункты назначения), фирм с целью организации допоставок в магазины.
+Практическое применение задачи можно встретить в сфере транспортировки пациентов между корпусами поликлиники\cite{Beaudry, Kergosien}, доставке газет и журналов, доставки топлива в частные дома \cite{Sarak}.
+Решение рассматриваемой задачи также имеет практическое применение в системах вертолетных перевозок людей между морскими нефтяными платформами \cite{Romero}.
 
-\subsection{Problem Statement}
-Let $C,dim(C)=n$ is a set of vehicles and $Q,dim(Q)=m$ is a set of requests received from customers at the current time.
+В последнее время все больше людей пользуются услугами курьерских служб.  Обычно заявки формируются днем ранее и
+перед диспетчером полный список заявок, которые необходимо выполнить в кратчайшие сроки.
+Тематическое исследование провел Attanasio(2007) \cite{Attanasio}, в котором описал преимущества использования компьютерных методов перед человеческой диспетчеризацией.
+В своей работе он рассмотрел примеры работы eCourier Ltd, лондонской компании, которая предлагает курьерские услуги.
+Их клиентами в основном являются юридические фирмы, финансовые учреждения, рекламные агентства и другие организации, которые заинтересованы в быстрой доставке товаров или оригинальных подписях на документах.
+После того как все заявки на конкретный момент времени сформированы, необходимо составить такой маршрут, который удовлетворял бы требованиям заказчиков.
+В зависимости от уровня обслуживания, который был задан клиентом, курьер может объединить или ни его доставку с другими, то есть это означает, что товары могут быть доставлены по очереди, или курьер может выполнить ускоренную доставку.
 
-Suppose that the following information is known about vehicles from set $C$: \\
-$\vec{P}_c$ -- vehicle position, $c \in C$;\\
-$L_c$ -- vehicle capacity, $c \in C$.
+Компании, которые предлагают курьерские услуги достаточно часто имеют смешанный парк транспортных средств, который состоит из велосипедов, мотоциклов, автомобилей и небольших грузовиков.
+В зависимости от типа запроса, расположение и временных окон будет выбран тот вид транспортного средства, который сможет выполнить заказ в кратчайшие сроки.
+Исследования \cite{Attanasio} показали, что использование компьютерных методов, в том числе алгоритмов оптимизации, были очень выгодными для курьерских компаний.
 
-Let $S$ is a set of senders, ($dim(S) = dim(Q) = m$), $R$ is a set of receivers, $dim(R) = dim(Q) = m$. Then each client request $q \in Q$ includes the following information:\\
-$s_q$ -- a sender of client shipment, $s \in S$; \\
-$r_q$ -- a receiver of client shipment, $r \in R$; \\
-$\vec{P}_{s_q}$ -- sender position;\\
-$\vec{P}_{r_q}$ -- receiver position; \\
-$w_q$ -- client shipment that is required to deliver from the sender to the recipient; \\
-$[t_{s}^{q}, t_{s}^{q}+ \Delta{t_{s}^{q}}]$ -- the time window within which worker must to pick up the goods from the sender;\\ 
-$[t_{r}^{q}, t_{r}^{q}+ \Delta{t_{r}^{q}}]$ -- the time window within which worker must deliver the goods to the receiver.
+Таким образом, использование автоматизированной системы позволяет улучшить качество обслуживания, время доставки, повысить курьерскую эффективность и сократить расходы на доставку, тем самым обеспечить повышение конкурентного преимущества.
 
-Thus, each application can be represented as a tuple:
+\subsection{Постановка задачи}
+Пусть $C,dim(C)=n$ -- множество транспортных средств, $Q,dim(Q)=m$ -- множество запросов клиентов, которые требуется обработать на текущий момент времени.
+
+Предположим, что следующая информация известна о транспортных средствах из множества $C$: \\
+$\vec{P}_c$ -- местоположение ТС, $c \in C$;\\
+$L_c$ -- грузоподъемность ТС, $c \in C$.
+
+Пусть $S$ -- множество отправителей груза, ($dim(S) = dim(Q) = m$), $R$ -- множество получателей, $dim(R) = dim(Q) = m$.
+Тогда каждый клиентский запрос $q \in Q$ включает в себя следующую информацию:\\
+$s_q$ -- отправитель груза клиента, $s \in S$; \\
+$r_q$ -- получатель клиентского груза, $r \in R$; \\
+$\vec{P}_{s_q}$ -- Местоположение отправителя;\\
+$\vec{P}_{r_q}$ -- Местоположение получателя; \\
+$w_q$ -- клиентский груз, который требуется перевезти от отправителя к получателю; \\
+$[t_{s}^{q}, t_{s}^{q}+ \Delta{t_{s}^{q}}]$ -- временное окно, в течение которого нужно забрать груз у отправителя;\\ 
+$[t_{r}^{q}, t_{r}^{q}+ \Delta{t_{r}^{q}}]$ -- временное окно, в течение которого нужно доставить груз получателю.
+
+Таким образом, каждая заявка может быть представлена в виде кортежа:
 \begin{equation}
@@
 \end{equation}
-To estimate the cost of transportation between destinations defined cost function $\Omega$:
+Для того, чтобы оценить стоимость перевозки между пунктами назначения, определим функцию стоимости $\Omega$:
 \begin{equation}
@@
 
-Necessary to construct the optimal routes of vehicles movement for the transportation of goods from the sender to the receiver for all client requests.
+Необходимо построить оптимальные маршруты движения транспортных средств для перевозки грузов от отправителя к получателю для всех клиентских запросов.
 
@@
 
-\subsection{Discrete model}
-Routing problem can be represented as a directed graph $G=G(V,E)$.
-The set $V=C\cup{S}\cup{R}$ -- nodes of the graph $G$, consisting of elements of vehicles, senders and receivers.
-$E$ -- dynamic set of arcs of the graph $G$, such that:
+\subsection{Дискретная модель}
+Задачу маршрутизации можно представить в виде ориентированного графа $G=G(V,E)$.
+Пусть множество $V=C\cup{S}\cup{R}$ -- узлы графа $G$,  состоящее из элементов множеств транспортных средств $C$, отправителей $S$ и получателей $R$.
+$E$ -- динамическое множество дуг графа $G$, такое, что:
 \begin{equation}
@@
 
-Let $X = \{X^k\}^n_{k=1}$ is a sequence of variable matrices for each vehicle $k \in C$. Elements of the matrices take the following values:
+Пусть $X = \{X^k\}^n_{k=1}$ представляет собой последовательность матриц для каждого ТС $k \in C$. Элементы матрицы принимают следующие значения:
 \begin{equation}
@@
     \begin{cases}
-	  1,&\text{the vehicle $k \in C$ moves from $i$ node to the $j$}\\
-	  0,&\text{otherwise.}
+	  1,&\text{АТС $k \in C$ движется с узла $i$ в вершину $j$}\\
+	  0,&\text{в противном случае.}
     \end{cases}
 \end{equation}
-where: $i\in{V/C \cup {k}}, j \in V/C$.
+где: $i\in{V/C \cup {k}}, j \in V/C$.
 
-Let us introduce the vector  of variables $ \vec{Y}  ^ k(X) $ for each vehicle $ k \ in C $. Vector elements have the following values:
+Введем вектор переменных $ \vec{Y}  ^ k(X) $ для каждого транспортного средства  $ k \ in C $. Векторные элементы принимают следующие значения:
 \begin{equation}
@@
 \begin{cases}
-1,&\text{the request $j \in Q$ is processed by vehicle $k \in C$}\\
-0,&\text{otherwise.}
+1,&\text{Запрос $j \in Q$ обслуживается ТС $k \in C$}\\
+0,&\text{в противном случае}
 \end{cases}
 \end{equation}
-where: $i\in{V/C \cup {k}}, j \in V/C$.
+где: $i\in{V/C \cup {k}}, j \in V/C$.
 
-Let $t^k_j(X^{(k)})$ -- arrival time of the vehicle $k \in C$ at the destination $j \in S \cup R$.
+Пусть $t^k_j(X^{(k)})$ -- время прибытия автомобиля $k \in C$ в пункт назначения $j \in S \cup R$.
 
-The objective function takes the following form:
+Целевая функция принимает следующий вид:
 \begin{equation} \label{main_objective}
@@
 
-We define constraints on the objective function (\ref{main_objective}), which provided the continuity of routes:
+Наложим ограничения на целевую функцию  (\ref{main_objective}) для обеспечения непрерывности маршрутов:
 \begin{align} 
@@
 
-The restriction (\ref{main_cond_1}) prohibits to a node in the graph $G$ has more than one output arc.
-The restriction (\ref{main_cond_2}) prohibits to a node has more than one input arc.
-The constraint (\ref{main_cond_3}) indicates that the number of input arcs to the node can not be less than the output arcs (this constraint considers the fact that the vehicle can leave the destination only if it has visited this node).
-The restriction (\ref{main_cond_4}) excludes local loops.
+Ограничение (\ref{main_cond_1}) запрещает, чтобы у узла графа $G$ было более чем одна выходящая дуга.
+Ограничение (\ref {main_cond_2}) запрещает узлу иметь более чем одну входную дугу.
+Ограничение (\ref {main_cond_3}) указывает на то, что у узла количество входных дуг не может быть меньше количества выходных дуг (это ограничение учитывает тот факт, что транспортное средство может покинуть пункт назначения, если только он посетил этот узел).
+Ограничение (\ref {main_cond_4}) исключает локальные циклы.
 
-Then next constraints synchronize values of variables $X$ and $\vec{y}$ for each request $q \in Q$ and prohibits the service of receiver before a visit to the sender:
+Следующие ограничения синхронизируют значения переменных $X$ и $ \Vec{у} $ для каждого запроса $ Q \ в Q $ и запрещают посещение получателя перед отправителем:
+
 \begin{align} 
@@
 
-Define restrictions for accounting of vehicle capacity and time windows:
+Определим ограничения для учета грузоподъемности транспортных средств и временные окна:
 \begin{align} 
@@
 
-\subsection{Description of the Algorithm}
-Present the set of matrices $\{X^k\}$ in the form of a vector defined on the hypercube $E^\eta=\{0.1\}^\eta$, where $\eta=n\cdot(2m+1)\cdot 2m$.
-The aim of the task is to minimize the objective function:
+\subsection{Описание алгоритма}
+Представим множество матриц $\{X^k\}$ в виде вектора, определенного на гиперкубе $E^\eta=\{0.1\}^\eta$, где $\eta=n\cdot(2m+1)\cdot 2m$. Цель задачи состоит в минимизации следующей целевой функции:
+
 \begin{align} 
@@
 
-Denote by $\delta(\vec{u},\vec{v})$ the Hamming distance between $\vec{u}$ and $\vec{v}$. Through $N_l(\vec{u})$ we denote the neighborhood of a point $\vec{u}$ radius $l$, ie \cite{vrp_shafeyev}:
+Обозначим через $\delta(\vec{u},\vec{v})$ расстояние Хэмминга между $\vec{u}$ и $\vec{v}$. Через $N_l(\vec{u})$ обозначим окрестность точки $\vec{u}$ дариусом $l$ \cite{vrp_shafeyev}:
 \begin{align} 
@@
 
-When $l=\eta$ a set $ N_l(\vec{u})$ for any vector $\vec{u}$ coincides with the set $E^{\eta}$ and being in this neighborhood vector with a minimum value of the objective function is equivalent to solving the original problem.
-The standard algorithm local search algorithm starts with a randomly selected vector $\vec{u^0}$.
-On the $i$ step of the algorithm move from the current vector in the neighboring that the minimum value of the objective function in the neighborhood of the given vector:
+При $ л = \ ETA $ множество $ N_l (\ VEC {и}) $ для любого вектора $ \ VEC {и} $ совпадает с множеством $ E ^ {\ эты} $ и быть в этой окрестности вектора с минимальным значение целевой функции эквивалентна решению исходной задачи.
+Стандартный алгоритм Локальный алгоритм поиска начинается с выбранного случайным образом вектор $ \ VEC {и ^ 0} $.
+На $ I $ шаг алгоритма перехода от текущего вектора в соседней, что минимальное значение целевой функции в окрестности данного вектора:
+
+Когда $l=\eta$, тогда множество $ N_l(\vec{u})$ для любого вектора $\vec{u}$ совпадает с множеством $E^{\eta}$  и поиск в этой окрестности минимума эквивалентно решению исходной задачи.
+Классический алгоритм локального поиска начинает поиск оптимума с  инициализации вектора $\vec{u^0}$.
+На $i$ шаге алгоритма  алгоритма выполняется перемещение текущего вектора в минимум его окрестности:
 \begin{align} 
@@
 
-The algorithm terminates at a local optimum, when $F(\vec{x^{i+1}})=F(\vec{x^i})$.
-When solving the transportation problem is a typical situation when many local optimums and only one of them is global:
+Алгоритм завершается в локальном оптимуме, когда $F(\vec{x^{i+1}})=F(\vec{x^i})$.
+В задачах маршрутизации, как правило, имеется большое множество локальных оптимумов и только один из них носит глобальный характер:
 \begin{align} 
@@
 
-\subsection{Initialization}
-To use a search algorithm is restricted by the need to build the initial solution.
-To obtain the initial solution $\vec{u^0}$ you can use a heuristic method to construct the route.
-The essence of these methods is the following.
-Every car has consistently built a route by adding another is not considered clients according to specified rules.
+\subsection{Ініціалізація}
+Для того, щоб використовувати алгоритм імітації відпалу необхідно побудувати початкове рішення.
+Для отримання початкового рішення $\vec{u^0}$ можна скористатися евристичним методом конструювання маршруту.
+Суть цих методів полягає в наступному.
+Для кожного автомобіля послідовно будується маршрут шляхом додавання ще не розглянутих клієнтських заявок за заданими правилами.
+У нашому випадку, для збереження порядку відвідування вершин, під замовленням розуміється пара "відправник-одержувач".
 
-The most known method of construction of the route proposed by Solomon in 1987 year and found in the literature as a method of Solomon.
-This algorithm is fast enough (computational complexity $O(n^3)$, therefore, it is useful to initialize a vector $\vec{u^0}$.
+Найвідоміший метод конструювання маршруту був запропонований Соломоном у 1987 році \cite{Cordeau2001,Braysy} і в літературі зустрічається як метод Соломона.
+Даний алгоритм є досить швидким (обчислювальна складність $O(n^3)$ \cite{Babb}), тому його зручно використовувати для ініціалізації вектора $\vec{u^0}$.
 
-At the beginning of the algorithm, you add one to a customer on the route according to time limits.
-The choice of the first client is randomly or chosen one, you want to serve first.
-Then, each valid client $u$ from the set of outstanding clients $Q^{'}$ (the set $Q^{'}\subset{Q}$) evaluated insertion between two adjacent nodes route.
-For selection of inserts used criterion:
+На початку роботи алгоритму відбувається додавання по одному запиту на маршрут у відповідності з тимчасовими обмеженнями.
+Вибір першої заявки відбувається випадковим чином, або вибирається та, яку необхідно виконати раніше за інших.
+Далі кожна можлива заявка $u$ з множини нерозглянутих заявок $Q$ (множина $Q' \in Q$) оцінюється на предмет вставити її в початок маршруту, кінець або між двома сусідніми вузлами маршруту.
+Для вибору місця вставки використовується наступний критерій:
 \begin{align} 
-& c(v_i,u,v_i+1)=[\min_{p=2,\dots,n^k}(\Omega_{p-1,u}^k+\Omega_{u,p}^k-\mu \Omega_{p-1,p}^k) ]
+	\label{insert_kr}
+	c(k,v_i,u,v_{i+1}) = \min =
+	\begin{cases}
+		\min_{q=2,\dots,n^k}(\Omega_{q-1,u}^k+\Omega_{u}^{k}+\Omega_{u,q}^{k}-\mu \Omega_{q-1,q}^k),\\
+		\Omega_{k,u}^k+\Omega_{u,q}^k+\Omega_{u}^k\mu \Omega_{q,q}^k), q=1\\
+		\Omega_{q,u}^k+\Omega_{u}^k, q=n^k
+	\end{cases}
 \end{align} 
-where:
+де:\\
+$k$ -- маршрут транспортного засобу $k\in C$;\\
+$n^k$ - кількість вузлів у маршруті;\\
+$\mu$ -- довільний параметр, $\mu \ge 0 $;\\
+$\Omega_{u}^k=\Omega^k[\vec{P}_{s_q}, \vec{P}_{r_u}]$ -- варіант вставки нової заявки між вже існуючими відправником і одержувачем;\\
+$\Omega_{u,q}^k=\Omega^k[\vec{P}_{r_u}, \vec{P}_{q}]$ -- варіант вставки нової заявки на початок маршруту;\\
+$\Omega_{q,u}^k=\Omega^k[\vec{P}_{q}, \vec{P}_{s_u}]$ -- варіант вставки нової заявки в кінець маршруту.
 
-$k$ - vehicle route $k\in C$;
-$n^k$ - the number of nodes in the route;
-$\mu$ - arbitrary parameter, $\mu \ge 0 $.
+Додається та заявка, у якій $c(k,v_i,u,v_{i+1})$ буде мінімальною:
+\begin{align} 
+	\label{kriterij}
+	& u^*:c(k,v_i,u^*,v_{i+1})=\min_{u}[c(k,v_i,u,v_{i+1})]
+\end{align} 
+В результаті $u^*$ додається до поточного маршруту $k$ на найвигіднішу позицію.
+\begin{figure}[h]
+	\center{\includegraphics[width=1\linewidth]{images/pict1}}
+	\caption{Приклад виконання одного кроку методом Соломона}
+	\label{ris:pict1}
+\end{figure}
 
-Added the client, which is $c(v_i,u,v_i+1)$ least:
-\begin{align} 
-& u^*:c(v_i,u^*,v_i+1)=min[c(v_i,u,V_i+1)]
-\end{align}
+На рис. \ref{ris:pict1} представлений зважений граф, на прикладі якого показано виконання одного кроку методом Соломона:\\
+а) вихідний граф, на якому показані варіанти додавання нових заявок на маршрут;\\
+b) маршрут, який був побудований методом Соломона після виконання одного кроку.
 
-As a result the client $u^*$ is added to the current route $k$ the position between nodes $v_i$ and $v_i+1$.
+\textit{Приклад.} Спочатку дана заявка (яка складається з відправника і одержувача).
+Значення функції вартості $\Omega$ представлені на дугах графа (рис.\ref{ris:pict1}).
+Необхідно додати нові заявки в маршрут, використовуючи критерії \ref{insert_kr} і \ref{kriterij}.
+Коефіцієнт $\mu=0.5$.
 
-\subsection{A Method of Forming Solutions Neighborhood}
-The most difficult part of the local search algorithm with the prohibitions are viewing the surroundings $N_l(\vec{u^i},\vec{\phi^i})$ of the current point $\vec{us^i}$.
-For even larger values of $l$ and the parameter $p=1$ in the outskirts plenty of points.
-But due to the nature of the model's constraints many of them are not the solution.
-Therefore, due to the nature of the considered discrete tasks, we used the following method of determining the area.
-On the $i$ step of the algorithm compiled from routes pair: $\{R_1,R_2\},\{R_1,R_3\},\dots,\{R_1,R_n\},\{R_2,R_1\}, \{R_2,R_3\},\dots,\{R_2,R_n\},\dots,\{R_{n-1},R_{n}\}$.
-Then over each pair of routes is performed in moving requests.
-The resulting displacement vectors $\vec{u}$ are neighborhood $N_l(\vec{u^i},\vec{\phi^i})$.
-For moving  we used change-move  and swap-move operations.
+\textit{Рішення.} Розглянемо процес додавання нових заявок.
+З двох запропонованих заявок $(\{S_2,R_2\},\{S_3,R_3\})$ виберемо ту, яку потрібно додати в першу чергу.
+В залежності від місця куди можна додати заявку (початок, кінець, середина) виберемо потрібний критерій.
+Таким чином, знайдемо оптимальне рішення вставки нової заявки $\{S_2,R_2\}$:\\
+$\min[c(k,S_1, \{S_2, R_2\}, R_1), c(k,\{S_2, R_2\}, S_1, R_1), c(k, S_1, R_1, \{S_2, R_2\})]=\min[3+11+4-0.5 \cdot 8; 17+11+8-0.5 \cdot 20; 2+11]=\min[14;26;13]$\\
+Знайдемо оптимальне рішення вставки нової заявки $\{S_3,R_3\}$:\\ 
+$\min[c(k,S_1, \{S_3, R_3\}, R_1), c(k,\{S_3, R_3\}, S_1, R_1), c(k, S_1, R_1, \{S_3, R_3\})]=\min[6+9+12-0.5 \cdot 8; 10+9+13-0.5 \cdot 20; 5+9]=\min[23;22;14]$.\\
+Із знайдених рішень виберемо рішення з найменшим значенням $c(k,v_i,u,v_{i+1})$. У нашому прикладі це $c(k, S_1, R_1, \{S_2, R_2\}=13$.
+
+Таким чином, новий маршрут буде складатися з початкових заявок і нової заявки $\{S_2,R_2\}$, яка буде додана в кінець маршруту.
+
+\textit{Відповідь}: в маршрут $\{S_1,R_1\}$ необхідно додати заявку $\{S_2,R_2\}$, яка буде розміщена в кінці.
+
+На рис. \ref{ris:pict2} представлені варіанти додавання в маршрут нової заявки $\{S_2,R_2\}$:\\
+a) варіант додавання заявки $\{S_2,R_2\}$ в кінець маршруту;\\
+b) варіант додавання заявки $\{S_2,R_2\}$ в початок маршруту;\\
+c) варіант додавання заявки $\{S_2,R_2\}$ між вершинами $\{S_1,R_1\}$.
+
+На рис. \ref{ris:pict3} представлені варіанти додавання в маршрут нової заявки $\{S_3,R_3\}$:\\
+a) варіант додавання заявки $\{S_3,R_3\}$ в кінець маршруту;\\
+b) варіант додавання заявки $\{S_3,R_3\}$ в початок маршруту;\\
+c) варіант додавання заявки $\{S_3,R_3\}$ між вершинами $\{S_1,R_1\}$.
+
+\begin{figure}[t]
+	\center{\includegraphics[width=1\linewidth]{images/pict2}}
+	\caption{Варіанти додавання в маршрут заявки $\{S_2,R_2\}$}
+	\label{ris:pict2}
+\end{figure}
+
+\begin{figure}[t]
+	\center{\includegraphics[width=1\linewidth]{images/pict3}}
+	\caption{Варіанти додавання в маршрут заявки $\{S_3,R_3\}$}
+	\label{ris:pict3}
+\end{figure}
+
+\clearpage
+
+\subsection{Метод формування околиці}
+Під час роботи алгоритму імітації відпалу необхідно проводити перегляд околиці $N_{l}(\vec{u^i})$ поточної точки $\vec{u^i}$.В ході обчислення маршруту в околицю потрапляє досить велика кількість точок.
+Але в зв'язку з особливостями обмежень нашої моделі багато з цих точок не є рішенням задачі.
+Тому в зв'язку з особливостями розглянутої дискретної задачі ми використали наступний прийом визначення околиці.
+
+Нехай $x^{k}(v)$ -- індекс вузла $v$, $v \in R \cup S$ в маршруті автотранспортного засобу $k \in C$.
+Кожна заявка складається з пари $\{S,R\}$, де $S$ -- це відправник, а $R$ -- одержувач.
+Під час визначення околиці, повинна зберігатися цілісність маршруту, тобто виконуватися обмеження (\ref{main_cond_1} --\ref{main_cond_4}).
+Над заявками маршруту $Route_i$, $i=1 \dots n$ виконуються операції перестановки.
+Отримані під час перестановок вектори $\vec{u}$ і буде околицею $N_{l}(\vec{u^i})$.
+Для перестановки використовувалися операції переміщення і поглинання.
+
+Операція поглинання представлена на рис.\ref{ris:figures}(a), на якому представлено два маршрути $Route_1$ і $Route_2$.
+Маршрут $Route_1$ складається з вузлів $\{C_1,S_1,R_1,S_2,\\R_2,S_3,R_3\}$, а маршрут $Route_2$ складається з вузлів $\{C_2,S_4,R_4,S_5,R_5,S_6,R_6\}$.
+Під час виконання операції поглинання заявки $\{S_1,R_1\}$ в маршруті $Route_1$, вставка даної заявки в маршрут $Route_2$ може бути виконана на $n+1$ позицій.
+
+\begin{figure}[h]
+	\center{\includegraphics[width=0.75\linewidth]{images/neighboard.png}}
+	\caption{Приклад операції a)поглинання та b)переміщення}
+	\label{ris:figures}
+\end{figure}
+
+Операція переміщення представлена на рис.\ref{ris:figures}(b) На даному рисунку представлений маршрут $Route_1$, який складається з вузлів $\{C_1,S_1,R_1,S_2,R_2,S_3,R_3\}$.
+Для збереження цілісності маршруту вузол $S_i$ (відправник) має бути відвіданий до $R_i$ (одержувач), виходячи з цього під час операції переміщення, вузол $S_i$ може переміщатися поки буде виконуватися умова $0<x^{new}(S_i)<\widetilde{u}(R_i)$, а вузол $R_i$ може переміщатися поки буде виконуватися умова $n^k \geq x^{new}(R_i)>\widetilde{u}(S_i)$.
 
@@
 
+\bibitem{Attanasio} A.~Attanasio, J.~Bregman,G.~Ghiani, E.~Manni.
+Real-time fleet management at Ecourier Ltd.
+Dynamic Fleet Management, volume 38 of Operations Research/Computer Science Interfaces, chapter 10, p.219--238, 2007.
+
+\bibitem{Beaudry} A.~Beaudry, G.~Laporte, T.~Melo, Nickel. Dynamic transportation of patients in hospitals. Berichte des Fraunhofer ITWM, Nr. 104 :p.1--34, 2010.
+
+\bibitem{Romero} M.Romero, L.Sheremetov and A.Soriano.
+A genetic algorithm for the pickup and delivery problem: An application to the helicopter offshore transportation.
+In Theoretical Advances and Applications of Fuzzy Logic and Soft Computing, volume 42 of Advances in Soft Computing, 2007, p.
+35--44.
+
+\bibitem{Kergosien} M.Romero, L.Sheremetov and A.Soriano.
+Yannick Kergosien, Christophe Lent, D.
+Piton, Jean-Charles Billaut.
+A tabu search heuristic for a dynamic transportation problem of patients between care units.
+29 pages.
+2010.
+
+\bibitem{Sarak} H.~Sarak, A.~Satman. The degree-day method to estimate the residential heating natural gas consumption in Turkey: a case study. Pergamon, Energy 28 (2003) 929--939.
+
 \end{thebibliography}
 
-\end{document}
+\end{document} \grid
+\grid

### 2016-05-12 article_ru.tex
translate to rus

@@
 \section{
-``Исследование  настроечных параметров алгоритма поиска с запретами и его модификация для решения курьерской задачи маршрутизации.'' 
+``Исследование  настроечных параметров алгоритма поиска с запретами и его модификация для решения курьерской задачи маршрутизации'' 
 }
@@
 Данная статья посвящена рассмотрению задачи курьерской маршрутизации, которая является подзадачей ЗМТ.
-Данная задача состоит в нахождении маршрутов для посещения заданного множества адресов некоторым количеством единиц транспортных средств.
-Такая проблема наиболее часто встречается среди компаний, выполняющих развозки товара от отправителя к получателю.
+Данная задача состоит в нахождении маршрутов для посещения заданного множества адресов некоторым количеством единиц транспортных средств, выполняющих развозки товара от отправителя к получателю.
 
@@
 
-В последнее время все больше людей пользуются услугами курьерских служб.  Обычно заявки формируются днем ранее и
-перед диспетчером полный список заявок, которые необходимо выполнить в кратчайшие сроки.
-Тематическое исследование провел Attanasio(2007) \cite{Attanasio}, в котором описал преимущества использования компьютерных методов перед человеческой диспетчеризацией.
+В последнее время все больше людей пользуются услугами курьерских служб.
+Обычно заявки формируются днем ранее и у диспетчера имеется полный список заявок, которые необходимо выполнить в кратчайшие сроки.
+Attanasio(2007) \cite{Attanasio} провел тематическое исследование, в котором описал преимущества использования компьютерных методов перед человеческой диспетчеризацией.
 В своей работе он рассмотрел примеры работы eCourier Ltd, лондонской компании, которая предлагает курьерские услуги.
 Их клиентами в основном являются юридические фирмы, финансовые учреждения, рекламные агентства и другие организации, которые заинтересованы в быстрой доставке товаров или оригинальных подписях на документах.
-После того как все заявки на конкретный момент времени сформированы, необходимо составить такой маршрут, который удовлетворял бы требованиям заказчиков.
-В зависимости от уровня обслуживания, который был задан клиентом, курьер может объединить или ни его доставку с другими, то есть это означает, что товары могут быть доставлены по очереди, или курьер может выполнить ускоренную доставку.
+После того, как все заявки на конкретный момент времени сформированы, необходимо составить такой маршрут, который удовлетворял бы требованиям заказчиков.
+В зависимости от уровня обслуживания, который был задан клиентом, курьер может объединить товары нескольких клиентов в процессе развоза.
 
 Компании, которые предлагают курьерские услуги достаточно часто имеют смешанный парк транспортных средств, который состоит из велосипедов, мотоциклов, автомобилей и небольших грузовиков.
-В зависимости от типа запроса, расположение и временных окон будет выбран тот вид транспортного средства, который сможет выполнить заказ в кратчайшие сроки.
+В зависимости от типа запроса, расположения транспортных средств и временных окон будет выбран тот вид транспортного средства, который сможет выполнить заказ в кратчайшие сроки.
 Исследования \cite{Attanasio} показали, что использование компьютерных методов, в том числе алгоритмов оптимизации, были очень выгодными для курьерских компаний.
@@
     \begin{cases}
-	  1,&\text{АТС $k \in C$ движется с узла $i$ в вершину $j$}\\
+	  1,&\text{АТС $k \in C$ движется с узла $i$ в узел $j$}\\
 	  0,&\text{в противном случае.}
@@
 
-Обозначим через $\delta(\vec{u},\vec{v})$ расстояние Хэмминга между $\vec{u}$ и $\vec{v}$. Через $N_l(\vec{u})$ обозначим окрестность точки $\vec{u}$ дариусом $l$ \cite{vrp_shafeyev}:
+Обозначим через $\delta(\vec{u},\vec{v})$ расстояние Хэмминга между $\vec{u}$ и $\vec{v}$. Через $N_l(\vec{u})$ обозначим окрестность точки $\vec{u}$ радиусом $l$ \cite{vrp_shafeyev}:
 \begin{align} 
@@
 
-При $ л = \ ETA $ множество $ N_l (\ VEC {и}) $ для любого вектора $ \ VEC {и} $ совпадает с множеством $ E ^ {\ эты} $ и быть в этой окрестности вектора с минимальным значение целевой функции эквивалентна решению исходной задачи.
-Стандартный алгоритм Локальный алгоритм поиска начинается с выбранного случайным образом вектор $ \ VEC {и ^ 0} $.
-На $ I $ шаг алгоритма перехода от текущего вектора в соседней, что минимальное значение целевой функции в окрестности данного вектора:
+При $l=\eta$ множество $N_l(\vec{u})$ для любого вектора $ \vec{u} $ совпадает с множеством $E^{\eta}$ и поиск в этой окрестности вектора с минимальным значение целевой функции эквивалентно решению исходной задачи.
+Стандартный алгоритм локального поиска начинается с выбранного случайным образом вектора $\vec{u^0}$.
+На $i$ шаге алгоритма выполняется переход от текущего вектора в минимум окрестности:
 
@@
 
-To ensure that the algorithm did not stop in a local minimum, and passed from one local minimum to another, with edges removed the Central point and when searching for the minimum of the following rule applies.
-Let $l=2$ and in the transition from $\vec{u^{i+1}}$ to $\vec{u^i}$ change the values in the coordinate $(u_\lambda^{i}, u_\omega^{i})$.
-The algorithm stores such pair for the last $h$ couple of steps and in the next step prohibits the movement in these directions.
-An ordered list of such pairs
+Для того, чтобы убедиться, что алгоритм не остановится в локальном минимуме, и перемещается от одного локального минимума в другой, с окрестности удаляется центральная точка и при поиске минимума окрестности применяется следующее правило.
+Пусть $l=2$ и при переходе от $\vec{u^{i+1}}$ к $\vec{u^i}$ изменяются значения в координате $(u_\lambda^{i}, u_\omega^{i})$.
+Алгоритм хранит такие пары за последние $h$ шагов и на следующем этапе запрещает движение в этих направлениях.
+Упорядоченный список таких пар:
 \begin{align} 
@@
 \end{align}	
-called the list of prohibitions.
-When building the list of all pairs of different and a pair of $(u_\lambda,u_\omega),\lambda \ne \omega$ does not prohibit the movement of pairs $(u_\lambda,u_\lambda)$ and $(u_\omega,u_\omega)$.
-When $l>2$ similarly constructed three coordinates, fours, etc.
-A set of non-restricted vectors denote by $N_l(\vec{u^i},\vec{\phi^i})$.
-In order that the search was effectively, it is advisable to use small values $h$ and to control this parameter in the course of the algorithm.
+является списком запретом.
+При создании списка всех пар  и пару $(u_\lambda,u_\omega),\lambda \ne \omega$ не запрещает движение пар $(u_\lambda,u_\lambda)$ and $(u_\omega,u_\omega)$.
+Когда $l>2$ аналогичным образом построены три координаты, четыре и т.д.
+Набор неограниченная векторов обозначим через $N_l(\vec{u^i},\vec{\phi^i})$.
+Для того чтобы поиск был эффективным, то целесообразно использовать малые значения  $h$ и контролировать этот параметр в ходе алгоритма.
 
@@
 
-Denote by $N_l(\vec{u^i},\vec{\phi^i},p)$ probabilistic edge that goes from a deterministic $N_l(\vec{u^i},\vec{\phi^i})$ as follows.
-Each vector $\vec{v} \in N_l(\vec{u^i},\vec{\phi^i})$ with probability $p$ included in the neighborhood $N_l(\vec{u^i},\vec{\phi^i},p)$ regardless of other points.
-Note that this set may be empty or contain only one point.
-The General scheme of the probabilistic search algorithm with a list of prohibitions referred to as pseudo-code in algorithm \\
-scheme \ref{alg:TabuSearch}.
+Обозначим через $N_l(\vec{u^i},\vec{\phi^i},p)$ вероятностный край, который идет от детерминированной $N_l(\vec{u^i},\vec{\phi^i})$ следующим образом.
+Каждый вектор $\vec{v} \in N_l(\vec{u^i},\vec{\phi^i})$ с вероятностью $p$ включенных в окрестности $N_l(\vec{u^i},\vec{\phi^i},p)$ независимо от других точек.
+Обратите внимание, что этот набор может быть пустым или содержать только одну точку.
+Общая схема вероятностного алгоритма поиска со списком запретов упоминается как псевдо-кода в алгоритме \\
+представлена на схеме \ref{alg:TabuSearch}.
 
-As the stopping criterion is based on the total number of steps $N_stop$, in the course of which does not change the value $F_{opt}$.
-Values $l,p,h$ are the control parameters of the algorithm.
-Their choice depends on the problem dimension.
+По мере того как критерий останова основан на общем количестве шагов  $N_stop$,  в ходе которого не изменяет значение $F_{opt}$.
+Значения $l,p,h$ являются настроечными параметрами алгоритма.
+Их выбор зависит от размерности задачи.
 
-In the presented scheme, it was assumed that the value of $h$ (the dimension of the deny list) does not change in the course of the algorithm.
-This creates certain difficulties in the implementation of the scheme, as it is unknown how long is to take a list of prohibitions.
-At small $h$ the algorithm can start walking cycle.
-At large $h$ the search becomes inefficient.
-One of the simple rules regulating the length of the list of prohibitions is the following.
-
-If at the next step of the algorithm, the edge  $N_l(\vec{u^i},\vec{\phi^i},p)$ turned out to be empty, it is added to the list of fictitious vector $\vec{0}$.
-This vector nothing forbids, but reduces the number of bans per unit.
+В представленной схеме, предполагалось, что величина $h$ (размерность списка запретов)  не изменяется в ходе алгоритма.
+Это создает определенные трудности в реализации схемы, поскольку неизвестно, каким должен быть размер списка запретов.
+При малых $h$ алгоритм может зациклиться.
+При больших  $h$ поиск становится неэффективным.
 
 \subsection{Ініціалізація}
-Для того, щоб використовувати алгоритм імітації відпалу необхідно побудувати початкове рішення.
-Для отримання початкового рішення $\vec{u^0}$ можна скористатися евристичним методом конструювання маршруту.
-Суть цих методів полягає в наступному.
-Для кожного автомобіля послідовно будується маршрут шляхом додавання ще не розглянутих клієнтських заявок за заданими правилами.
-У нашому випадку, для збереження порядку відвідування вершин, під замовленням розуміється пара "відправник-одержувач".
+Для того, чтобы использовать алгоритм поиска с запретами, необходимо построить первоначальное решение.
+Для получения первоначального решения $\vec{u^0}$ можно воспользоваться эвристическим методом конструирования маршрута.
+Суть этих методов состоит в следующем.
+Для каждого автомобиля последовательно строится маршрут путем добавления еще не рассмотренных клиентских заявок по заданным правилам.
+В нашем случае, для сохранения порядка посещения вершин, под заказом понимается пара "отправитель-получатель".
 
-Найвідоміший метод конструювання маршруту був запропонований Соломоном у 1987 році \cite{Cordeau2001,Braysy} і в літературі зустрічається як метод Соломона.
-Даний алгоритм є досить швидким (обчислювальна складність $O(n^3)$ \cite{Babb}), тому його зручно використовувати для ініціалізації вектора $\vec{u^0}$.
+Самый известный метод конструирования маршрута был предложен Соломоном в 1987 году \cite{Cordeau2001,Braysy} и в литературе встречается как метод Соломона.Данный алгоритм является достаточно быстрым (вычислительная сложность $O(n^3)$ \cite{Babb}), поэтому его удобно использовать для инициализации вектора $\vec{u^0}$.
 
-На початку роботи алгоритму відбувається додавання по одному запиту на маршрут у відповідності з тимчасовими обмеженнями.
-Вибір першої заявки відбувається випадковим чином, або вибирається та, яку необхідно виконати раніше за інших.
-Далі кожна можлива заявка $u$ з множини нерозглянутих заявок $Q$ (множина $Q' \in Q$) оцінюється на предмет вставити її в початок маршруту, кінець або між двома сусідніми вузлами маршруту.
-Для вибору місця вставки використовується наступний критерій:
+В начале работы алгоритма происходит добавление по одному запросу на маршрут в соответствии с временными ограничениями.
+Выбор первой заявки происходит случайным образом, или выбирается та, которую необходимо выполнить раньше других.
+Далее каждая возможная заявка  $u$ из множества нерассмотренных заявок $Q$ (множество $Q' \in Q$) оценивается на предмет вставить ее в начало маршрута, конец или между двумя соседними узлами маршрута.
+Для выбора места вставки используется следующий критерий:
 \begin{align} 
@@
 \end{align} 
-де:\\
-$k$ -- маршрут транспортного засобу $k\in C$;\\
-$n^k$ - кількість вузлів у маршруті;\\
-$\mu$ -- довільний параметр, $\mu \ge 0 $;\\
-$\Omega_{u}^k=\Omega^k[\vec{P}_{s_q}, \vec{P}_{r_u}]$ -- варіант вставки нової заявки між вже існуючими відправником і одержувачем;\\
-$\Omega_{u,q}^k=\Omega^k[\vec{P}_{r_u}, \vec{P}_{q}]$ -- варіант вставки нової заявки на початок маршруту;\\
-$\Omega_{q,u}^k=\Omega^k[\vec{P}_{q}, \vec{P}_{s_u}]$ -- варіант вставки нової заявки в кінець маршруту.
+где:\\
+$k$ -- маршрут транспортного средства $k\in C$;\\
+$n^k$ - количество узлов в маршруте;\\
+$\mu$ --настроечный параметр, $\mu \ge 0 $;\\
+$\Omega_{u}^k=\Omega^k[\vec{P}_{s_q}, \vec{P}_{r_u}]$ -- вариант вставки новой заявки между уже существующими отправителем и получателем;\\
+$\Omega_{u,q}^k=\Omega^k[\vec{P}_{r_u}, \vec{P}_{q}]$ -- вариант вставки новой заявки на начало маршрута;\\
+$\Omega_{q,u}^k=\Omega^k[\vec{P}_{q}, \vec{P}_{s_u}]$ -- вариант вставки новой заявки в конец маршрута.
 
-Додається та заявка, у якій $c(k,v_i,u,v_{i+1})$ буде мінімальною:
+Добавляется та заявка, у которой $c(k,v_i,u,v_{i+1})$ будет минимальным:
 \begin{align} 
@@
 \end{align} 
-В результаті $u^*$ додається до поточного маршруту $k$ на найвигіднішу позицію.
+В результате $u^*$ добавляется к текущему маршруту $k$ на наиболее выгодную позицию.
 \begin{figure}[h]
 	\center{\includegraphics[width=1\linewidth]{images/pict1}}
-	\caption{Приклад виконання одного кроку методом Соломона}
+	\caption{Пример выполнения одного шага методом Соломона}
 	\label{ris:pict1}
@@
 
-На рис. \ref{ris:pict1} представлений зважений граф, на прикладі якого показано виконання одного кроку методом Соломона:\\
-а) вихідний граф, на якому показані варіанти додавання нових заявок на маршрут;\\
-b) маршрут, який був побудований методом Соломона після виконання одного кроку.
+На рис. \ref{ris:pict1} представлен взвешенный граф, на примере которого показано выполнение одного шага методом Соломона: \\
+а) исходный граф, на котором показаны варианты добавления новых заявок на маршрут; \\
+b) маршрут, который был построен методом Соломона после выполнения одного шага.
 
-\textit{Приклад.} Спочатку дана заявка (яка складається з відправника і одержувача).
-Значення функції вартості $\Omega$ представлені на дугах графа (рис.\ref{ris:pict1}).
-Необхідно додати нові заявки в маршрут, використовуючи критерії \ref{insert_kr} і \ref{kriterij}.
-Коефіцієнт $\mu=0.5$.
+\textit{Пример.} Изначально данна заявка (которая состоит из отправителя и получателя).
+Значение функции стоимости $\Omega$ представлены на дугах графа (рис.\ref{ris:pict1}).
+Необходимо добавить новые заявки в маршрут, используя критерии \ref{insert_kr} і \ref{kriterij}.
+Коэффициент $\mu=0.5$.
 
-\textit{Рішення.} Розглянемо процес додавання нових заявок.
-З двох запропонованих заявок $(\{S_2,R_2\},\{S_3,R_3\})$ виберемо ту, яку потрібно додати в першу чергу.
-В залежності від місця куди можна додати заявку (початок, кінець, середина) виберемо потрібний критерій.
-Таким чином, знайдемо оптимальне рішення вставки нової заявки $\{S_2,R_2\}$:\\
+\textit{Решение.}Рассмотрим процесс добавления новых заявок.
+Из двух предложенных заявок $(\{S_2,R_2\},\{S_3,R_3\})$ выберем ту, которую нужно добавить в первую очередь.
+В зависимости от места куда можно добавить заявку (начало, конец, середина) выберем нужный критерий.
+Таким образом, найдем оптимальное решение вставки новой заявки $\{S_2,R_2\}$:\\
 $\min[c(k,S_1, \{S_2, R_2\}, R_1), c(k,\{S_2, R_2\}, S_1, R_1), c(k, S_1, R_1, \{S_2, R_2\})]=\min[3+11+4-0.5 \cdot 8; 17+11+8-0.5 \cdot 20; 2+11]=\min[14;26;13]$\\
-Знайдемо оптимальне рішення вставки нової заявки $\{S_3,R_3\}$:\\ 
+Найдем оптимальное решение вставки новой заявки $\{S_3,R_3\}$:\\ 
 $\min[c(k,S_1, \{S_3, R_3\}, R_1), c(k,\{S_3, R_3\}, S_1, R_1), c(k, S_1, R_1, \{S_3, R_3\})]=\min[6+9+12-0.5 \cdot 8; 10+9+13-0.5 \cdot 20; 5+9]=\min[23;22;14]$.\\
-Із знайдених рішень виберемо рішення з найменшим значенням $c(k,v_i,u,v_{i+1})$. У нашому прикладі це $c(k, S_1, R_1, \{S_2, R_2\}=13$.
+Из найденных решений выберем решение с наименьшим значением $c(k,v_i,u,v_{i+1})$. В нашем примере это $c(k, S_1, R_1, \{S_2, R_2\}=13$.
 
-Таким чином, новий маршрут буде складатися з початкових заявок і нової заявки $\{S_2,R_2\}$, яка буде додана в кінець маршруту.
+Таким образом, новый маршрут будет состоять из начальных заявок и новой заявки $\{S_2,R_2\}$,которая будет добавлена в конец маршрута.
 
-\textit{Відповідь}: в маршрут $\{S_1,R_1\}$ необхідно додати заявку $\{S_2,R_2\}$, яка буде розміщена в кінці.
+\textit{Ответ}: в маршрут $\{S_1,R_1\}$ необходимо подать заявку $\{S_2,R_2\}$, которая будет размещена в конце.
 
-На рис. \ref{ris:pict2} представлені варіанти додавання в маршрут нової заявки $\{S_2,R_2\}$:\\
-a) варіант додавання заявки $\{S_2,R_2\}$ в кінець маршруту;\\
-b) варіант додавання заявки $\{S_2,R_2\}$ в початок маршруту;\\
-c) варіант додавання заявки $\{S_2,R_2\}$ між вершинами $\{S_1,R_1\}$.
+На рис. \ref{ris:pict2} представлены варианты добавления в маршрут новой заявки $\{S_2,R_2\}$:\\
+a) вариант добавления заявки $\{S_2,R_2\}$ в конец маршрута;\\
+b) вариант добавления заявки $\{S_2,R_2\}$ в начало маршрута;\\
+c) вариант добавления заявки $\{S_2,R_2\}$ между узлами $\{S_1,R_1\}$.
 
-На рис. \ref{ris:pict3} представлені варіанти додавання в маршрут нової заявки $\{S_3,R_3\}$:\\
-a) варіант додавання заявки $\{S_3,R_3\}$ в кінець маршруту;\\
-b) варіант додавання заявки $\{S_3,R_3\}$ в початок маршруту;\\
-c) варіант додавання заявки $\{S_3,R_3\}$ між вершинами $\{S_1,R_1\}$.
+На рис. \ref{ris:pict3} представлены варианты добавления в маршрут новой заявки $\{S_3,R_3\}$:\\
+a) вариант добавления заявки $\{S_3,R_3\}$ в кінець маршруту;\\
+b) вариант добавления заявки $\{S_3,R_3\}$ в початок маршруту;\\
+c) вариант добавления заявки $\{S_3,R_3\}$ между узлами $\{S_1,R_1\}$.
 
@@
 	\center{\includegraphics[width=1\linewidth]{images/pict2}}
-	\caption{Варіанти додавання в маршрут заявки $\{S_2,R_2\}$}
+	\caption{Варианты добавления в маршрут заявки $\{S_2,R_2\}$}
 	\label{ris:pict2}
@@
 	\center{\includegraphics[width=1\linewidth]{images/pict3}}
-	\caption{Варіанти додавання в маршрут заявки $\{S_3,R_3\}$}
+	\caption{Варианты добавления в маршрут заявки $\{S_3,R_3\}$}
 	\label{ris:pict3}
@@
 
-\subsection{Метод формування околиці}
-Під час роботи алгоритму імітації відпалу необхідно проводити перегляд околиці $N_{l}(\vec{u^i})$ поточної точки $\vec{u^i}$.В ході обчислення маршруту в околицю потрапляє досить велика кількість точок.
-Але в зв'язку з особливостями обмежень нашої моделі багато з цих точок не є рішенням задачі.
-Тому в зв'язку з особливостями розглянутої дискретної задачі ми використали наступний прийом визначення околиці.
+\subsection{Метод формирования окрестности}
+Во время работы алгоритма имитации отжига необходимо проводить просмотр окрестности $N_{l}(\vec{u^i})$ текущей точки $\vec{u^i}$.
+В ходе вычисления маршрута в окрестность попадает достаточно большое количество точек.
+Но в связи с особенностями ограничений нашей модели многие из этих точек не является решением задачи.
+Поэтому в связи с особенностями рассматриваемой дискретной задачи мы использовали следующий прием определения окрестности.
 
-Нехай $x^{k}(v)$ -- індекс вузла $v$, $v \in R \cup S$ в маршруті автотранспортного засобу $k \in C$.
-Кожна заявка складається з пари $\{S,R\}$, де $S$ -- це відправник, а $R$ -- одержувач.
-Під час визначення околиці, повинна зберігатися цілісність маршруту, тобто виконуватися обмеження (\ref{main_cond_1} --\ref{main_cond_4}).
-Над заявками маршруту $Route_i$, $i=1 \dots n$ виконуються операції перестановки.
-Отримані під час перестановок вектори $\vec{u}$ і буде околицею $N_{l}(\vec{u^i})$.
-Для перестановки використовувалися операції переміщення і поглинання.
+Пусть $x^{k}(v)$ -- индекс узла $v$, $v \in R \cup S$ в маршруте автотранспортного средства $k \in C$.
+Каждая заявка состоит из пары $\{S,R\}$, где $S$ -- отправитель, а $R$ -- получатель.
+В процессе формирования окрестности, должна сохраняться целостность маршрута, то есть выполняться ограничения (\ref{main_cond_1} --\ref{main_cond_4}).
+Над заявками маршрута $Route_i$, $i=1 \dots n$выполняются операции перестановки.
+Полученные в ходе перестановок векторы $\vec{u}$ и будет окрестностью $N_{l}(\vec{u^i})$.
+Для перестановки использовались операции перемещения и поглощения.
 
-Операція поглинання представлена на рис.\ref{ris:figures}(a), на якому представлено два маршрути $Route_1$ і $Route_2$.
-Маршрут $Route_1$ складається з вузлів $\{C_1,S_1,R_1,S_2,\\R_2,S_3,R_3\}$, а маршрут $Route_2$ складається з вузлів $\{C_2,S_4,R_4,S_5,R_5,S_6,R_6\}$.
-Під час виконання операції поглинання заявки $\{S_1,R_1\}$ в маршруті $Route_1$, вставка даної заявки в маршрут $Route_2$ може бути виконана на $n+1$ позицій.
+Операция поглащения представлена на рис.\ref{ris:figures}(a), на котором представлены два маршрута $Route_1$ і $Route_2$.
+Маршрут $Route_1$ состоит из улов $\{C_1,S_1,R_1,S_2,\\R_2,S_3,R_3\}$, а маршрут $Route_2$ состоит из узлов $\{C_2,S_4,R_4,S_5,R_5,S_6,R_6\}$.
+Во время выполнения операции поглощения $\{S_1,R_1\}$ в маршруте $Route_1$, вставка данной заявки в маршрут $Route_2$ може бути виполнена на $n+1$ позиций.
 
@@
 	\center{\includegraphics[width=0.75\linewidth]{images/neighboard.png}}
-	\caption{Приклад операції a)поглинання та b)переміщення}
+	\caption{Пример операции a)поглащения и b)перемещения}
 	\label{ris:figures}
@@
 
-Операція переміщення представлена на рис.\ref{ris:figures}(b) На даному рисунку представлений маршрут $Route_1$, який складається з вузлів $\{C_1,S_1,R_1,S_2,R_2,S_3,R_3\}$.
-Для збереження цілісності маршруту вузол $S_i$ (відправник) має бути відвіданий до $R_i$ (одержувач), виходячи з цього під час операції переміщення, вузол $S_i$ може переміщатися поки буде виконуватися умова $0<x^{new}(S_i)<\widetilde{u}(R_i)$, а вузол $R_i$ може переміщатися поки буде виконуватися умова $n^k \geq x^{new}(R_i)>\widetilde{u}(S_i)$.
+Операция перемещения представлена на рис.\ref{ris:figures}(b).
+На данном рисунке представлен маршрут $Route_1$, который состоит из узлов $\{C_1,S_1,R_1,S_2,R_2,S_3,R_3\}$.
+Для сохранения целостности маршрута: узел $S_i$ (отправитель) может быть посещен только до $R_i$ (получатель),исходя из этого во время операции перемещения, узел $S_i$ может перемещаться пока будет выполняться условие $0<x^{new}(S_i)<\widetilde{u}(R_i)$,а узел $R_i$ может перемещаться пока будет выполняться условие $n^k \geq x^{new}(R_i)>\widetilde{u}(S_i)$.
 
-\subsection{Analysis of the algorithm parameters}
-The Tabu Search algorithm has two main parameters: $h$ -- size of the list of prohibitions and $N_{neighbors}$ -- accepted count limit of size of neighbordhood $N_l(\vec{u})$.
-For the experiments, we constructed the test tasks on the basis of the capacity routing problems developed by Breedam, Fisher, Christofides and Eilon.
-To determine trends in time and cost of the decision we used exponential smoothing by a factor of $0.1$.
+\subsection{Анализ параметров алгоритма}
+The Tabu Search алгоритм имеет два основных настроечных параметра: $h$ -- размер списка запретов и $N_{neighbors}$ --Предел допускаемого подсчет размера окрестности $N_l(\vec{u})$.
+Для экспериментов, мы построили тестовые задания на основе проблем маршрутизации мощности, разработанных Breedam, Fisher, Christofides and Eilon.
+Для определения тенденций во времени и стоимости решения мы используем экспоненциальное сглаживание с коэффициентом $0.1$.
 
@@
 
-In this experiment we use two different breakpoint strategies: 
+В этом эксперименте мы используем две различные точки останова стратегии:
 \begin{enumerate}
-	\item $StepCount$ termination strategy --  terminates when an amount of steps has been reached;
-	\item $TimeSpent$ termination strategy -- terminates when an amount of time has been reached.
+	\item $StepCount$ Стратегия завершения - заканчивается, когда количество шагов было достигнуто;
+	\item $TimeSpent$ Стратегия завершения - заканчивается, когда количество времени было достигнуто.
 \end{enumerate}
 
-As should have been expected, for large values of the parameter a jam occurs in local optimum due to the large number of restrictions of movement in space.
-Otherwise if you select a too small tabu size, algorithm can still get stuck in a local optimum.
-In the first computational experiment we made finding the best solutions for different values of the tabu size parameter from interval $h \in [0, 40]$ (pict.
-\ref{aba:fig2} an pict.
+Как и следовало ожидать, при больших значениях параметра застревание происходит в локальный оптимум из-за большого количества ограничений движения в космосе.
+В противном случае, если вы выбрали слишком маленький размер Табу алгоритм может все еще застревают в локальный оптимум.
+В первом вычислительного эксперимента мы сделали находить оптимальные решения для различных значений параметра размера Табу из интервала $h \in [0, 40]$ (рис.
+\ref{aba:fig2} рис.
 \ref{aba:fig3}).
@@
 	{
-	  The dependence of the quality of solutions on the size of the tabu list when using $ TimeSpent $ termination strategy
+	Зависимость качества решений от размера списка при использовании табу $ TimeSpent $ стратегии завершения
 	}
@@
 	{
-	  The dependence of the quality of solutions on the neighborhood size when using $ TimeSpent $ termination strategy
+	 Зависимость качества решений от размера окрестности при использовании $ TimeSpent $ стратегии завершения
 	}

### 2016-05-13 article_ru.tex
translate to russian

@@
 
-In the second experiment (see figure \ref{aba:fig4}) we built a relationship between the number $N_{neigh.}$ of viewed solutions neighborhood size $N_l(\vec{u})$ using $TimeSpent$ termination strategy.
-In this experiment we made finding the best solutions for different values of the neighborhood size parameter from interval $N_{neigh.}\in[750, 850]$.
+Во втором эксперименте (смотри рисунок  \ref{aba:fig4}) мы построили зависимость между параметром $N_{neigh.}$ просматриваемой размера решения окрестности $N_l(\vec{u})$ используя $TimeSpent$ стратегии завершения.
+В этом эксперименте мы сделали находить оптимальные решения для различных значений параметра размера окрестности от интервала $N_{neigh.}\in[750, 850]$.
 
 \subsection{Modification}
+Первая обобщенная схема выглядит следующим образом.
+Во-первых, множество решений строится Соломона.
+Тогда каждое решение улучшается поиск с запретами и выбрать оптимальное решение в соответствии с целевой функцией.
 
-The first generalized diagram is as follows.
-First, the set of solutions is constructed by Solomon.
-Then every solution is improved by tabu search and choose the best solution according to the objective function.
+Модификация схемы основана на гипотезе «О большой долине" \cite{opt_Kochetov}.
+Согласно этой гипотезе, в среднем локальных оптимумов расположены гораздо ближе к глобальным, чем случайно выбранной точке.
+Существует определенная концентрация локального оптимума в небольшой части допустимой области, которая образно называется большая долина.
+Если это предположение верно, то желательно, чтобы вспомнить лучшие решения и на их основе разработать новое оригинальное решение.
+Мы используем эту идею для решения маршрутизации Courier проблемы доставки.
 
-Modification of the scheme is based on the hypothesis of "On a large valley"\cite{opt_Kochetov}.
-According to this hypothesis, the average local optima are located much closer to the global than a randomly chosen point.
-There is a certain concentration of local optimum in a small part of the feasible region, which is figuratively called a large valley.
-If this assumption is true, then it is advisable to remember the best solutions and based on them design new original decision.
-We use this idea to solve the  Routing Courier Delivery Problem.
-
-We proceed to the description of the algorithm in scheme \ref{alg:constructSolution}. Let $U^{opt}$ is a sorted array of optimal solutions by value of the objective function ascending,  ie:
+Перейдем к описанию алгоритма в схеме \ref{alg:constructSolution}. Пусть $U^{opt}$ является отсортированный массив оптимальных решений по значению целевой функции по возрастанию, то есть:
 \begin{equation} \label{u_sorted}
@@
 \end{equation}
-Each solution $u^{opt}_i$ gets into the population with a given probability $p^i_u$, the probability of selection decreases with increasing sequence number:
+Каждое решение $u^{opt}_i$ попадает в популяции с заданной вероятностью $p^i_u$, вероятность выбора уменьшается с увеличением порядкового номера:
 \begin{equation} \label{p_sorted}
@@
 
-It then crosses solutions by the following rule: with the first solution is randomly selected route $R_1$.
-Then another solution is selected from a route $R_2$ that does not include client requests from the first  route etc.
-If the client requests were not included in the solution $u^0$, then these requests are added by the construction heuristics algorithm (for example, Solomon algorithm).
+Затем она пересекает решения по следующему правилу: с первым решением случайным образом выбирается маршрут $R_1$.
+Тогда другое решение выбирается из маршрута $R_2$ который не включает в себя клиентские запросы от первого маршрута и т.д.
+Если клиентские запросы не были включены в решение  $u^0$, t то эти запросы добавляются в алгоритм построения эвристики ( например, Соломон алгоритм).
 
-Now we describe the primary function $SolveCDP()$ of the modified algorithm presented in pseudocode form in Scheme \ref{alg:modTabuSearch}.
-In this function using a loop is a sequential formation of an array of the best solutions $U^{opt}$ by means TabuSearch algorithm.
+Теперь мы опишем основную функцию $SolveCDP()$  модифицированного алгоритма представлено в виде псевдокода на схеме \ref{alg:modTabuSearch}.
+В этой функции с помощью цикла является последовательное формирование массива из лучших решений $U^{opt}$ с помощью алгоритма TabuSearch.
 
-Based on the results of the experiments described in the preceding section, we decided to set the parameters of the algorithm are not fixed values, but as a random variable of predetermined period.
-For example, the size of the tabu list $h$ is defined as the interval $[10,35]$(see pict.
+На основании результатов экспериментов, описанных в предыдущем параграфе, мы решили установить параметры алгоритма не являются фиксированными значениями, но, как случайной величины заданного периода.
+Например, размер списка $h$  определяется как интервал  $[10,35]$(see pict.
 \ref{aba:fig2} and  pict.
-\ref{aba:fig3}) and the neighbordhood size $l=N_{neighbordhood}$ is defined as the interval $[600\cdot n,850\cdot n]$, where $n$ -- problem size (see pict.
-\ref{aba:fig4} ).
+\ref{aba:fig3}) и размер окрестности $l=N_{neighbordhood}$ определяется как интервал $[600\cdot n,850\cdot n]$, где $n$ -- размерность задачи (см.
+рисунок \ref{aba:fig4} ).
 
-Let $\vec{\psi_i} = (\psi^1_i,\psi^2_i,\psi^3_i) = (\tilde{l_i},\tilde{p_i},\tilde{l_h})$ -- a set of values of the TabuSearch configuration parameters which was using for solving the problem in the step $i$.
+Пусть $\vec{\psi_i} = (\psi^1_i,\psi^2_i,\psi^3_i) = (\tilde{l_i},\tilde{p_i},\tilde{l_h})$ -- набор значений параметров конфигурации TabuSearch, который использовался для решения задачи на этапе $i$.
 
-Let $[\psi^j_{begin}, \psi^j_{end}]$ -- an optimal interval of the TabuSearch $j$ -parameter.
-(For example: if $j=3$, then: $[\psi^3_{begin}, \psi^3_{end}]= [h_{begin}, h_{end}] = [10, 40] $).
-These intervals are initial data and they are set in the initialization block.
+Пусть $[\psi^j_{begin}, \psi^j_{end}]$ -- оптимальный интервал TabuSearch $j$ -parameter. (For example: if $j=3$, then: $[\psi^3_{begin}, \psi^3_{end}]= [h_{begin}, h_{end}] = [10, 40] $).
+Эти интервалы являются исходные данные, и они установлены в блоке инициализации.
 
-The configuration parameters  $\psi^j_i $ are randomly selected from the interval $[\psi^j_{begin}, \psi^j_{end}]$ according to a distribution law.
-We proposed to build the distribution density $f_j(\psi^j_i)$ of the random variable $\psi^j_i$ as follows.
-At first, we define anchor points:
+Настроечные параметры  $\psi^j_i $ выбираются случайным образом из интервала $[\psi^j_{begin}, \psi^j_{end}]$ в соответствии с законом распределения.
+Мы предложили построить плотность распределения $f_j(\psi^j_i)$ случайной величины $\psi^j_i$следующим образом.
+Во-первых, мы определяем опорные точки:
 
@@
  \end{equation}
- The number of anchor points $n_{\rho}$  is smaller than the set of optimal solutions, because a set of  anchor points are removed the same points $(n_{\rho} \leq  n_u)$.
+Число опорных точек $n_{\rho}$  меньше, чем множество оптимальных решений, так как набор опорных точек, удаляются те же точки $(n_{\rho} \leq  n_u)$.
 
-The main idea is that the distribution density should be larger at those points (parameters $\psi^j_i$) in which high quality solutions than the points at which the low quality of solutions.
-Therefore, we presented the distribution function $f_j(z)$ as a sum of kernel functions:
+Основная идея заключается в том, что плотность распределения должна быть больше в тех точках (параметры  $\psi^j_i$) в которых высококачественные решения, чем точек, в которых низкое качество решений.
+Таким образом, мы представили функцию распределения $f_j(z)$ в виде суммы ядерных функций:
  \begin{equation} \label{dist_density}
@@
  \end{equation}
- where: $K(x)$ -- kernel function, $ \alpha$ -- a normalizing parameter
+ где: $K(x)$ -- ядерная функция, $ \alpha$ -- нормирующий параметр
 
- We have chosen as the kernel the function of parabolic type(known as an Epanechnikov function),  because during the experiments the best results were obtained using this function:
+Мы выбрали в качестве ядра функции параболического типа (известный как функция Епанечникова), так как в ходе экспериментов лучшие результаты были получены с помощью этой функции:
  \begin{equation} \label{dist_density}
@@
 
- The parameter $\alpha_j$ was introduced in the density function to perform the normalization condition:
+Параметр $\alpha_j $ была введена в функции плотности для выполнения условия нормировки:
 \begin{equation} \label{norm_cond}
@@
 
- The parameter $\alpha_j$ can be written in explicit form:
+ Параметр $\alpha_j$можно записать в явном виде:
   \begin{equation} \label{norm_cond}
@@
 
- In figure \ref{aba:fig5} the results of an experiment in which  were presented a dependence of the quality of the solutions obtained by the classical and modified algorithms and  the breaking point time.
-As seen from the graph, with an increase in operating time of the computational scheme, the quality of solutions obtained using the modified algorithm is growing faster than using classical Tabu Search algorithm.
+На рисунке \ref{aba:fig5} результаты эксперимента, в котором были представлены зависимость качества получаемых решений с помощью классических и модифицированных алгоритмов и времени переломной точкой.
+Как видно из графика, с увеличением времени работы расчетной схемы, качество получаемых решений с использованием модифицированного алгоритма растет быстрее, чем при использовании классического алгоритма Табу поиска.
 
@@
  	\caption
- 	{
- 		The dependence of the quality of solutions and  the breaking point time by classic and modified algorithms using $ TimeSpent $ termination strategy
- 	}
+ 	{Зависимость качества растворов и времени Breaking Point с помощью классических и модифицированных алгоритмов с помощью $ TimeSpent $ стратегии завершения}
  	\label{aba:fig5}
@@
 \subsection{Conclusion}
-During the research has been investigated and implemented tabu search algorithm to solve the Static Routing Courier Delivery problem with time constraints.
-Also a modified algorithm was developed  based on the Tabu Search algorithm to improve the quality of the solutions.
-Modified algorithm gave the best solutions in terms of the balance between the number of vehicles and the cost traveled.
-In practice this algorithm can be used for decision support in intelligent systems for improving the quality of customer service and for reducing waiting the time to use vehicle, this will reduce fuel costs and depreciation of transport.
+В ходе исследования был исследован и реализован алгоритм поиска с запретами решить проблему доставки статической маршрутизации Courier с нехваткой времени.
+Кроме того, модифицированный алгоритм был разработан на основе алгоритма Табу поиска, чтобы улучшить качество решений.
+Модифицированный алгоритм дал лучшие решения с точки зрения баланса между количеством транспортных средств и стоимости пройденного.
+На практике этот алгоритм может быть использован для поддержки принятия решений в интеллектуальных системах для повышения качества обслуживания клиентов и сокращения ждут время, чтобы использовать транспортное средство, это позволит сократить расходы на топливо и амортизацию транспорта.
 
-The analysis of the parameters of implemented algorithms allowed us to determine their optimal values for this class of routing problems.
-With the modified algorithm was found solutions of model problems, which in most cases have an acceptable deviation from the global optimum.
+Анализ параметров реализованных алгоритмов позволили определить их оптимальные значения для этого класса задач маршрутизации.
+С помощью модифицированного алгоритма были найдены решения модельных задач, которые в большинстве случаев имеют приемлемое отклонение от глобального оптимума.
 

### 2016-05-15 article.tex
translate to en

@@
 %%%%%%%%%%%%%%% You have to specify name of your paper here
-``Investigation of tuning parameters of tabu search algorithm and its modification  for solving the static Routing Courier Delivery Problem.'' 
+``Investigation of tuning parameters of Tabu Search algorithm and its modification  for solving the Routing Courier Delivery Problem.'' 
 }
@@
 
-The paper considers the routing courier delivery problem with the service time for which the discrete model was constructed and the calculation scheme based on tabu search algorithm.
-The efficiency  of the proposed algorithm‘s scheme has been tested on a large-scale problems, which were generated on the basis of different classes known model routing problems.
+The paper considers the Routing Courier Delivery Problem with the service time for which the discrete model was constructed and the calculation scheme based on the Tabu search algorithm.
+This article provides a method for dynamically adjusting the parameters of Tabu Search algorithm in the process of solving the problem.
+The efficiency  of the proposed algorithm‘s scheme has been tested on large-scale problems, which were generated on the basis of well known model routing problems.
 
@@
 \subsection{Introduction}
-%%%%% Paragraph text. To separate paragraphs use empty lines.
-In the research we consider a static vehicle routing problem with time windows.
-This variant of the routing problem is a common problem in the courier delivery industry.
-Effective solution to this problem affects the quality of services provided by the company, as well as the magnitude of potential losses caused by the excess of the agreed terms of delivery.
-The proposed solution to the problem is based on the automation of the process of planning the route and involves the use of distributed information systems for the possible implementation of these approaches in practice.
-Another important criterion for good transport plan are low transport costs, which affect the cost of services.
+One means of saving resources for transportation of goods is the use of decision support systems in the field of transport logistics.
+One of the key functions of decision support systems in the field of transport logistics is the ability to calculate and construct efficient detour routes for various purposes in the transport network in terms of cost.
+The mathematical formulation of this problem is widely known as the Vehicle Routing Problem (VRP).
+There are several types of VRPs with various additional conditions such as the carrying vehicle`s capacity, time of deliveries and other limitations for a better description of the environment.
 
-The Routing Courier Delivery Problem with time windows requires that a set of vehicles satisfy a collection of customer requests.
-Each client request requires the use of a single vehicle to receive and load a specified amount of goods at one source location(sender) and to deliver them to destination location(receiver).
-All requests must be performed without violating either the vehicle capacity or the customer time window.
-The solution is a master plan that solves the problem, and takes into account the daily events that take place according to plan.
-In this paper, we present a discrete model and a modified tabu search algorithm  to solve the  Courier Delivery Problem with time windows.
+This article is devoted to the courier routing problem, which is a sub-problem of VRP.
+The problem is to find a route to visit a given set of addresses by a number of units of vehicles carrying goods from the sender to the recipient.
 
-In order to validate our strategies decisions technique effectiveness, we have constructed a new set of benchmark problems for the pickup and delivery problem with time windows based on which were generated on the basis of different classes known model routing problems.
-Computational results substantiate the solution quality and efficiency of our modified tabu search algorithm.
+Automation of the route planning process will be relevant for: online stores, large wholesale companies (e.g.
+for distribution of goods that are perishable or optimizing the delivery schemes of goods to destinations) and firms with the aim of organizing restocking in stores.
+A practical application of the problem can be found in the transportation of patients between the buildings of clinic \cite{Beaudry, Kergosien}, the delivery of newspapers and magazines and the delivery of fuel to homes \cite{Sarak}.
+The solution of these problems also has a practical application in the helicopter transportation system of people between offshore oil platforms \cite{Romero}.
+
+In recent years are an increasing amount of people turning to courier services for famous needs.Typically, client orders placed the day before and the dispatcher then has a full list of orders that must be delivered within the deedlines started by the client.
+Attanasio \cite{Attanasio} held a thematic research on courier services which described the advantages of using computer technologies to human dispatching.
+In his work he considered examples of work by eCourier Ltd, a London company that offers express services.
+Their clients are mainly law firms, financial institutions, advertising agencies and other organizations that are interested in the rapid delivery of goods or the sensitive documents.
+Once all of the orders at a particular time have been placed, you must create a route that would satisfy the requirements of the customers.
+Depending on the service level that was specified by the client, the courier can combine several products customers.
+
+Companies that offer courier services frequently have a mixed fleet of vehicles, which consists of bicycles, motorcycles, cars and small trucks.
+Depending on the type of request, the transport arrangement means and time slots will be selected the kind of vehicle that will be able to execute the order within a specified time.
+Researches \cite{Attanasio} showed that the using of computer methods, including optimization algorithms have been very profitable for courier companies.
+
+Thus, the use of an automated system allows to improve the quality of service, reduce delivery time, improve courier efficiency and reduce the cost of shipping, thereby providing a competitive advantage increase.
 
 \subsection{Problem Statement}
-Let $C,dim(C)=n$ is a set of vehicles and $Q,dim(Q)=m$ is a set of requests received from customers at the current time.
+Let $C,dim(C)=n$ be a set of vehicles and $Q,dim(Q)=m$ be a set of requests received from customers at the current time.
 
@@
 
-Let $S$ is a set of senders, ($dim(S) = dim(Q) = m$), $R$ is a set of receivers, $dim(R) = dim(Q) = m$. Then each client request $q \in Q$ includes the following information:\\
+Let $S$ be a set of senders, ($dim(S) = dim(Q) = m$), $R$ is a set of receivers, $dim(R) = dim(Q) = m$. Then each client request $q \in Q$ includes the following information:\\
 $s_q$ -- a sender of client shipment, $s \in S$; \\
@@
 $\vec{P}_{r_q}$ -- receiver position; \\
-$w_q$ -- client shipment that is required to deliver from the sender to the recipient; \\
-$[t_{s}^{q}, t_{s}^{q}+ \Delta{t_{s}^{q}}]$ -- the time window within which worker must to pick up the goods from the sender;\\ 
-$[t_{r}^{q}, t_{r}^{q}+ \Delta{t_{r}^{q}}]$ -- the time window within which worker must deliver the goods to the receiver.
+$w_q$ -- client shipment that is required to be delivered from the sender to the recipient; \\
+$[t_{s}^{q}, t_{s}^{q}+ \Delta{t_{s}^{q}}]$ -- the time window within which the worker must to pick up the goods from the sender;\\ 
+$[t_{r}^{q}, t_{r}^{q}+ \Delta{t_{r}^{q}}]$ -- the time window within which the worker must deliver the goods to the receiver.
 
@@
 \end{equation}
-To estimate the cost of transportation between destinations defined cost function $\Omega$:
+To estimate the cost of transportation between destinations defined by the following cost function $\Omega$:
 \begin{equation}
@@
 
-Necessary to construct the optimal routes of vehicles movement for the transportation of goods from the sender to the receiver for all client requests.
+It is necessary to construct the most optimal routes of vehicles movement for the transportation of goods from the sender to the receiver for all client requests.
 
@@
 \subsection{Discrete model}
-Routing problem can be represented as a directed graph $G=G(V,E)$.
-The set $V=C\cup{S}\cup{R}$ -- nodes of the graph $G$, consisting of elements of vehicles, senders and receivers.
+The Routing Courier Delivery Problem can be represented as a directed graph $G=G(V,E)$.
+The set $V=C\cup{S}\cup{R}$ are nodes of the graph $G$, with elements consists of vehicles, senders and receivers.
 $E$ -- dynamic set of arcs of the graph $G$, such that:
@@
     \begin{cases}
-	  1,&\text{the vehicle $k \in C$ moves from $i$ node to the $j$}\\
+	  1,&\text{the vehicle $k \in C$ moves from the $i$ node to the $j$}\\
 	  0,&\text{otherwise.}
@@
 
-The restriction (\ref{main_cond_1}) prohibits to a node in the graph $G$ has more than one output arc.
-The restriction (\ref{main_cond_2}) prohibits to a node has more than one input arc.
+The restriction (\ref{main_cond_1}) prohibits a node in the graph $G$ of having more than one output arc.
+The restriction (\ref{main_cond_2}) prohibits a node from having more than one input arc.
 The constraint (\ref{main_cond_3}) indicates that the number of input arcs to the node can not be less than the output arcs (this constraint considers the fact that the vehicle can leave the destination only if it has visited this node).
@@
 
-Then next constraints synchronize values of variables $X$ and $\vec{y}$ for each request $q \in Q$ and prohibits the service of receiver before a visit to the sender:
+The next constraints synchronize values of variables $X$ and $\vec{y}$ for each request $q \in Q$ and prohibits the courier service visiting the receiver before meeting with the sender first:
 \begin{align} 
@@
 
-Define restrictions for accounting of vehicle capacity and time windows:
+Defining restrictions for the accounting of vehicle capacity and delivery time windows:
 \begin{align} 
@@
 \subsection{Description of the Algorithm}
-Present the set of matrices $\{X^k\}$ in the form of a vector defined on the hypercube $E^\eta=\{0.1\}^\eta$, where $\eta=n\cdot(2m+1)\cdot 2m$.
+We present the set of matrices $\{X^k\}$ in the form of a vector defined on the hypercube $E^\eta=\{0.1\}^\eta$, where $\eta=n\cdot(2m+1)\cdot 2m$.
 The aim of the task is to minimize the objective function:
@@
 
-Denote by $\delta(\vec{u},\vec{v})$ the Hamming distance between $\vec{u}$ and $\vec{v}$. Through $N_l(\vec{u})$ we denote the neighborhood of a point $\vec{u}$ radius $l$, ie \cite{vrp_shafeyev}:
+Denote by $\delta(\vec{u},\vec{v})$ the Hamming distance between $\vec{u}$ and $\vec{v}$. Through $N_l(\vec{u})$ we denote the neighborhood of a point $\vec{u}$ radius $l$\cite{vrp_shafeyev}:
 \begin{align} 
@@
 When $l=\eta$ a set $ N_l(\vec{u})$ for any vector $\vec{u}$ coincides with the set $E^{\eta}$ and being in this neighborhood vector with a minimum value of the objective function is equivalent to solving the original problem.
-The standard algorithm local search algorithm starts with a randomly selected vector $\vec{u^0}$.
+The classic Local Search algorithm starts with a randomly selected vector $\vec{u^0}$.
 On the $i$ step of the algorithm move from the current vector in the neighboring that the minimum value of the objective function in the neighborhood of the given vector:
@@
 
-The algorithm terminates at a local optimum, when $F(\vec{u^{i+1}})=F(\vec{u^i})$.
-When solving the transportation problem is a typical situation when many local optimums and only one of them is global:
+The algorithm terminates at a local optimum, when $F(\vec{u^{i+1}})=F(\vec{u^i})$. In VRPs there are a typical situation when many local optimums and only one of them is global:
 \begin{align} 
@@
 
-To ensure that the algorithm did not stop in a local minimum, and passed from one local minimum to another, with edges removed the Central point and when searching for the minimum of the following rule applies.
-Let $l=2$ and in the transition from $\vec{u^{i+1}}$ to $\vec{u^i}$ change the values in the coordinate $(u_\lambda^{i}, u_\omega^{i})$.
+To ensure that the algorithm does not stop in a local minimum, and move from one local optimum to another, we must remove the central point  from the current neighborhood and when algorithm search the minimum of the following rule applies.
+Let $l=2$ and current position move from $\vec{u^{i+1}}$ to $\vec{u^i}$ than it changes the values in the coordinate $(u_\lambda^{i}, u_\omega^{i})$.
 The algorithm stores such pair for the last $h$ couple of steps and in the next step prohibits the movement in these directions.
-An ordered list of such pairs
+An ordered list of such pairs:
 \begin{align} 
@@
 \end{align}	
-called the list of prohibitions.
-When building the list of all pairs of different and a pair of $(u_\lambda,u_\omega),\lambda \ne \omega$ does not prohibit the movement of pairs $(u_\lambda,u_\lambda)$ and $(u_\omega,u_\omega)$.
+called the Tabu List.
+When building the list of all pairs and a pair of $(u_\lambda,u_\omega),\lambda \ne \omega$ does not prohibit the movement of pairs $(u_\lambda,u_\lambda)$ and $(u_\omega,u_\omega)$.
 When $l>2$ similarly constructed three coordinates, fours, etc.
 A set of non-restricted vectors denote by $N_l(\vec{u^i},\vec{\phi^i})$.
-In order that the search was effectively, it is advisable to use small values $h$ and to control this parameter in the course of the algorithm.
+In order that the search was efficiently, it is advisable to use small values $h$ and to control this parameter in the course of the algorithm.
 
@@
 \subsection{Initialization}
-To use a search algorithm is restricted by the need to build the initial solution.
-To obtain the initial solution $\vec{u^0}$ you can use a heuristic method to construct the route.
-The essence of these methods is the following.
-Every car has consistently built a route by adding another is not considered clients according to specified rules.
+To use the Tabu Search Algorithm, you must build the initial decision.
+For the initial solution $ \vec{u^0} $, you can use the heuristic method of route constructing\cite {Cordeau2001, Braysy}.
+The essence of these methods is as follows.
+For each car built in series by adding a route client orders on pre-defined rules are not yet considered.
+In this case, to preserve the order of visits vertices pair is meant by the order <<sender-receiver>>.
+This algorithm is fast enough (computational complexity $O(n^3)$ \cite{Babb}), so it is convenient to use for initialization vector $\vec{u^0}$.
 
-The most known method of construction of the route proposed by Solomon in 1987 year and found in the literature as a method of Solomon.
-This algorithm is fast enough (computational complexity $O(n^3)$, therefore, it is useful to initialize a vector $\vec{u^0}$.
+At the beginning of the algorithm going to add a single request to the route in accordance with the time constraints.
+The choice of the first application occurs randomly, or select the one you want to perform before others.
+Then every possible claim $u$ from the set of unchecked orders $ Q $ (the set $ Q'\in Q$) is estimated for inserting it in the beginning of the route, or the end of the route between two adjacent nodes.
+the following criteria are used to select the insertion point:
+\begin{align} 
+\label{insert_kr}
+c(k,v_i,u,v_{i+1}) = \min =
+\begin{cases}
+\min_{q=2,\dots,n^k}(\Omega_{q-1,u}^k+\Omega_{u}^{k}+\Omega_{u,q}^{k}-\mu \Omega_{q-1,q}^k),\\
+\Omega_{k,u}^k+\Omega_{u,q}^k+\Omega_{u}^k\mu \Omega_{q,q}^k), q=1\\
+\Omega_{q,u}^k+\Omega_{u}^k, q=n^k
+\end{cases}
+\end{align} 
+where:\\
+$k$ --  a vehicle route $k\in C$;\\
+$n^k$ - a number of nodes in the route;\\
+$\mu$ --a tuning parameter, $\mu \ge 0 $;\\
+$\Omega_{u}^k=\Omega^k[\vec{P}_{s_q}, \vec{P}_{r_u}]$ -- insert a new request between existing sender and receiver;\\
+$\Omega_{u,q}^k=\Omega^k[\vec{P}_{r_u}, \vec{P}_{q}]$ -- insert a new request between existing sender and receiver to the begin of the route;\\
+$\Omega_{q,u}^k=\Omega^k[\vec{P}_{q}, \vec{P}_{s_u}]$ -- insert a new request between existing sender and receiver to end of the route.
 
-At the beginning of the algorithm, you add one to a customer on the route according to time limits.
-The choice of the first client is randomly or chosen one, you want to serve first.
-Then, each valid client $u$ from the set of outstanding clients $Q^{'}$ (the set $Q^{'}\subset{Q}$) evaluated insertion between two adjacent nodes route.
-For selection of inserts used criterion:
+Adds that the request, which $c(k,v_i,u,v_{i+1})$ It will be minimal:
 \begin{align} 
-& c(v_i,u,v_i+1)=[\min_{p=2,\dots,n^k}(\Omega_{p-1,u}^k+\Omega_{u,p}^k-\mu \Omega_{p-1,p}^k) ]
+\label{kriterij}
+& u^*:c(k,v_i,u^*,v_{i+1})=\min_{u}[c(k,v_i,u,v_{i+1})]
 \end{align} 
-where:
+As a result, $u^*$ is added to the current route $k$ on the most advantageous position.
 
-$k$ - vehicle route $k\in C$;
-$n^k$ - the number of nodes in the route;
-$\mu$ - arbitrary parameter, $\mu \ge 0 $.
+\begin{figure}[h]
+	\center{\includegraphics[width=1\linewidth]{images/pict1}}
+	\caption{An example embodiment of a method of route constructing a single step}
+	\label{ris:pict1}
+\end{figure}
 
-Added the client, which is $c(v_i,u,v_i+1)$ least:
-\begin{align} 
-& u^*:c(v_i,u^*,v_i+1)=min[c(v_i,u,V_i+1)]
-\end{align}
+Fig. \ref {ris:pict1} represented by a weighted graph, which shows the example of the execution of one step by designing routes: \\
+a) the original graph, which shows the options to add new requests to the route; \\
+b) the route, which was built by construction method after first step.
 
-As a result the client $u^*$ is added to the current route $k$ the position between nodes $v_i$ and $v_(i+1)$.
+\textit {Example.} Initially we have the request, which consists of a sender and receiver.
+The value of the cost function $\Omega $ are shown in the graph arcs (Figure \ref {ris:pict1}.).
+You need to add new applications to the route, using the criteria \ref{insert_kr} and \ref{kriterij}.
+The coefficient $\mu=0.5$.
+
+\textit{Solution.}Consider the process of adding new requests.
+Of the two proposed requests $(\{S_2,R_2\},\{S_3,R_3\})$ we choose the one you want to add in the first place.
+Depending on the location where you can add application (beginning, end, middle), select the desired criteria.
+Thus, we find the optimal solution to insert a new order $\{S_2,R_2\}$:\\
+$\min[c(k,S_1, \{S_2, R_2\}, R_1), c(k,\{S_2, R_2\}, S_1, R_1), c(k, S_1, R_1, \{S_2, R_2\})]=\min[3+11+4-0.5 \cdot 8; 17+11+8-0.5 \cdot 20; 2+11]=\min[14;26;13]$\\
+We find the optimal place insert a new request $\{S_3,R_3\}$:\\ 
+$\min[c(k,S_1, \{S_3, R_3\}, R_1), c(k,\{S_3, R_3\}, S_1, R_1), c(k, S_1, R_1, \{S_3, R_3\})]=\min[6+9+12-0.5 \cdot 8; 10+9+13-0.5 \cdot 20; 5+9]=\min[23;22;14]$.\\
+From the solutions found will choose the solution with the lowest cost $c(k,v_i,u,v_{i+1})$.In this example, $c(k, S_1, R_1, \{S_2, R_2\}=13$.
+
+Thus, the new route will consist of initial requests and new requests  $\{S_2,R_2\}$, which will be added to the end of the route.
+
+\textit{Conclusion}: route in $\{S_1,R_1\}$ it is necessary to add  request $\{S_2,R_2\}$, which will be added to the end of the route
+
+Fig. \ref{ris:pict2}presented the options of adding a new request to route $\{S_2,R_2\}$:\\
+a) the option of adding request $\{S_2,R_2\}$ in the end of the route;\\
+b) the option of adding request $\{S_2,R_2\}$ in the beginning of the route;\\
+c) the option of adding request $\{S_2,R_2\}$ between nodes $\{S_1,R_1\}$.
+
+Fig. \ref{ris:pict3} presented the options of adding a new request to route $\{S_3,R_3\}$:\\
+a) the option of adding request $\{S_3,R_3\}$ in the end of the route;\\
+b) the option of adding request $\{S_3,R_3\}$ in the beginning of the route;\\
+c) the option of adding request $\{S_3,R_3\}$ between nodes $\{S_1,R_1\}$.
+
+\begin{figure}[t]
+	\center{\includegraphics[width=1\linewidth]{images/pict2}}
+	\caption{Options added  the request to route $\{S_2,R_2\}$}
+	\label{ris:pict2}
+\end{figure}
+
+\begin{figure}[t]
+	\center{\includegraphics[width=1\linewidth]{images/pict3}}
+	\caption{Options added the request to route $\{S_3,R_3\}$}
+	\label{ris:pict3}
+\end{figure}
 
 \subsection{A Method of Forming Solutions Neighborhood}
-The most difficult part of the local search algorithm with the prohibitions are viewing the surroundings $N_l(\vec{u^i},\vec{\phi^i})$ of the current point $\vec{us^i}$.
-For even larger values of $l$ and the parameter $p=1$ in the outskirts plenty of points.
-But due to the nature of the model's constraints many of them are not the solution.
-Therefore, due to the nature of the considered discrete tasks, we used the following method of determining the area.
-On the $i$ step of the algorithm compiled from routes pair: $\{R_1,R_2\},\{R_1,R_3\},\dots,\{R_1,R_n\},\{R_2,R_1\}, \{R_2,R_3\},\dots,\{R_2,R_n\},\dots,\{R_{n-1},R_{n}\}$.
-Then over each pair of routes is performed in moving requests.
-The resulting displacement vectors $\vec{u}$ are neighborhood $N_l(\vec{u^i},\vec{\phi^i})$.
-For moving  we used change-move  and swap-move operations.
+
+During the work of the Tabu Search Algorithm should be carried out watching the neighborhood $N_{l}(\vec {u^i})$ current point $ \vec{u^i}$.
+In the calculation of the route in the neighborhood gets a sufficiently large number of points.
+But due to the limitations of our model features many of these points is not a solution to the problem.
+Therefore, due to the peculiarities of the considered discrete problem, we used the following method definition of a neighborhood.
+
+Let $x^{k}(v)$ - node index $v$, $ v \in R \cup S $ in the path of the vehicle $ k \in C $.
+Each request consists of pairs $\{S, R \}$, where $S$ - sender, and $R$ - recipient.
+During the formation of the neighborhood, should be preserved the integrity of the route that is performed restrictions (\ref {main_cond_1} -- \ref {main_cond_4}).
+Above route requests $ Route_i $, $ i = 1, \dots n $ are performed permutation operation.
+The findings of the permutations of the vectors $ \vec {u} $ and is a neighborhood of $ N_{l} (\vec{u^i}) $.
+To swap transactions used movement and absorption.
+
+Operation swallowing is shown in Figure \ref {ris: figures}.
+(A), which shows the two routes $Route_1 $ i $ Route_2 $.
+Route $ Route_1 $ consists of catch $ \{C_1, S_1, R_1, S_2, \\ R_2, S_3, R_3 \}$, and the route is $ Route_2 $ node $ \{C_2, S_4, R_4, S_5, R_5, S_6, R_6 \} $.
+During the operation of the absorption $ \{S_1, R_1 \} $ in the route $ Route_1 $, paste this application to route $ Route_2 $ can be performed on $ n + 1 $ positions.
+
+\begin{figure}[h]
+	\center{\includegraphics[width=0.75\linewidth]{images/neighboard.png}}
+	\caption{Example of operations: a) absorption and b) moving}
+	\label{ris:figures}
+\end{figure}
+
+The move operation is shown in figure \ref{ris: figures}.
+(B).
+This figure shows the route $ Route_1 $, which consists of nodes $ \{C_1, S_1, R_1, S_2, R_2, S_3, R_3 \} $.
+To save the route integrity: node $S_i $ (sender) can be visited only to $ R_i $ (recipient), on this basis during the move operation, the node $ S_i $ can move is to satisfy the condition $ 0 < x^{new} ( S_i) < \widetilde{u} (R_i) $, and  node $R_i$ can be moved until the condition is $ n^k \geq x^{new} (R_i)> \widetilde{u} (S_i) $.
 
@@
 
-\end{document}
+\end{document} \grid
+\grid
+\grid

### 2016-05-15 article_ru.tex
translate to en

@@
 В статье рассматривается задача маршрутизации курьерской доставки  с долгим сроком службы, для которой была построена дискретная модель и вычислительная схема, в основу которой был положен алгоритм поиска с запретами.
+В статье предложен способ динамической настройки параметров алгоритма поиска с запретами в процессе решения задачи.
 Эффективность схемы предложенного алгоритма была проверена на задачах большой размерности, которые были составлены на базе известных модельных задачах маршрутизации на транспорте различных классов.
@@
 Каждый вектор $\vec{v} \in N_l(\vec{u^i},\vec{\phi^i})$ с вероятностью $p$ включенных в окрестности $N_l(\vec{u^i},\vec{\phi^i},p)$ независимо от других точек.
-Обратите внимание, что этот набор может быть пустым или содержать только одну точку.
-Общая схема вероятностного алгоритма поиска со списком запретов упоминается как псевдо-кода в алгоритме \\
-представлена на схеме \ref{alg:TabuSearch}.
+Обратим внимание, что этот набор может быть пустым или содержать только одну точку.
+Общая схема вероятностного алгоритма поиска со списком запретов представлен в виде псевдо-кода на схеме \ref{alg:TabuSearch}.
 
@@
 Значения $l,p,h$ являются настроечными параметрами алгоритма.
-Их выбор зависит от размерности задачи.
+Их выбор зависит от размерности и исходных данных задачи.
 
@@
 При больших  $h$ поиск становится неэффективным.
+Ниже будет рассмотрен способ подгонки параметров в процессе решения задачи.
 
-\subsection{Ініціалізація}
+\subsection{Инициализация}
 Для того, чтобы использовать алгоритм поиска с запретами, необходимо построить первоначальное решение.
-Для получения первоначального решения $\vec{u^0}$ можно воспользоваться эвристическим методом конструирования маршрута.
+Для получения первоначального решения $\vec{u^0}$ можно воспользоваться эвристическим методом конструирования маршрута \cite{Cordeau2001,Braysy} .
 Суть этих методов состоит в следующем.
@@
 В нашем случае, для сохранения порядка посещения вершин, под заказом понимается пара "отправитель-получатель".
-
-Самый известный метод конструирования маршрута был предложен Соломоном в 1987 году \cite{Cordeau2001,Braysy} и в литературе встречается как метод Соломона.Данный алгоритм является достаточно быстрым (вычислительная сложность $O(n^3)$ \cite{Babb}), поэтому его удобно использовать для инициализации вектора $\vec{u^0}$.
+Данный алгоритм является достаточно быстрым (вычислительная сложность $O(n^3)$ \cite{Babb}), поэтому его удобно использовать для инициализации вектора $\vec{u^0}$.
 
@@
 Выбор первой заявки происходит случайным образом, или выбирается та, которую необходимо выполнить раньше других.
-Далее каждая возможная заявка  $u$ из множества нерассмотренных заявок $Q$ (множество $Q' \in Q$) оценивается на предмет вставить ее в начало маршрута, конец или между двумя соседними узлами маршрута.
+Далее каждая возможная заявка  $u$ из множества нераcсмотренных заявок $Q$ (множество $Q' \in Q$) оценивается на предмет вставить ее в начало маршрута, конец или между двумя соседними узлами маршрута.
 Для выбора места вставки используется следующий критерий:
@@
 
-На основании результатов экспериментов, описанных в предыдущем параграфе, мы решили установить параметры алгоритма не являются фиксированными значениями, но, как случайной величины заданного периода.
+На основании результатов экспериментов, описанных в предыдущем разделе, мы решили установить параметры алгоритма не являются фиксированными значениями, но, как случайной величины заданного периода.
 Например, размер списка $h$  определяется как интервал  $[10,35]$(see pict.

### 2016-05-16 article.tex
correct engl p 1-7

@@
 The classic Local Search algorithm starts with a randomly selected vector $\vec{u^0}$.
-On the $i$ step of the algorithm move from the current vector in the neighboring that the minimum value of the objective function in the neighborhood of the given vector:
+On the $i$ step of the algorithm the current vector moves to the minimum value of the objective function in the neighborhood:
 \begin{align} 
@@
 
-To ensure that the algorithm does not stop in a local minimum, and move from one local optimum to another, we must remove the central point  from the current neighborhood and when algorithm search the minimum of the following rule applies.
-Let $l=2$ and current position move from $\vec{u^{i+1}}$ to $\vec{u^i}$ than it changes the values in the coordinate $(u_\lambda^{i}, u_\omega^{i})$.
-The algorithm stores such pair for the last $h$ couple of steps and in the next step prohibits the movement in these directions.
+To ensure that the algorithm does not stop at a local minimum and move from one local optimum to another, we must remove the central point  from the current neighborhood and when the algorithm searches for the minimum, the following rule applies.
+Let $l=2$ and the current position moves from $\vec{u^{i}}$ to $\vec{u^{i+1}}$ therefore changing the values in the coordinate $(u_\lambda^{i}, u_\omega^{i})$.
+The algorithm stores such pairs for the last $h$ amount of steps and in the next step prohibits the movement in these directions.
 An ordered list of such pairs:
@@
 \end{align}	
-called the Tabu List.
-When building the list of all pairs and a pair of $(u_\lambda,u_\omega),\lambda \ne \omega$ does not prohibit the movement of pairs $(u_\lambda,u_\lambda)$ and $(u_\omega,u_\omega)$.
-When $l>2$ similarly constructed three coordinates, fours, etc.
-A set of non-restricted vectors denote by $N_l(\vec{u^i},\vec{\phi^i})$.
-In order that the search was efficiently, it is advisable to use small values $h$ and to control this parameter in the course of the algorithm.
+is called the Tabu List.
+The pair $(u_\lambda,u_\omega),\lambda \ne \omega$ does not prohibit the movement of pairs $(u_\lambda,u_\lambda)$ and $(u_\omega,u_\omega)$.
+When $l > 2$, the Tabu List is created accordingly for 3 coordinates, 4 coordinates, etc.
+A set of non-restricted vectors are denoted by $N_l(\vec{u^i},\vec{\phi^i})$.
+In order for the search to be efficient, it is advisable to use small values of $h$ and to control this parameter throughout the algorithm.
 
@@
 	\textbf{function TabuSearch}($u^0,l,p,h$) \\
-	// Initialize variables:	\\
+	// Initial variables:	\\
 	$u^{opt} = u^0$
@@
 
-Denote by $N_l(\vec{u^i},\vec{\phi^i},p)$ probabilistic edge that goes from a deterministic $N_l(\vec{u^i},\vec{\phi^i})$ as follows.
-Each vector $\vec{v} \in N_l(\vec{u^i},\vec{\phi^i})$ with probability $p$ included in the neighborhood $N_l(\vec{u^i},\vec{\phi^i},p)$ regardless of other points.
+Denoted by $N_l(\vec{u^i},\vec{\phi^i},p)$ a probabilistic neighborhood which stems from a deterministic $N_l(\vec{u^i},\vec{\phi^i})$ as follows; each vector $\vec{v} \in N_l(\vec{u^i},\vec{\phi^i})$ with probability $p$ is included in the neighborhood $N_l(\vec{u^i},\vec{\phi^i},p)$ regardless of other points.
 Note that this set may be empty or contain only one point.
-The General scheme of the probabilistic search algorithm with a list of prohibitions referred to as pseudo-code in algorithm \\
-scheme \ref{alg:TabuSearch}.
+The General scheme of the probabilistic Tabu Search algorithm in scheme \ref{alg:TabuSearch} is referred to as the pseudo-code.
 
-As the stopping criterion is based on the total number of steps $N_stop$, in the course of which does not change the value $F_{opt}$.
+The stopping criterion is based on the total number of steps $N_{stop}$, which does not change the value $F_{opt}$.
 Values $l,p,h$ are the control parameters of the algorithm.
-Their choice depends on the problem dimension.
+Their choice depends on the problems dimension.
 
-In the presented scheme, it was assumed that the value of $h$ (the dimension of the deny list) does not change in the course of the algorithm.
-This creates certain difficulties in the implementation of the scheme, as it is unknown how long is to take a list of prohibitions.
-At small $h$ the algorithm can start walking cycle.
+In the presented scheme, it was assumed that the value of $h$ (the dimension of the Tabu List) does not change throughout the course of the algorithm.
+This creates certain difficulties in the implementation of the scheme, as it is unknown how long the Tabu List size should be.
+At small $h$ the algorithm can get into an infinite loop.
 At large $h$ the search becomes inefficient.
-One of the simple rules regulating the length of the list of prohibitions is the following.
-
-If at the next step of the algorithm, the edge  $N_l(\vec{u^i},\vec{\phi^i},p)$ turned out to be empty, it is added to the list of fictitious vector $\vec{0}$.
-This vector nothing forbids, but reduces the number of bans per unit.
 
 \subsection{Initialization}
-To use the Tabu Search Algorithm, you must build the initial decision.
+To use the Tabu Search Algorithm, you must first build the initial solution.
 For the initial solution $ \vec{u^0} $, you can use the heuristic method of route constructing\cite {Cordeau2001, Braysy}.
 The essence of these methods is as follows.
-For each car built in series by adding a route client orders on pre-defined rules are not yet considered.
-In this case, to preserve the order of visits vertices pair is meant by the order <<sender-receiver>>.
-This algorithm is fast enough (computational complexity $O(n^3)$ \cite{Babb}), so it is convenient to use for initialization vector $\vec{u^0}$.
+According to the rules of the algorithm routes for each vehicle are constructed as client requests come in order to ensure efficiency.
+To preserve the order of visits of the nodes by the vehicles in the pair are ordered ''sender-receiver''.This algorithm is fast enough (computational complexity $O(n^3)$ \cite{Babb}), so it is convenient to use for the initialization of the vector $\vec{u^0}$.
 
-At the beginning of the algorithm going to add a single request to the route in accordance with the time constraints.
-The choice of the first application occurs randomly, or select the one you want to perform before others.
-Then every possible claim $u$ from the set of unchecked orders $ Q $ (the set $ Q'\in Q$) is estimated for inserting it in the beginning of the route, or the end of the route between two adjacent nodes.
-the following criteria are used to select the insertion point:
+At the beginning of the algorithm a single request is added to each route in accordance with the time constraints.
+The first node(sender) is chosen randomly or is selected the one you want to perform before others.
+Then every possible request $u$ from the set of unchecked requests $ Q' $ (the set $ Q'\in Q$) is estimated by inserting it in the beginning of the route, or the end of the route between two adjacent nodes.
+the following criteria is used to select the insertion point:
 \begin{align} 
@@
 $\Omega_{u}^k=\Omega^k[\vec{P}_{s_q}, \vec{P}_{r_u}]$ -- insert a new request between existing sender and receiver;\\
-$\Omega_{u,q}^k=\Omega^k[\vec{P}_{r_u}, \vec{P}_{q}]$ -- insert a new request between existing sender and receiver to the begin of the route;\\
-$\Omega_{q,u}^k=\Omega^k[\vec{P}_{q}, \vec{P}_{s_u}]$ -- insert a new request between existing sender and receiver to end of the route.
+$\Omega_{u,q}^k=\Omega^k[\vec{P}_{r_u}, \vec{P}_{q}]$ -- insert a new request between existing sender and receiver to the begining of the route;\\
+$\Omega_{q,u}^k=\Omega^k[\vec{P}_{q}, \vec{P}_{s_u}]$ -- insert a new request between existing sender and receiver to the end of the route.
 
-Adds that the request, which $c(k,v_i,u,v_{i+1})$ It will be minimal:
+Adds that request, which $c(k,v_i,u,v_{i+1})$ will be minimal:
 \begin{align} 
@@
 \end{align} 
-As a result, $u^*$ is added to the current route $k$ on the most advantageous position.
+As a result, $u^*$ is added to the current route $k$ in the most advantageous position.
 
@@
 
-Fig. \ref {ris:pict1} represented by a weighted graph, which shows the example of the execution of one step by designing routes: \\
-a) the original graph, which shows the options to add new requests to the route; \\
-b) the route, which was built by construction method after first step.
+Fig. \ref {ris:pict1} represented by a weighted graph, which shows the example of the execution of one step of designing routes: \\
+a) the original graph, which shows options to add new requests to the route; \\
+b) the route, which was built by the construction method after first step.
 
-\textit {Example.} Initially we have the request, which consists of a sender and receiver.
+\textit {Example.} Initially we have a request, which consists of a sender and receiver.
 The value of the cost function $\Omega $ are shown in the graph arcs (Figure \ref {ris:pict1}.).
-You need to add new applications to the route, using the criteria \ref{insert_kr} and \ref{kriterij}.
+You need to add new requests to the route, using the criteria \ref{insert_kr} and \ref{kriterij}.
 The coefficient $\mu=0.5$.
@@
 \textit{Solution.}Consider the process of adding new requests.
-Of the two proposed requests $(\{S_2,R_2\},\{S_3,R_3\})$ we choose the one you want to add in the first place.
+Of the two proposed requests $(\{S_2,R_2\},\{S_3,R_3\})$ it chooses the one which needs to be added firstly.
 Depending on the location where you can add application (beginning, end, middle), select the desired criteria.
@@
 \subsection{Analysis of the algorithm parameters}
-The Tabu Search algorithm has two main parameters: $h$ -- size of the list of prohibitions and $N_{neighbors}$ -- accepted count limit of size of neighbordhood $N_l(\vec{u})$.
+The Tabu Search algorithm has two basic parameters: $h$ -- size of the list of prohibitions and $N_{neighbors}$ -- accepted count limit of size of neighbordhood $N_l(\vec{u})$.
 For the experiments, we constructed the test tasks on the basis of the capacity routing problems developed by Breedam, Fisher, Christofides and Eilon.
@@
 The first generalized diagram is as follows.
-First, the set of solutions is constructed by Solomon.
-Then every solution is improved by tabu search and choose the best solution according to the objective function.
+First, the set of solutions is constructed by route construction algorithm.
+Then every solution is improved by Tabu Search and choose the best solution according to the objective function.
 
-Modification of the scheme is based on the hypothesis of "On a large valley"\cite{opt_Kochetov}.
+Modification of the scheme is based on the hypothesis of ''On a large valley''\cite{opt_Kochetov}.
 According to this hypothesis, the average local optima are located much closer to the global than a randomly chosen point.
@@
 \ref{aba:fig2} and  pict.
-\ref{aba:fig3}) and the neighbordhood size $l=N_{neighbordhood}$ is defined as the interval $[600\cdot n,850\cdot n]$, where $n$ -- problem size (see pict.
+\ref{aba:fig3}) and the neighbordhood size $l=N_{neighbordhood}$ is defined as the interval $[6\cdot n,8.5\cdot n]$, where $n$ -- problem size (see pict.
 \ref{aba:fig4} ).
 
-Let $\vec{\psi_i} = (\psi^1_i,\psi^2_i,\psi^3_i) = (\tilde{l_i},\tilde{p_i},\tilde{l_h})$ -- a set of values of the TabuSearch configuration parameters which was using for solving the problem in the step $i$.
+Let $\vec{\psi_i} = (\psi^1_i,\psi^2_i,\psi^3_i) = (\tilde{l_i},\tilde{p_i},\tilde{h_i})$ -- a set of values of the TabuSearch configuration parameters which was using for solving the problem in the step $i$.
 
 Let $[\psi^j_{begin}, \psi^j_{end}]$ -- an optimal interval of the TabuSearch $j$ -parameter.
-(For example: if $j=3$, then: $[\psi^3_{begin}, \psi^3_{end}]= [h_{begin}, h_{end}] = [10, 40] $).
+For example: if $j=3$, then: $[\psi^3_{begin}, \psi^3_{end}]= [h_{begin}, h_{end}] = [10, 40] $.
 These intervals are initial data and they are set in the initialization block.
@@
 
-		// Make solution using euristics construction algorithms \\
+		// Make solution using heuristics construction algorithms \\
 		$u^0 = constructSolution(U^{opt} ,n^{min}_s)$
@@
  	{
- 		The dependence of the quality of solutions and  the breaking point time by classic and modified algorithms using $ TimeSpent $ termination strategy
+ 		The dependence of the quality of solutions and  the breaking point time by classic algorithm('$\times$' mark) with random selection tuning parameters and modified('o' mark) algorithms using $ TimeSpent $ termination strategy.
  	}
@@
 \subsection{Conclusion}
-During the research has been investigated and implemented tabu search algorithm to solve the Static Routing Courier Delivery problem with time constraints.
-Also a modified algorithm was developed  based on the Tabu Search algorithm to improve the quality of the solutions.
-Modified algorithm gave the best solutions in terms of the balance between the number of vehicles and the cost traveled.
-In practice this algorithm can be used for decision support in intelligent systems for improving the quality of customer service and for reducing waiting the time to use vehicle, this will reduce fuel costs and depreciation of transport.
-
+During the research has been investigated and implemented Tabu Search algorithm to solve the Routing Courier Delivery problem with time windows.
+Also a modified algorithm was developed  based on the Tabu Search Algorithm to improve the quality of solutions.
+Modified algorithm was given the best solutions in terms of the balance between the number of vehicles and the cost traveled.
+In practice this algorithm can be used in intelligent systems for decision support, improving the quality of customer service and reducing waiting time, this will reduce fuel costs and depreciation of transport.
 The analysis of the parameters of implemented algorithms allowed us to determine their optimal values for this class of routing problems.
@@
 %In brackets we writes the name which is used for referencing.
-
 \bibitem{RCDS} F.~Ordonez, Chen Wang, A New Approach for Routing Courier Delivery Services //METRANS Transportation Center:University of Southern California
@@
 
-\bibitem{Goldberg_Kennedy} V.~Goldberg, R.~Kennedy, An Efficient cost scaling algorithm for the assignment problem, Math. Program., 1995, p. 153--177.
+\bibitem{Goldberg_Kennedy} V.~Goldberg, R.~Kennedy, An Efficient cost scaling algotirhm for the assignment problem, Math. Program., 1995, p. 153--177.
 
@@
 
+\bibitem{Attanasio} A.~Attanasio, J.~Bregman,G.~Ghiani, E.~Manni.
+Real-time fleet management at Ecourier Ltd.
+Dynamic Fleet Management, volume 38 of Operations Research/Computer Science Interfaces, chapter 10, p.219--238, 2007.
+
+\bibitem{Beaudry} A.~Beaudry, G.~Laporte, T.~Melo, Nickel. Dynamic transportation of patients in hospitals. Berichte des Fraunhofer ITWM, Nr. 104 :p.1--34, 2010.
+
+\bibitem{Romero} M.Romero, L.Sheremetov and A.Soriano.
+A genetic algorithm for the pickup and delivery problem: An application to the helicopter offshore transportation.
+In Theoretical Advances and Applications of Fuzzy Logic and Soft Computing, volume 42 of Advances in Soft Computing, 2007, p.
+35--44.
+
+\bibitem{Kergosien} M.Romero, L.Sheremetov and A.Soriano.
+Yannick Kergosien, Christophe Lent, D.
+Piton, Jean-Charles Billaut.
+A tabu search heuristic for a dynamic transportation problem of patients between care units.
+29 pages.
+2010.
+
+\bibitem{Sarak} H.~Sarak, A.~Satman. The degree-day method to estimate the residential heating natural gas consumption in Turkey: a case study. Pergamon, Energy 28 (2003) 929--939.
+
 \end{thebibliography}
